---
title: "Mice"
subtitle: "Fibroblasts dataset"
author: "Audrey"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <<- Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```

The goal of this script is to run analysis the fibroblasts dataset, to understand differences between pNF fibroblasts and dyNF fibroblasts.

```{r library}
library(patchwork)
library(ggplot2)
library(dplyr)

.libPaths()
```

# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "mice_fibroblasts"
out_dir = "."
n_threads = 3 # for tSNE
percent_mt_thresh = 7
```

We load the fibroblasts dataset :

```{r load_sobj}
sobj = readRDS(paste0(out_dir, "/", save_name, "_sobj.rds"))
name2D = "mnn_44_umap"

sobj
```


We load the sample information :

```{r custom_palette_sample, fig.width = 8, fig.height = 8}
sample_info = readRDS(paste0(out_dir, "/../1_metadata/mice_sample_info.rds"))
sample_info = sample_info %>%
  dplyr::filter(project_name %in% unique(sobj$project_name)) %>%
  base::droplevels()

graphics::pie(rep(1, nrow(sample_info)),
              col = sample_info$color,
              labels = sample_info$project_name)
```

Here are custom colors for each cell type :

```{r color_markers, fig.width = 10, fig.height = 1, class.source = "fold-hide"}
color_markers = readRDS(paste0(out_dir, "/../1_metadata/mice_color_markers.rds"))

data.frame(cell_type = names(color_markers),
           color = unlist(color_markers)) %>%
  ggplot2::ggplot(., aes(x = cell_type, y = 0, fill = cell_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers), breaks = names(color_markers)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank())
```

# Visualization

## Global

We visualize clustering and sample on the projection :

```{r see_sobj, fig.width = 9, fig.height = 5}
project_name = Seurat::DimPlot(sobj, group.by = "project_name",
                               reduction = name2D) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$project_name) +
  Seurat::NoAxes() + ggplot2::ggtitle("Sample of origin") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.position = "none")

clusters = Seurat::DimPlot(sobj, group.by = "seurat_clusters",
                           reduction = name2D, label = TRUE) +
  Seurat::NoLegend() +
  Seurat::NoAxes() + ggplot2::ggtitle("Clsuters ID") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

project_name | clusters
```

We visualize quality-related metrics :

```{r see_qc, fig.width = 15, fig.height = 4}
plot_list = lapply(c("nFeature_RNA", "log_nCount_RNA", "percent.mt", "percent.rb"),
                   FUN = function(one_metric) {
                     p = Seurat::FeaturePlot(sobj, features = one_metric, reduction = name2D) +
                       ggplot2::theme(aspect.ratio = 1) +
                       Seurat::NoAxes() +
                       ggplot2::scale_color_gradientn(colors  = aquarius::color_gene)
                     return(p)
                   })

patchwork::wrap_plots(plot_list, nrow = 1)
```

dyNF-associated fibroblasts express much more genes than in pNF.

```{r vln_feature, fig.width = 8, fig.height = 5}
Seurat::VlnPlot(sobj, features = "nFeature_RNA", pt.size = 0.5,
                group.by = "sample_identifier") +
  ggplot2::scale_fill_manual(values = sample_info$color) +
  Seurat::NoLegend() +
  ggplot2::theme(axis.title.x = element_blank())
```


## Gene expression

We will look those genes :

```{r fibroblats_gene}
peri = c("Sfrp5", "Itgb4", "Wnt6", "Lypd2", "Krt19", "Cav2", "Slc6a13",
         "Stxbp6", "Sdpr", "Foxd1", "Tenm2", "Mustn1", "Moxd1", "Plcd3", "Frmd4b")
epi = c("Clec3b", "Gdf10", "Dpt", "Pcolce2", "Ly6c1", "Comp", "Bmp4",
        "Gpnmb", "Clec11a", "Cfh", "Ogn", "Fbln7", "Fmod", "Crispld2",
        "Cxcl13", "Dpp4", "Cd55", "Adamts5", "Cd248")
endo = c("Enpp2", "Fam213a", "Myoc", "Cyp26b1", "Aspa", "P2ry14", "Nipal1",
         "Nt5e", "Mylk", "Cdh19", "Atp1a2", "Jam2", "Acot1", "Kif21a",
         "Foxd1", "1500009L16Rik", "Hlf", "Sox8", "Ptx3")
caf = c("Pdpn", "S100a4", "Fap", "Itgb1", "Acta2", "Pdgfrb", "Pdgfra",
        "Des", "Fbln1", "Clu", "Thy1", "Cav1", "Mme", "C5ar2")
```

### pNF fibroblasts

We make the plot for **epineurial**, **endoneurial** and **perineurial** fibroblasts :

```{r plot_genes_epi_peri_endo, fig.width = 12, fig.height = 18}
plot_list = lapply(c(peri, epi, endo),
                   FUN = function(one_gene) {
                     p = Seurat::FeaturePlot(sobj, features = one_gene, reduction = name2D) +
                       ggplot2::theme(aspect.ratio = 1) +
                       Seurat::NoAxes() +
                       ggplot2::scale_color_gradientn(colors  = aquarius::color_gene)
                     return(p)
                   })

patchwork::wrap_plots(plot_list, ncol = 6)
```

### dyNF fibroblasts

#### CAFs

Markers publication source : [https://onlinelibrary.wiley.com/doi/full/10.1111/cas.14537](https://onlinelibrary.wiley.com/doi/full/10.1111/cas.14537)

```{r plot_genes_caf, fig.width = 8, fig.height = 8}
plot_list = lapply(caf,
                   FUN = function(one_gene) {
                     p = Seurat::FeaturePlot(sobj, features = one_gene, reduction = name2D) +
                       ggplot2::theme(aspect.ratio = 1) +
                       Seurat::NoAxes() +
                       ggplot2::scale_color_gradientn(colors  = aquarius::color_gene)
                     return(p)
                   })

patchwork::wrap_plots(plot_list, ncol = 4)
```

#### Repair FB

Markers publication source : [https://www.cell.com/cell-stem-cell/pdf/S1934-5909(18)30535-6.pdf](https://www.cell.com/cell-stem-cell/pdf/S1934-5909(18)30535-6.pdf)
Postn has been identified by DE within this published dataset.

```{r plot_genes_repair, fig.width = 8, fig.height = 8}
plot_list = lapply(c("Dlk1", "Tnc", "Plagl1", "Mest",          # from publication
                     "Bmpr1b", "Cilp", "Runx2", "Ptn", "Sox9",  # from publication
                     "Postn"), # own analysis
                   FUN = function(one_gene) {
                     p = Seurat::FeaturePlot(sobj, features = one_gene,
                                             reduction = name2D, pt.size = 0.5) +
                       ggplot2::theme(aspect.ratio = 1) +
                       Seurat::NoAxes() +
                       ggplot2::scale_color_gradientn(colors  = aquarius::color_gene)
                     return(p)
                   })

patchwork::wrap_plots(plot_list, ncol = 3)
```


# Differential expression

We identify DE genes between pNF and dyNF :

```{r mark}
Seurat::Idents(sobj) = sobj$tumor_type
mark = Seurat::FindMarkers(sobj, ident.1 = "pNF")
mark$diff = abs(mark$pct.1 - mark$pct.2)

mark = mark %>% dplyr::arrange(-diff)

mark %>%
  dplyr::top_n(., n = 40, wt = diff) %>%
  dplyr::arrange(avg_logFC)
```

We set markers for dyNF-associated fibroblasts :

```{r dynf_markers}
dynf_markers = mark %>%
  dplyr::filter(pct.1 < 0.3 & pct.2 > 0.7) %>%
  dplyr::top_n(., n = 20, wt = -avg_logFC) %>%
  rownames()
```

We visualize those markers :

```{r plot_dynf_markers, fig.width = 12, fig.height = 8}
plot_list = lapply(dynf_markers,
                   FUN = function(one_gene) {
                     p = Seurat::FeaturePlot(sobj, features = one_gene,
                                             reduction = name2D, pt.size = 0.5) +
                       ggplot2::theme(aspect.ratio = 1) +
                       Seurat::NoAxes() +
                       ggplot2::scale_color_gradientn(colors  = aquarius::color_gene)
                     return(p)
                   })

patchwork::wrap_plots(plot_list, ncol = 5)
```


# Fibroblast subtypes annotation

We run a cell type annotation pipeline :

```{r annotation}
sobj = aquarius::cell_annot_custom(sobj,
                                   newname = "FB_type",
                                   markers = list("periFb" = peri,
                                                  "endoFb" = endo,
                                                  "epiFb" = epi,
                                                  "CAF" = dynf_markers),
                                   use_negative = TRUE,
                                   add_score = TRUE,
                                   verbose = TRUE)
```

We define colors for fibroblast subpopulations :

```{r color_fibro}
color_fibro = setNames(aquarius::c30_palette[c(4, 19, 18, 23) + 1],
                       nm = c("endoFb", "epiFb", "periFb", "CAF"))
color_fibro
```


We can visualize this :

```{r FB_type, fig.width = 12, fig.height = 6}
plot_list = lapply(grep(colnames(sobj@meta.data), pattern = "^score", value = TRUE),
                   FUN = function(one_gene) {
                     p = Seurat::FeaturePlot(sobj, features = one_gene,
                                             reduction = name2D, pt.size = 0.5) +
                       ggplot2::theme(aspect.ratio = 1) +
                       Seurat::NoAxes() +
                       ggplot2::scale_color_gradientn(colors  = aquarius::color_gene)
                     return(p)
                   })
plot_list[[length(plot_list) + 1]] = Seurat::DimPlot(sobj, group.by = "FB_type",
                                                     cols = color_fibro,
                                                     reduction = name2D) +
  ggplot2::labs(title = "Fibroblast annotation") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.position = "top",
                 legend.justification = "center",
                 legend.direction = "horizontal") +
  Seurat::NoAxes()

patchwork::wrap_plots(plot_list) +
  patchwork::plot_layout(design = "EEAB\nEECD")
```

What is the proportion of fibroblast type, by tumor type ?

```{r barplot_FB_type, fig.width = 5, fig.height = 6}
df_prop = table(sobj$FB_type,
                sobj$tumor_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("FB_type", "tumor_type", "nb_cells"))

aquarius::plot_barplot(df_prop,
                       x = "tumor_type",
                       y = "nb_cells",
                       fill = "FB_type",
                       position = ggplot2::position_fill()) +
  ggplot2::scale_fill_manual(values = color_fibro,
                             breaks = names(color_fibro))
```


# Save

We save this object :

```{r save_sobj, fig.width = 12, fig.height = 12}
saveRDS(sobj, file = paste0(out_dir, "/", save_name, "_annotated_sobj.rds"))
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```

