---
title: "MPNST project"
subtitle: "Zoom in mice TME"
author: "Audrey"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

```{r clean, echo = FALSE}
rm(list = ls())
```

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <= Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


This file is used to analyse the immune cells dataset.

```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
library(org.Mm.eg.db)
library(ComplexHeatmap)
```


# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "all_mice"
out_dir = "/home/aurelien/Documents/Audrey/git_analysis/mpnst_project/8_combined_mice/"
```


We load the dataset :

```{r load_sobj}
sobj = readRDS(paste0(out_dir, "/", save_name, "_sobj.rds"))
sobj
```

and sample information :

```{r load_sample_info}
sample_info = readRDS(paste0(out_dir, "/", save_name, "_sample_info.rds")) 
sample_info
```

We load specific colors for cell types :

```{r color_markers}
color_markers = readRDS(paste0(out_dir, "/../1_metadata/datasets_color_markers.rds"))
```


This is the projection of interest :

```{r name2D}
name2D = "harmony_38_tsne"
```


# Visualization

## Cell type

We visualize cell type annotation :

```{r cell_type, fig.width = 6, fig.height = 4}
Seurat::DimPlot(sobj, reduction = name2D, group.by = "cell_type") +
  ggplot2::scale_color_manual(values = color_markers,
                              breaks = names(color_markers)) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1)
```

We make a barplot to compare TME composition between fl/fl and fl/- :

```{r barplot_microenv, fig.width = 12, fig.height = 8}
tumor_cells = readRDS(paste0(out_dir, "/../7_tumor_cells_mice/mice_tumor_cells_sobj.rds")) %>%
  colnames() %>%
  stringr::str_split(., pattern = "_") %>%
  lapply(., `[[`, 3) %>% unlist()

df = sobj@meta.data %>%
  dplyr::mutate(cell_barcode = rownames(.)) %>%
  dplyr::mutate(cell_barcode = stringr::str_split(cell_barcode, pattern = "_") %>%
                  lapply(., `[[`, 1) %>% unlist()) %>%
  dplyr::filter(!(cell_barcode %in% tumor_cells)) %>%
  dplyr::filter(cell_type != "tumor cells") %>% ############ mis-annotation + high-mito
  dplyr::select(project_name, cell_type, Seurat.Phase) %>%
  dplyr::mutate(celltype_cycle = paste(cell_type, Seurat.Phase, sep = "zzz"))
df = table(df$project_name, df$celltype_cycle) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("project_name", "celltype_cycle", "nb_cells")) %>%
  dplyr::mutate(cell_type = stringr::str_split(celltype_cycle, pattern = "zzz") %>%
                  lapply(., `[[`, 1) %>% unlist()) %>%
  dplyr::mutate(cycling_status = stringr::str_split(celltype_cycle, pattern = "zzz") %>%
                  lapply(., `[[`, 2) %>% unlist())

aquarius::plot_barplot(df,
                       x = "project_name",
                       y = "nb_cells",
                       fill = "cell_type",
                       pattern = "cycling_status",
                       position = ggplot2::position_fill()) +
  ggplot2::scale_fill_manual(breaks = names(color_markers),
                             values = unlist(color_markers),
                             name = "Cell type") +
  ggplot2::theme(axis.title.y = element_blank())
```

We make the same barplot, only for cycle status :

```{r barplot_microenv_cycling, fig.width = 12, fig.height = 8}
df = sobj@meta.data %>%
  dplyr::mutate(cell_barcode = rownames(.)) %>%
  dplyr::mutate(cell_barcode = stringr::str_split(cell_barcode, pattern = "_") %>%
                  lapply(., `[[`, 1) %>% unlist()) %>%
  dplyr::filter(!(cell_barcode %in% tumor_cells)) %>%
  dplyr::filter(cell_type != "tumor cells") %>% ############ mis-annotation + high-mito
  dplyr::select(project_name, cell_type, Seurat.Phase) %>%
  dplyr::mutate(celltype_cycle = paste(cell_type, Seurat.Phase, sep = "zzz"))
df = table(df$project_name, df$Seurat.Phase) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("project_name", "cycling_status", "nb_cells"))

aquarius::plot_barplot(df,
                       x = "project_name",
                       y = "nb_cells",
                       fill = "cycling_status",
                       position = ggplot2::position_fill()) +
  ggplot2::scale_fill_manual(values = aquarius::gg_color_hue(3),
                             name = "Cycling status") +
  ggplot2::theme(axis.title.y = element_blank())
```


We make the same only for tumor cells (mis-annotation seeable) :

```{r barplot_tumorcells, fig.width = 12, fig.height = 8}
df = sobj@meta.data %>%
  dplyr::mutate(cell_barcode = rownames(.)) %>%
  dplyr::mutate(cell_barcode = stringr::str_split(cell_barcode, pattern = "_") %>%
                  lapply(., `[[`, 1) %>% unlist()) %>%
  dplyr::filter(cell_barcode %in% tumor_cells | cell_type == "tumor cells") %>%
  dplyr::select(project_name, cell_type, Seurat.Phase) %>%
  dplyr::mutate(celltype_cycle = paste(cell_type, Seurat.Phase, sep = "zzz"))
df = table(df$project_name, df$celltype_cycle) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("project_name", "celltype_cycle", "nb_cells")) %>%
  dplyr::mutate(cell_type = stringr::str_split(celltype_cycle, pattern = "zzz") %>%
                  lapply(., `[[`, 1) %>% unlist()) %>%
  dplyr::mutate(cycling_status = stringr::str_split(celltype_cycle, pattern = "zzz") %>%
                  lapply(., `[[`, 2) %>% unlist())

aquarius::plot_barplot(df,
                       x = "project_name",
                       y = "nb_cells",
                       fill = "cell_type",
                       pattern = "cycling_status",
                       position = ggplot2::position_fill()) +
  ggplot2::scale_fill_manual(breaks = names(color_markers),
                             values = unlist(color_markers),
                             name = "Cell type") +
  ggplot2::theme(axis.title.y = element_blank())
```

We visualize proliferation on the projection :

```{r see_birc5, fig.width = 12, fig.height = 8}
source("/home/aurelien/Documents/Audrey/git_aquarius/R/plot_split_dimred.R")

plot_list = plot_split_dimred(sobj, reduction = name2D,
                              split_by = "project_name",
                              split_color = sample_info$color,
                              color_by = "Birc5")
patchwork::wrap_plots(plot_list, nrow = 3) + patchwork::plot_layout(guides = "collect")
```


We make a ratio of cycling cells, for each dataset, between tumor cells and micro-environment. We use cell type annotation to compute the proportion. We suppose that mis-annotation of some cells will not drive the result, so if there is a difference between datasets, this is due to the well annotated cells.

```{r ratio_cycling, fig.width = 8, fig.height = 8, class.source = "fold-hide"}
sobj$is_cycling = ifelse(sobj$Seurat.Phase == "G1",
                         yes = "no", no = "yes")
sobj$is_tumor_cells = ifelse(sobj$cell_type == "tumor cells",
                             yes = "yes", no = "no")

## As proportion
# Proportion of cycling cells in TME
df_tme = sobj@meta.data %>%
  dplyr::filter(is_tumor_cells == "no") %>%
  dplyr::select(project_name, is_cycling)
df_tme = table(df_tme$project_name, df_tme$is_cycling) %>%
  prop.table(., margin = 1) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("project_name", "is_cycling", "prop_tme")) %>%
  dplyr::filter(is_cycling == "yes") %>%
  dplyr::select(project_name, prop_tme)

# Proportion of cycling cells in tumor cells
df_tum = sobj@meta.data %>%
  dplyr::filter(is_tumor_cells == "yes") %>%
  dplyr::select(project_name, is_cycling)
df_tum = table(df_tum$project_name, df_tum$is_cycling) %>%
  prop.table(., margin = 1) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("project_name", "is_cycling", "prop_tum")) %>%
  dplyr::filter(is_cycling == "yes") %>%
  dplyr::select(project_name, prop_tum)

# Join and make ratio
df = dplyr::left_join(df_tme, df_tum, by = "project_name") %>%
  dplyr::mutate(ratio = prop_tum / prop_tme) %>%
  dplyr::left_join(., y = sample_info[, c("project_name", "genotype")],
                   by = "project_name")

ggplot2::ggplot(df, aes(x = prop_tum, y = prop_tme, label = project_name,
                        col = project_name)) +
  ggplot2::geom_abline(intercept = 0, slope = 1, col = "gray80", lty = 2) +
  ggplot2::geom_point(aes(fill = genotype), pch = 21, stroke = 0, size = 3) +
  ggrepel::geom_label_repel(min.segment.length = 0, segment.color = "black", label.size = 0) +
  # ggplot2::lims(x = c(0,1), y = c(0,1)) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$project_name,
                              guide = "none") +
  ggplot2::labs(x = "Proportion of cycling cells in tumor cells",
                y = "Proportion of cycling cells in tumor micro-environment") +
  ggplot2::theme_classic()

## As number
# Number of cycling cells in TME
df_tme = sobj@meta.data %>%
  dplyr::filter(is_tumor_cells == "no") %>%
  dplyr::select(project_name, is_cycling)
df_tme = table(df_tme$project_name, df_tme$is_cycling) %>%
  # prop.table(., margin = 1) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("project_name", "is_cycling", "num_tme")) %>%
  dplyr::filter(is_cycling == "yes") %>%
  dplyr::select(project_name, num_tme)

# Number of cycling cells in tumor cells
df_tum = sobj@meta.data %>%
  dplyr::filter(is_tumor_cells == "yes") %>%
  dplyr::select(project_name, is_cycling)
df_tum = table(df_tum$project_name, df_tum$is_cycling) %>%
  # prop.table(., margin = 1) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("project_name", "is_cycling", "num_tum")) %>%
  dplyr::filter(is_cycling == "yes") %>%
  dplyr::select(project_name, num_tum)

# Join and make ratio
df = dplyr::left_join(df_tme, df_tum, by = "project_name") %>%
  dplyr::mutate(ratio = num_tum / num_tme) %>%
  dplyr::left_join(., y = sample_info[, c("project_name", "genotype")],
                   by = "project_name")

ggplot2::ggplot(df, aes(x = num_tum, y = num_tme, label = project_name,
                        col = project_name)) +
  ggplot2::geom_abline(intercept = 0, slope = 1, col = "gray80", lty = 2) +
  ggplot2::geom_point(aes(fill = genotype), pch = 21, stroke = 0, size = 3) +
  ggrepel::geom_label_repel(min.segment.length = 0, segment.color = "black", label.size = 0) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$project_name,
                              guide = "none") +
  ggplot2::labs(x = "Number of cycling cells in tumor cells",
                y = "Number of cycling cells in tumor micro-environment") +
  ggplot2::theme_classic()
```

## Tumor cells

We load the tumor cells dataset :

```{r}
sobj_tum = readRDS(paste0(out_dir, "/../7_tumor_cells_mice/mice_tumor_cells_sobj.rds"))
sobj_tum
```











## Clusters

We visualize clusters :

```{r clustering, fig.width = 6, fig.height = 4}
cluster_plot = Seurat::DimPlot(sobj, reduction = name2D, label = TRUE) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1)
cluster_plot
```


and split by origin :

```{r plot_split_dimred, fig.width = 12, fig.height = 9}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "project_name",
                                        group_by = "seurat_clusters",
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name),
                                        group_color = aquarius::gg_color_hue(length(levels(sobj$seurat_clusters))))

plot_list[[length(plot_list) + 1]] = cluster_plot

patchwork::wrap_plots(plot_list, ncol = 6) &
  Seurat::NoLegend()
```


# Differential expression

What is the identity of specific clusters ?

We design a custom functions to represent cells of interest on the projection :

```{r see_clusters, class.source = "fold-hide"}
# Major cell type by cluster
cell_type_clusters = sobj@meta.data[, c("cell_type", "seurat_clusters")] %>%
  table() %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
cell_type_clusters = setNames(levels(sobj$cell_type)[cell_type_clusters],
                              nm = names(cell_type_clusters))

see_clusters = function(pop_oi = "macrophages") {
  # Clusters containing population of interest
  clusters_oi = names(which(cell_type_clusters == pop_oi))
  
  # Colors for clusters
  custom_colors = aquarius::gg_color_hue(length(levels(sobj$seurat_clusters)))
  names(custom_colors) = levels(sobj$seurat_clusters)
  custom_colors[!(names(custom_colors) %in% clusters_oi)] = "gray92"
  
  p1 = Seurat::DimPlot(sobj, reduction = name2D, cols = custom_colors,
                       group.by = "seurat_clusters", label = TRUE) +
    ggplot2::labs(title = "Clusters of interest",
                  subtitle = paste0(clusters_oi, collapse = ", ")) +
    ggplot2::theme(aspect.ratio = 1,
                   plot.title = element_text(hjust = 0.5),
                   plot.subtitle = element_text(hjust = 0.5)) +
    Seurat::NoAxes() + Seurat::NoLegend()
  
  # Color for cell type
  custom_colors = unlist(color_markers)
  custom_colors[!(names(custom_colors) %in% pop_oi)] = "gray92"
  
  p2 = Seurat::DimPlot(sobj, reduction = name2D,
                       group.by = "cell_type", cols = custom_colors) +
    ggplot2::labs(title = "Annotation of interest",
                  subtitle = pop_oi) +
    ggplot2::theme(aspect.ratio = 1,
                   plot.title = element_text(hjust = 0.5),
                   plot.subtitle = element_text(hjust = 0.5)) +
    Seurat::NoAxes() + Seurat::NoLegend()
  
  # Piechart for each cluster, by sample name
  plot_list = lapply(clusters_oi, FUN = function(one_cluster) {
    df = sobj@meta.data %>%
      dplyr::filter(.data$seurat_clusters == .env$one_cluster) %>%
      dplyr::select(sample_identifier, seurat_clusters)
    
    p = aquarius::plot_piechart(df,
                                grouping_var = "sample_identifier") +
      ggplot2::scale_fill_manual(values = sample_info$color,
                                 breaks = sample_info$sample_identifier) +
      ggplot2::labs(title = paste0("Cluster ", one_cluster),
                    subtitle = paste0(nrow(df), " cells")) +
      ggplot2::theme(plot.title = element_text(hjust = 0.5),
                     plot.subtitle = element_text(hjust = 0.5))
    
    return(p)
  })
  p3 = patchwork::wrap_plots(plot_list)
  
  # Patchwork
  p = patchwork::wrap_plots(p2, p1, p3, nrow = 1)
  
  return(p)
}
```

We make over-representation analysis for each group of genes. We load gene sets from MSigDB :

```{r gene_sets}
gene_sets = aquarius::get_gene_sets()
gene_sets = gene_sets$gene_sets

head(gene_sets)
```

We get gene name and gene ID correspondence :

```{r gene_corresp}
gene_corresp = aquarius::get_gene_id(
  gtf_path = "/mnt/Data/GENOME/mm10_IREScre_tdTomato_eYFP/genes/genes.gtf")

head(gene_corresp)
```

## Macrophages

Macrophages are those clusters :

```{r clusters_oimacro}
clusters_oi = names(which(cell_type_clusters == "macrophages"))
clusters_oi
```

We represent those clusters on the projection :

```{r see_macro, fig.width = 12, fig.height = 7}
see_clusters("macrophages")
```


### Cluster 27 (mito)

```{r cluster27_see_blue_red, fig.width = 5, fig.height = 5}
group1 = 27
group2 = clusters_oi

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for these group :

```{r cluster5_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
dim(mark)

head(mark, n = 20)
```


### Cluster 22 (cycling)

```{r cluster22_see_blue_red, fig.width = 5, fig.height = 5}
group1 = 22
group2 = clusters_oi

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for these group :

```{r cluster22_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
dim(mark)

head(mark, n = 20)
```


### fl/fl vs fl/-

We set specific idents to compare this population between fl/fl and fl/- genotype :

```{r macrophages_idents, fig.width = 5, fig.height = 5}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flfl" ~ "flfl",
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" ~ "flmn")

group1 = "flfl"
group2 = "flmn"

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for both groups :

```{r macrophages_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```


#### Up in flfl

```{r macrophages_flfl}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r macrophages_flfl_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Up in fl/-

```{r macrophages_flmn}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r macrophages_flmn_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


## mDC

mDC are those clusters :

```{r clusters_oimDC}
clusters_oi = names(which(cell_type_clusters == "mDC"))
clusters_oi
```

We represent those clusters on the projection :

```{r see_mDC, fig.width = 12, fig.height = 7}
see_clusters("mDC")
```

### fl/fl vs fl/-

We set specific idents to compare this population between fl/fl and fl/- genotype :

```{r mDC_idents, fig.width = 5, fig.height = 5}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flfl" ~ "flfl",
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" ~ "flmn")

group1 = "flfl"
group2 = "flmn"

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for both groups :

```{r mDC_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```


#### Up in flfl

```{r mDC_flfl}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r mDC_flfl_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Up in fl/-

```{r mDC_flmn}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r mDC_flmn_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


## cDC

cDC are those clusters :

```{r clusters_cDC}
clusters_oi = names(which(cell_type_clusters == "cDC"))
clusters_oi
```

We represent those clusters on the projection :

```{r see_cDC, fig.width = 12, fig.height = 7}
see_clusters("cDC")
```

### fl/fl vs fl/-

We set specific idents to compare this population between fl/fl and fl/- genotype :

```{r cDC_idents, fig.width = 5, fig.height = 5}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flfl" ~ "flfl",
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" ~ "flmn")

group1 = "flfl"
group2 = "flmn"

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for both groups :

```{r cDC_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```


#### Up in flfl

```{r cDC_flfl}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r cDC_flfl_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Up in fl/-

```{r cDC_flmn}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r cDC_flmn_ora, fig.width = 12, fig.height = 4}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


## pDC

pDC are those clusters :

```{r clusters_pDC}
clusters_oi = names(which(cell_type_clusters == "pDC"))
clusters_oi
```

We represent those clusters on the projection :

```{r see_pDC, fig.width = 12, fig.height = 7}
see_clusters("pDC")
```

### fl/fl vs fl/-

We set specific idents to compare this population between fl/fl and fl/- genotype :

```{r pDC_idents, fig.width = 5, fig.height = 5}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flfl" ~ "flfl",
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" ~ "flmn")

group1 = "flfl"
group2 = "flmn"

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for both groups :

```{r pDC_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```


#### Up in flfl

```{r pDC_flfl}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r pDC_flfl_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Up in fl/-

```{r pDC_flmn}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r pDC_flmn_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


## T cells

T cells are those clusters :

```{r clusters_T_cells}
clusters_oi = names(which(cell_type_clusters == "T cells"))
clusters_oi
```

We represent those clusters on the projection :

```{r see_T_cells, fig.width = 12, fig.height = 7}
see_clusters("T cells")
```


### Cluster 17 - fl/fl vs fl/-

We set specific idents to compare this population between fl/fl and fl/- genotype :

```{r T_cells_idents, fig.width = 5, fig.height = 5}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi[1] & sobj$genotype == "flfl" ~ "flfl",
  sobj$seurat_clusters %in% clusters_oi[1] & sobj$genotype == "flmn" ~ "flmn")

group1 = "flfl"
group2 = "flmn"

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for both groups :

```{r T_cells_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```


#### Up in flfl

```{r T_cells_flfl}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r T_cells_flfl_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Up in fl/-

```{r T_cells_flmn}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r T_cells_flmn_ora, fig.width = 12, fig.height = 5}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


### Cluster 14 - fl/fl vs fl/-

We set specific idents to compare this population between fl/fl and fl/- genotype :

```{r T_cells2_idents, fig.width = 5, fig.height = 5}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi[2] & sobj$genotype == "flfl" ~ "flfl",
  sobj$seurat_clusters %in% clusters_oi[2] & sobj$genotype == "flmn" ~ "flmn")

group1 = "flfl"
group2 = "flmn"

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for both groups :

```{r T_cells2_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```


#### Up in flfl

```{r T_cells2_flfl}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r T_cells2_flfl_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Up in fl/-

```{r T_cells2_flmn}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r T_cells2_flmn_ora, fig.width = 12, fig.height = 5}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


## NK cells

NK cells are those clusters :

```{r clusters_NK_cells}
clusters_oi = names(which(cell_type_clusters == "NK cells"))
clusters_oi
```

We represent those clusters on the projection :

```{r see_NK_cells, fig.width = 12, fig.height = 7}
see_clusters("NK cells")
```

### fl/fl vs fl/-

We set specific idents to compare this population between fl/fl and fl/- genotype :

```{r NK_cells_idents, fig.width = 5, fig.height = 5}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flfl" ~ "flfl",
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" ~ "flmn")

group1 = "flfl"
group2 = "flmn"

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for both groups :

```{r NK_cells_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```


#### Up in flfl

```{r NK_cells_flfl}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r NK_cells_flfl_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Up in fl/-

```{r NK_cells_flmn}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r NK_cells_flmn_ora, fig.width = 12, fig.height = 6}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


## Neutrophils

Neutrophils are those clusters :

```{r clusters_neutrophils}
clusters_oi = names(which(cell_type_clusters == "neutrophils"))
clusters_oi
```

We represent those clusters on the projection :

```{r see_neutrophils, fig.width = 12, fig.height = 7}
see_clusters("neutrophils")
```


### fl/fl vs fl/-

We set specific idents to compare this population between fl/fl and fl/- genotype :

```{r neutrophils_idents, fig.width = 5, fig.height = 5}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flfl" ~ "flfl",
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" ~ "flmn")

group1 = "flfl"
group2 = "flmn"

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for both groups :

```{r neutrophils_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```


#### Up in flfl

```{r neutrophils_flfl}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r neutrophils_flfl_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Up in fl/-

```{r neutrophils_flmn}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r neutrophils_flmn_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


## B cells

B cells are those clusters :

```{r clusters_B_cells}
clusters_oi = names(which(cell_type_clusters == "B cells"))
clusters_oi = clusters_oi[clusters_oi != "34"]
clusters_oi
```

We represent those clusters on the projection :

```{r see_B_cells, fig.width = 12, fig.height = 7}
see_clusters("B cells")
```

### fl/fl vs fl/-

We set specific idents to compare this population between fl/fl and fl/- genotype :

```{r B_cells_idents, fig.width = 5, fig.height = 5}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flfl" ~ "flfl",
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" ~ "flmn")

group1 = "flfl"
group2 = "flmn"

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for both groups :

```{r B_cells_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```


#### Up in flfl

```{r B_cells_flfl}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r B_cells_flfl_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Up in fl/-

```{r B_cells_flmn}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r B_cells_flmn_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

## Endothelial cells

Endothelial cells are those clusters :

```{r clusters_endothelial_cells}
clusters_oi = names(which(cell_type_clusters == "endothelial cells"))
clusters_oi
```

We represent those clusters on the projection :

```{r see_endothelial_cells, fig.width = 12, fig.height = 7}
see_clusters("endothelial cells")
```

### fl/fl vs fl/-

We set specific idents to compare this population between fl/fl and fl/- genotype :

```{r endothelial_cells_idents, fig.width = 5, fig.height = 5}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flfl" ~ "flfl",
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" ~ "flmn")

group1 = "flfl"
group2 = "flmn"

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for both groups :

```{r endothelial_cells_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```


#### Up in flfl

```{r endothelial_cells_flfl}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r endothelial_cells_flfl_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Up in fl/-

```{r endothelial_cells_flmn}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r endothelial_cells_flmn_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


## Fibroblasts

Fibroblasts are those clusters :

```{r clusters_fibroblasts}
clusters_oi = names(which(cell_type_clusters == "fibroblasts"))
clusters_oi
```

We represent those clusters on the projection :

```{r see_fibroblasts, fig.width = 12, fig.height = 7}
see_clusters("fibroblasts")
```

### Cluster 10 vs 21

We set specific idents to compare this population between fl/fl and fl/- genotype :

```{r fibroblasts_AB_idents, fig.width = 5, fig.height = 5}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% "10" ~ "A",
  sobj$seurat_clusters %in% "21" ~ "B")

group1 = "A"
group2 = "B"

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for both groups :

```{r fibroblasts_AB_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```


#### Up in 10

```{r fibroblasts_A}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r fibroblasts_A_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Up in 21

```{r fibroblasts_B}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r fibroblasts_B_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


### fl/fl vs fl/-

We set specific idents to compare this population between fl/fl and fl/- genotype :

```{r fibroblasts_idents, fig.width = 5, fig.height = 5}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flfl" ~ "flfl",
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" ~ "flmn")

group1 = "flfl"
group2 = "flmn"

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We make identify specific markers for both groups :

```{r fibroblasts_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```


#### Up in flfl

```{r fibroblasts_flfl}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r fibroblasts_flfl_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Up in fl/-

```{r fibroblasts_flmn}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

dim(genes_of_interest)
head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r fibroblasts_flmn_ora, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(
  gene_names = rownames(genes_of_interest),
  gene_corresp = gene_corresp[, c("NAME", "NAME")] %>% `colnames<-`(c("NAME", "ENTREZID")),
  gene_sets = gene_sets[, c("gs_name", "gene_symbol")] %>% `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```


