---
title: "Figures for the paper"
subtitle: "Version 14"
author: "Audrey"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <<- Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE,
                      
                      # save figures in PDF in a separate folder
                      # https://stackoverflow.com/questions/27992239/knitr-include-figures-in-report-and-output-figures-to-separate-files
                      dev = c('png', 'pdf'), # tiff or pdf alone renders bad in html
                      # dpi = 300,
                      fig.path = "figures_detail/",
                      pdf.options(encoding = "ISOLatin9.enc"))
```


In this file, we make the scRNAseq figures for the paper.

```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
library(ComplexHeatmap)

.libPaths()
```

# Settings

Analyses are stored here :

```{r locations}
analysis_dir = "."

mouse_dir = paste0(analysis_dir, "/../1_mice/")
patient_dir = paste0(analysis_dir, "/../2_patients/")
graft_dir = paste0(analysis_dir, "/../3_graft/")
```

These are all the samples analysed :

```{r figZ_1, class.source = 'fold-hide', fig.width = 10, fig.height = 15}
# Mice
sample_info = readRDS(paste0(mouse_dir, "/1_metadata/mice_sample_info.rds"))
plot_list = aquarius::fig_plot_gb(sample_info, title = "Nf1 fl/fl mice datasets")

# Patients
sample_info = readRDS(paste0(patient_dir, "/1_metadata/patients_sample_info.rds"))
plot_list = c(plot_list, aquarius::fig_plot_gb(sample_info, title = "NF1 patients datasets"))

# Graft experiment
sample_info = readRDS(paste0(graft_dir, "/1_metadata/graft_sample_info.rds"))
plot_list = c(plot_list, aquarius::fig_plot_gb(sample_info, title = "Graft experiment"))

# patchwork
patchwork::wrap_plots(plot_list) +
  patchwork::plot_layout(nrow = length(plot_list), heights = c(0.1,4, 0.1, 4, 0.1, 0.5)) &
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15))
```

These are the custom colors for cell populations :

```{r figZ_2, fig.width = 12, fig.height = 1, class.source = "fold-hide"}
color_markers = readRDS(paste0(mouse_dir, "/1_metadata/mice_color_markers.rds"))

color_markers[["mural cells"]] = "gray60"
names(color_markers)[which(names(color_markers) == "tumor cells")] = "TOM+ cells"
names(color_markers)[which(names(color_markers) == "macrophages")] = "M\U03A6/Mo"

data.frame(cell_type = names(color_markers),
           color = unlist(color_markers)) %>%
  ggplot2::ggplot(., aes(x = cell_type, y = 0, fill = cell_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers), breaks = names(color_markers)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank())
```

```{r clean_settings, echo = FALSE}
rm(sample_info, plot_list)
```


# Mice stat (UPDATE THIS TABLE)

This is a barplot to representing the frequencies of tumors developed by both backgrounds.

```{r fig1_B, fig.width = 4, fig.height = 5}
df_proportion = data.frame(genotype = rep(c("Nf1fl/fl", "Nf1fl/-"), 5),
                           tumor = vctrs::vec_rep_each(c("ANNUBP/MPNST", "HG-MPNST", "LG-MPNST", "cellular NF", "ANNUBP"), 2),
                           propo = c(13, 8, 17, 15, 13, 8, 9, 0, 48, 69))
df_proportion$tumor = factor(df_proportion$tumor, levels = c("ANNUBP/MPNST", "HG-MPNST", "LG-MPNST", "cellular NF", "ANNUBP"))
df_proportion$genotype = factor(df_proportion$genotype, levels = c("Nf1fl/fl", "Nf1fl/-"))
tumor_colors = c("#614A00", "#C00000", "#ED7D31", "#BFBFBF", "#FFC000")

ggplot2::ggplot(df_proportion, ggplot2::aes(x = genotype, y = propo, fill = tumor)) +
  ggplot2::geom_bar(stat = "identity", position = ggplot2::position_fill(),
                    color = "black", size = 0.5, width = 0.7) +
  ggplot2::theme_classic() +
  ggplot2::scale_fill_manual(values = tumor_colors, name = NULL) +
  ggplot2::scale_y_continuous(breaks = seq(0, 1, by = 0.1),
                              labels = paste0(seq(0, 100, by = 10), sep = " %"),
                              expand = c(0,0)) +
  ggplot2::theme(axis.title = element_blank(),
                 axis.ticks.x = element_blank(),
                 axis.line.x = element_line(colour = "lightgray"))
```

```{r clean_first_bp, echo = FALSE}
rm(df_proportion, tumor_colors)
```


# SH figure

We open the table :

```{r load_sh_tab}
sh_tab = openxlsx::read.xlsx(paste0(analysis_dir, "/Validation target SH.xlsx"), sheet = 3)
sh_tab$id = paste(sh_tab$sh_id, sh_tab$jour, sep = "_DAY_")
sh_tab = sh_tab %>%
  dplyr::filter(!(id %in% c("Sox9 clone 22-5_DAY_14", "Sox9 clone 22-6_DAY_14")))

head(sh_tab)
```

For each condition (`id`), we make a paired t-test.

```{r sh_ttest}
my_ttest = lapply(sort(unique(sh_tab$id)), FUN = function(one_id) {
  sub_tab = sh_tab %>%
    dplyr::filter(id == one_id)
  the_test = stats::t.test(x = sub_tab$mass_ctrl,
                           y = sub_tab$mass_sh,
                           paired = TRUE)
  p_value = the_test$p.value
  return(p_value)
})
my_ttest = data.frame(id = sort(unique(sh_tab$id)),
                      pvalue = unlist(my_ttest))
my_ttest$x = c(1:nrow(my_ttest))
head(my_ttest)
```

We make a long dataframe for ggplot :

```{r long_sh_tab}
long_sh_tab = rbind(sh_tab[, c("mouse_id", "sh_id", "sh_gene", "mass_ctrl", "jour", "mono_poly", "id")] %>%
                      `colnames<-`(c("mouse_id", "sh_id", "sh_gene", "mass", "jour", "mono_poly", "id")),
                    sh_tab[, c("mouse_id", "sh_id", "sh_gene", "mass_sh", "jour", "mono_poly", "id")] %>%
                      `colnames<-`(c("mouse_id", "sh_id", "sh_gene", "mass", "jour", "mono_poly", "id")))
long_sh_tab$sh_ctrl = c(rep("ctrl", nrow(long_sh_tab)/2),
                        rep("sh", nrow(long_sh_tab)/2))

head(long_sh_tab)
```

We make the figure :

```{r sh_figure, fig.width = 12, fig.height = 5, class.source = "fold-hide"}
plot_content = setNames(nm = names(table(long_sh_tab$sh_gene)),
                        as.vector(table(long_sh_tab$sh_gene)))

plot_list = lapply(names(plot_content), FUN = function(one_gene) {
  # Data for gene of interest
  sub_long_sh_tab = long_sh_tab %>%
    dplyr::filter(.data$sh_gene == .env$one_gene)
  sub_long_sh_tab$id_simple = c(1:nrow(sub_long_sh_tab))
  sub_my_ttest = my_ttest %>%
    dplyr::filter(.data$id %in% sub_long_sh_tab$id)
  sub_my_ttest$x = c(1:nrow(sub_my_ttest))
  
  # Plot
  p = ggplot2::ggplot(sub_long_sh_tab, aes(x = id, y = mass, color = sh_ctrl)) +
    # Box plot and point inside
    ggplot2::geom_boxplot(alpha = 0.1, width = 0.75) +
    ggplot2::geom_point(position = position_dodge(width = 0.75)) +
    # Horizontal bar for paired t-test
    ggplot2::geom_segment(data = sub_my_ttest, inherit.aes = FALSE,
                          aes(x = x - 0.4, xend = x + 0.4,
                              y = 1.15*max(sub_long_sh_tab$mass),
                              yend = 1.15*max(sub_long_sh_tab$mass))) +
    # p-value
    ggplot2::geom_label(data = sub_my_ttest, inherit.aes = FALSE,
                        aes(x = x,
                            y = 1.10*max(sub_long_sh_tab$mass),
                            label = format(round(pvalue, 2), nsmall = 2)),
                        fill = NA, label.size = 0) +
    # Significance
    ggplot2::geom_label(data = sub_my_ttest, inherit.aes = FALSE,
                        aes(x = x,
                            y = case_when(pvalue > 0.05 ~ 1.21*max(sub_long_sh_tab$mass),
                                          pvalue < 0.05 ~ 1.17*max(sub_long_sh_tab$mass)),
                            size = case_when(pvalue > 0.05 ~ 1,
                                             pvalue < 0.05 ~ 2),
                            label = case_when(pvalue > 0.05 ~ "n.s.",
                                              pvalue > 0.01 & pvalue < 0.05 ~ "*",
                                              pvalue > 0.001 & pvalue < 0.01 ~ "**",
                                              pvalue < 0.001 ~ "***")),
                        fill = NA, label.size = 0, size = 5) +
    ggplot2::scale_radius(range = c(4, 7), guide = "none") +
    # Theme
    ggplot2::scale_x_discrete(labels = sub_long_sh_tab$id_simple) +
    ggplot2::scale_y_continuous(breaks = seq(0, 1.1*max(sub_long_sh_tab$mass), by = 0.25),
                                labels = format(seq(0, 1.1*max(sub_long_sh_tab$mass), by = 0.25), nsmall = 2),
                                limits = c(0, 1.23*max(sub_long_sh_tab$mass)),
                                expand = c(0,0)) +
    ggplot2::labs(title = one_gene) +
    ggplot2::guides(color = guide_legend(title = "Tumor")) +
    ggplot2::theme_classic() +
    ggplot2::theme(plot.title = element_text(hjust = 0.5),
                   text = element_text(size = 12),
                   axis.line.x = element_line(colour = "lightgray"),
                   axis.title = element_blank())
  
  # Output
  return(p)
})

plot_list[[1]] = plot_list[[1]] +
  ggplot2::labs(y = "Tumor mass (g)") +
  ggplot2::theme(axis.title.y = element_text(size = 12))

# Patchwork
patchwork::wrap_plots(plot_list, widths = plot_content, nrow = 1) +
  patchwork::plot_layout(guides = "collect")
```


```{r clean_sh, echo = FALSE}
rm(sh_tab, long_sh_tab, my_ttest, plot_content, plot_list)
```


# fl/fl mice

In this section, we make the figures related to Nf1 fl/fl mice. Below are global settings.

```{r mouse_settings}
sample_info = readRDS(paste0(mouse_dir, "/1_metadata/mice_sample_info.rds")) # sample info
dotplot_markers = readRDS(paste0(mouse_dir, "/1_metadata/mice_dotplot_markers.rds")) # dotplot markers
names(dotplot_markers) = names(color_markers)
reduction_tsne = "harmony_38_tsne" # projection

# Genes with IHC validation
genes_tom_traj = c("Sox10", "S100b", "Ngfr", "L1cam", "Chl1", "Pdgfra",
                   "Pdgfrb", "Sox9", "Cd34", "Spp1", "Gas1", "Crabp1")
genes_patients = c("SOX10", "S100B", "NGFR", "L1CAM", "CHL1", "PDGFRA",
                   "PDGFRB", "SOX9", "CD34", "SPP1", "GAS1", "CRABP1")

# Genes for the heatmap
#### MAYBE ADD periostin AND 'rSC' markers
genes_heatmap = list(group1 = c("Mpz", "Mbp", "S100b", "Rxrg", "Kcna1", "Rarres2", "Ptn",
                                "C4b", "Vwa1", "Scn7a", "Adam23", "Art3", "Fxyd1", "Cadm1",
                                "Cd59a", "Chl1", "Cmtm5", "Itga1", "Cdh19", "Aatk",
                                "Lgi4", "Plp1", "Erbb3", "L1cam", "Cadm4", "Sox10", "Foxd3",
                                "Mcam", "Ngfr", "Cdkn2a", "Moxd1", "Pdgfb", "Uchl1", "Tagln",
                                "Pdgfa", "Cd151",  "Gpc1", "Crip1", "Sep15", "Rhoa"),
                     group2 = c("Ccnd1",  "Tnfrsf12a", "Snrpd3", "Tgfbi", "Actr3",
                                "Nme1", "Mif", "Pfn1", "Ybx1", "Ran", "Lgals1",
                                "Snrpf", "Nudc", "Gapdh", "Snrpd1",
                                "S100a6", "Nucks1", "Ube2s", "H2afz", "Banf1",
                                "Hmga2", "Crabp1", "Acta2", "Cd34", "Cks1b",
                                "Dut", "Tubb6", "Lyar", "Tyms", "Dctpp1", "Mcm6",
                                "Hmga1", "Fosl1", "Mcm5", 
                                "Top2a", "Birc5", "Mki67", "Gmnn",  "Anln", "Cenpa"),
                     group3 = c("Nfkbia", "Bgn", "Spp1", "Pdlim2", "Fn1", "Cpe", "Tcf4", "Cst3",
                                "Errfi1", "Grn", "Nr2f1",  "Trps1", "Nav1",
                                "Prrx1", "Pbx1", "Wls", "Pcolce",
                                "Loxl1", "Twist1",  "Igfbp4", "Serpinf1", "Ddr2", "Mgp",
                                "Dcn", "Gas1", "Pdgfra", "Sesn3", "Mmp2", "Cd302", "Cdh11", 
                                "Mpzl1", "Itgb5", "Nbl1", "Pdgfrb", "Nrep",
                                "Sox9", "Cdh13", "Spon2", "Bicc1", "Rora"))

# Cell families
custom_order_cell_type = data.frame(
  cell_type = c("pDC", "mDC", "cDC", "mast cells", "neutrophils",
                "B cells", "NK cells", "T cells", "M\U03A6/Mo", "skeletal muscle cells",
                "mural cells", "endothelial cells", "fibroblasts", "TOM+ cells"),
  cell_type_old = c("pDC", "mDC", "cDC", "mast cells", "neutrophils",
                    "B cells", "NK cells", "T cells", "macrophages", "skeletal muscle cells",
                    "mural cells", "endothelial cells", "fibroblasts", "tumor cells"),
  cell_family = c(rep("immune cells", 9),
                  rep("stromal cells", 4),
                  rep("tumor cells", 1)),
  stringsAsFactors = FALSE)
rownames(custom_order_cell_type) = custom_order_cell_type$cell_type

family_color = c("immune cells" = "gray90",
                 "stromal cells" = "gray50",
                 "tumor cells" = "gray0")
```

These are the two samples sorted for Tomato expression, by FACS. Associated datasets contain tumor cells. They will not be represented on the cell type composition figures :

```{r mouse_sample_tomato}
sample_tomato = c("2020_14", "2020_02")
sample_tomato = sample_info %>%
  dplyr::filter(project_name %in% sample_tomato) %>%
  dplyr::pull(sample_identifier) %>%
  as.character()
```

## Dataset preparation

We load the mouse dataset :

```{r load_mouse_dataset}
sobj_full = readRDS(paste0(mouse_dir, "/3_combined/all_mice_sobj.rds"))
sobj_full
```

We change the levels of `cell_type`, for the figures legend to be in a nice order.

```{r change_levels}
sobj_full$cell_type = factor(
  sobj_full$cell_type,
  levels = levels(sobj_full$cell_type)[match(custom_order_cell_type$cell_type_old,
                                             levels(sobj_full$cell_type))]) # reorder cell type
levels(sobj_full$cell_type) = custom_order_cell_type$cell_type # change level names
```


We group cell type by cluster, to correct mis-annotation :

```{r mouse_cluster_type}
cluster_type = table(sobj_full$cell_type, sobj_full$seurat_clusters) %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
cluster_type = setNames(nm = names(cluster_type),
                        levels(sobj_full$cell_type)[cluster_type])

sobj_full$cluster_type = cluster_type[sobj_full$seurat_clusters]
sobj_full$cluster_type = factor(sobj_full$cluster_type,
                                levels = levels(sobj_full$cell_type))
```

We add a `cell_family` columns in metadata :

```{r sobj_cell_family}
sobj_full$cell_family = custom_order_cell_type[sobj_full$cluster_type, "cell_family"]
```

This is the number of cells in each dataset :

```{r figS2_D2, fig.width = 5, fig.height = 6}
table(sobj_full$sample_identifier) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("sample_id", "nb_cells")) %>%
  dplyr::mutate(color = sample_info[sample_info$sample_identifier == sample_id, "color"]) %>%
  aquarius::fig_plot_gb(column_names = c("Sample ID", "Nb cells", "Color")) %>%
  patchwork::wrap_plots(ncol = 1) +
  patchwork::plot_layout(heights = c(0,5))
```


## Cell type composition

To quantify cell types in each dataset, we remove the two datasets corresponding to samples that have been sorted for Tomato expression (`r paste(sample_tomato, collapse = ", ")`).

```{r sobj_bool}
bool = !(sobj_full$sample_identifier %in% sample_tomato)
subsobj = subset(sobj_full, cells = colnames(sobj_full)[bool])
subsobj$sample_identifier = factor(subsobj$sample_identifier,
                                   levels = intersect(levels(subsobj$sample_identifier), subsobj$sample_identifier))
subsobj
```


This figure represent cell types annotation from all mouse datasets.

```{r fig3_A1, class.source = 'fold-hide', fig.width = 7, fig.height = 5}
aquarius::fig_plot_dimplot(sobj_full, reduction_tsne, "cluster_type", color_markers, title = "PNST atlas")
```

```{r fig3_A1_cell_type, class.source = 'fold-hide', fig.width = 7, fig.height = 5}
aquarius::fig_plot_dimplot(sobj_full, reduction_tsne, "cell_type", color_markers, title = "PNST atlas")
```

This figure represent cell types annotation split by genotype :

```{r fig3_A1_split_grey, class.source = 'fold-hide', fig.width = 12, fig.height = 5}
plot_list = aquarius::plot_split_dimred(sobj_full,
                                        reduction = reduction_tsne,
                                        split_by = "genotype",
                                        group_by = "cluster_type",
                                        group_color = unlist(color_markers),
                                        main_pt_size = 0.1,
                                        bg_pt_size = 0.1)

patchwork::wrap_plots(plot_list, nrow = 1) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::guides(color = guide_legend(override.aes = list(size = 3),
                                       title = "Cell type"))
```


```{r fig3_A1_split_grey_cell_type, class.source = 'fold-hide', fig.width = 12, fig.height = 5}
plot_list = aquarius::plot_split_dimred(sobj_full,
                                        reduction = reduction_tsne,
                                        split_by = "genotype",
                                        group_by = "cell_type",
                                        group_color = unlist(color_markers),
                                        main_pt_size = 0.1,
                                        bg_pt_size = 0.1)

patchwork::wrap_plots(plot_list, nrow = 1) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::guides(color = guide_legend(override.aes = list(size = 3),
                                       title = "Cell type"))
```

```{r fig3_A1_split_white, class.source = 'fold-hide', fig.width = 12, fig.height = 5}
plot_list = aquarius::plot_split_dimred(sobj_full,
                                        reduction = reduction_tsne,
                                        split_by = "genotype",
                                        group_by = "cluster_type",
                                        group_color = unlist(color_markers),
                                        main_pt_size = 0.1,
                                        bg_pt_size = 0.1,
                                        bg_color = "white")

patchwork::wrap_plots(plot_list, nrow = 1) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::guides(color = guide_legend(override.aes = list(size = 3),
                                       title = "Cell type"))
```


```{r fig3_A1_split_white_cell_type, class.source = 'fold-hide', fig.width = 12, fig.height = 5}
plot_list = aquarius::plot_split_dimred(sobj_full,
                                        reduction = reduction_tsne,
                                        split_by = "genotype",
                                        group_by = "cell_type",
                                        group_color = unlist(color_markers),
                                        main_pt_size = 0.1,
                                        bg_pt_size = 0.1,
                                        bg_color = "white")

patchwork::wrap_plots(plot_list, nrow = 1) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::guides(color = guide_legend(override.aes = list(size = 3),
                                       title = "Cell type"))
```


This figure represent Tomato expression to justify tumor cells annotation.

```{r fig3_A2, class.source = 'fold-hide', fig.width = 6, fig.height = 5}
aquarius::fig_plot_feature(sobj_full, reduction_tsne, gene = "dtTomato", gene_expression_colors = aquarius::color_gene)
```

This figure represent Prss56 expression :

```{r figS2_F, class.source = 'fold-hide', fig.width = 6, fig.height = 5}
aquarius::fig_plot_feature(sobj_full, reduction_tsne, gene = "Prss56", gene_expression_colors = aquarius::color_gene)
```

This figure represent cells, split tumor type and by genotype :

```{r fig3_B, class.source = 'fold-hide', fig.width = 12, fig.height = 8}
sobj_full$tumor_type_genotype = paste(sobj_full$tumor_type, sobj_full$genotype, sep = "_") %>%
  factor(., levels = c("pNF_flfl", "dyNF_flfl", "MPNST_flfl", "dyNF_flmn", "MPNST_flmn"))

plot_list = aquarius::plot_split_dimred(sobj_full,
                                        reduction = reduction_tsne,
                                        split_by = "tumor_type_genotype",
                                        group_by = "cluster_type",
                                        group_color = unlist(color_markers))

patchwork::wrap_plots(plot_list, nrow = 2) +
  patchwork::plot_layout(design = "ABC\n#DE") &
  Seurat::NoLegend()
```

```{r fig3_B_cell_type, class.source = 'fold-hide', fig.width = 12, fig.height = 8}
sobj_full$tumor_type_genotype = paste(sobj_full$tumor_type, sobj_full$genotype, sep = "_") %>%
  factor(., levels = c("pNF_flfl", "dyNF_flfl", "MPNST_flfl", "dyNF_flmn", "MPNST_flmn"))

plot_list = aquarius::plot_split_dimred(sobj_full,
                                        reduction = reduction_tsne,
                                        split_by = "tumor_type_genotype",
                                        group_by = "cell_type",
                                        group_color = unlist(color_markers))

patchwork::wrap_plots(plot_list, nrow = 2) +
  patchwork::plot_layout(design = "ABC\n#DE") &
  Seurat::NoLegend()
```

We represent also Tomato expression split by genotype and tumor type. Cells annotated as tumor cells in "MPNST" zone, in pNF samples, are Tom-, thus not tumor cells but fibroblasts.

```{r fig3_B2, class.source = 'fold-hide', fig.width = 12, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj_full,
                                        reduction = reduction_tsne,
                                        split_by = "tumor_type_genotype",
                                        color_by = "dtTomato")

patchwork::wrap_plots(plot_list, nrow = 2) +
  patchwork::plot_layout(design = "ABC\n#DE", guides = "collect")
```


This figure shows that batch-effect is removed on this projection. It also show cell cycle phase, and a quality metric : the percentage of UMI by cell, for mitochondrial genes 

```{r figS2_B, class.source = 'fold-hide', fig.width = 13, fig.height = 7}
aquarius::fig_plot_quality(sobj_full, reduction_tsne,
                           gene_expression_colors = aquarius::color_gene,
                           sample_info = sample_info) %>%
  patchwork::wrap_plots(., nrow = 1)
```


This barplot represents the composition of each dataset in cell populations :

```{r preparation_barplot_mouse, class.source = 'fold-hide'}
df_proportion = table(subsobj$sample_identifier,
                      subsobj$cluster_type) %>%
  prop.table(., margin = 1) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("sample_identifier", "cluster_type", "prop_cells")) %>%
  dplyr::left_join(., y = sample_info, by = "sample_identifier")

plot_content = sample_info %>%
  dplyr::filter(!(sample_identifier %in% sample_tomato)) %>%
  dplyr::pull(sample_type) %>% table()
plot_content = setNames(nm = names(plot_content), as.vector(plot_content))

plot_list = lapply(names(plot_content), FUN = function(st) {
  p = df_proportion %>%
    dplyr::filter(.data$sample_type == .env$st) %>%
    aquarius::plot_barplot(.,
                           x = "sample_identifier",
                           y = "prop_cells",
                           fill = "cluster_type",
                           position = ggplot2::position_fill()) +
    ggplot2::scale_fill_manual(values = color_markers,
                               breaks = names(color_markers),
                               name = "Cell type") +
    ggplot2::scale_y_continuous(breaks = seq(0, 1, by = 0.1),
                                labels = paste0(seq(0, 100, by = 10), sep = " %"),
                                expand = c(0,0)) +
    ggplot2::theme(axis.title.y = ggplot2::element_blank(),
                   axis.line.x = ggplot2::element_line(colour = "lightgray"),
                   plot.title = ggtext::element_textbox(hjust = 0.5,
                                                        margin = margin(t = 5, b = 5)))
  
  return(p)
})

# Reorder
names(plot_list) = names(plot_content)
plot_list = plot_list[c("flfl_pNF", "flfl_dyNF", "flmn_dyNF", "flfl_MPNST", "flmn_MPNST")]
plot_content = plot_content[names(plot_list)]

plot_list_all = plot_list
```

Barplot for all samples :

```{r figS2_D1_all, fig.width = 10, fig.height = 6}
# Remove legend except for the first one
plot_list[c(2:length(plot_list))] = lapply(plot_list[c(2:length(plot_list))], FUN = function(p) {
  p = p +
    ggplot2::theme(axis.line.y = element_blank(),
                   axis.text.y = element_blank(),
                   axis.ticks.y = element_blank())
  return(p)
})

# Patchwork
patchwork::wrap_plots(plot_list, widths = plot_content, nrow = 1) +
  patchwork::plot_layout(guides = "collect")
```

Barplot for fl/fl samples :

```{r figS2_D1_flfl, fig.width = 9, fig.height = 6}
plot_list = plot_list[c("flfl_pNF", "flfl_dyNF", "flfl_MPNST")]

# Patchwork
patchwork::wrap_plots(plot_list, widths = plot_content[c("flfl_pNF", "flfl_dyNF", "flfl_MPNST")], nrow = 1) +
  patchwork::plot_layout(guides = "collect")
```

Barplot for fl/- samples :

```{r figS2_D1_flmn, fig.width = 5, fig.height = 6}
plot_list = plot_list_all[c("flmn_dyNF", "flmn_MPNST")]

# Remove legend except for the first one
plot_list[c(2:length(plot_list))] = lapply(plot_list[c(2:length(plot_list))], FUN = function(p) {
  p = p +
    ggplot2::theme(axis.line.y = element_blank(),
                   axis.text.y = element_blank(),
                   axis.ticks.y = element_blank())
  return(p)
})

# Patchwork
patchwork::wrap_plots(plot_list, widths = plot_content, nrow = 1) +
  patchwork::plot_layout(guides = "collect")

# Patchwork
patchwork::wrap_plots(plot_list, widths = c(1.1, 3), nrow = 1) +
  patchwork::plot_layout(guides = "collect")
```

This barplot represents the proportion of cell populations, by tumor type for all mice, except the two samples filtered for Tomato expression by FACS :

```{r fig3_C, class.source = 'fold-hide', fig.width = 4.5, fig.height = 5.5}
subsobj = subset(sobj_full, cells = colnames(sobj_full)[bool])
subsobj$sample_identifier = factor(subsobj$sample_identifier,
                                   levels = intersect(levels(subsobj$sample_identifier), subsobj$sample_identifier))

family_color[["cycling tumor cells"]] = family_color[["tumor cells"]]
family_color = family_color[c("immune cells", "stromal cells", "tumor cells", "cycling tumor cells")]

subsobj$cell_family = ifelse(subsobj$cluster_type == "TOM+ cells" & subsobj$Seurat.Phase %in% c("G2M", "S"),
                             yes = "cycling tumor cells",
                             no = subsobj$cell_family)

aquarius::fig_plot_bp_tt(subsobj, groupby = "cluster_type",
                         colors = color_markers, family_color = family_color) +
  Seurat::NoLegend()
```

This is the same figure, but for **fl/fl mice** only :

```{r fig3_C_flfl, class.source = 'fold-hide', fig.width = 4.5, fig.height = 5.5}
subsubsobj = subset(subsobj, genotype == "flfl")
subsubsobj$sample_identifier = factor(subsubsobj$sample_identifier,
                                      levels = intersect(levels(subsubsobj$sample_identifier), subsubsobj$sample_identifier))

family_color[["cycling tumor cells"]] = family_color[["tumor cells"]]
family_color = family_color[c("immune cells", "stromal cells", "tumor cells", "cycling tumor cells")]

subsubsobj$cell_family = ifelse(subsubsobj$cluster_type == "TOM+ cells" & subsubsobj$Seurat.Phase %in% c("G2M", "S"),
                                yes = "cycling tumor cells",
                                no = subsubsobj$cell_family)

aquarius::fig_plot_bp_tt(subsubsobj, groupby = "cluster_type",
                         colors = color_markers, family_color = family_color) +
  Seurat::NoLegend()
```

This is the same figure, but for **fl/- mice** only :

```{r fig3_C_flmn, class.source = 'fold-hide', fig.width = 4.5, fig.height = 5.5}
subsubsobj = subset(subsobj, genotype == "flmn")
subsubsobj$sample_identifier = factor(subsubsobj$sample_identifier,
                                      levels = intersect(levels(subsubsobj$sample_identifier), subsubsobj$sample_identifier))

family_color[["cycling tumor cells"]] = family_color[["tumor cells"]]
family_color = family_color[c("immune cells", "stromal cells", "tumor cells", "cycling tumor cells")]

subsubsobj$cell_family = ifelse(subsubsobj$cluster_type == "TOM+ cells" & subsubsobj$Seurat.Phase %in% c("G2M", "S"),
                                yes = "cycling tumor cells",
                                no = subsubsobj$cell_family)

aquarius::fig_plot_bp_tt(subsubsobj, groupby = "cluster_type",
                         colors = color_markers, family_color = family_color) +
  Seurat::NoLegend()
```

This dotplot assesses the cell type annotation :

```{r figS2_A, class.source = 'fold-hide', fig.width = 8, fig.height = 8}
markers = dotplot_markers[levels(sobj_full$cluster_type)] # new order
markers = unique(unlist(markers)) # as vector
markers = markers[!(markers %in% c("Ptprc", "dtTomato"))]
markers = c("Ptprc", markers, "dtTomato")

aquarius::fig_plot_dotplot(sobj_full,
                           groupby = "cluster_type",
                           markers = markers,
                           gene_expression_colors = aquarius::color_gene) +
  ggplot2::geom_vline(xintercept = 0.5 + c(5, 7, 9, 11, 13, 15,
                                           18, 20, 22, 24, 26, 28, 30, 32),
                      color = "gray92")
```

```{r clean_mouse, echo = FALSE}
rm(dotplot_markers, bool, markers, sample_tomato,
   df_proportion, cluster_type, subsubsobj)
```


How many cells for each cell type, by tumor type ?

```{r figS2_E, fig.height = 5, fig.width = 8, class.source = "fold-hide"}
df_quantities = table(subsobj$cluster_type,
                      subsobj$tumor_type) %>%
  prop.table(., margin = 2) %>%
  "*"(100) %>% round(., 2) %>%
  as.data.frame.array(stringAsFactors = FALSE) %>%
  dplyr::mutate(cluster_type = rownames(.)) %>%
  dplyr::relocate(cluster_type)
df_quantities$nb_cells = rowSums(table(subsobj$cluster_type, subsobj$tumor_type))
df_quantities = rbind(df_quantities,
                      c("Number of cells", as.numeric(table(subsobj$tumor_type)), ncol(subsobj)))

columns_color = unlist(color_markers[levels(subsobj$cluster_type)])
columns_color = c(columns_color, NA, rep(NA, times = nrow(df_quantities)*(ncol(df_quantities) - 1)))
font_color = ifelse(is.na(columns_color), yes = "black", no = "white")

# Cell type
custom_theme1 = gridExtra::ttheme_minimal(base_colour = font_color,
                                          core = list(bg_params = list(fill = columns_color,
                                                                       col = "black")),
                                          colhead = list(fg_params = list(col = "black")))
gb1 = gridExtra::tableGrob(df_quantities,
                           rows = NULL,
                           cols = c("Cell type", "pNF", "dyNF", "MPNST", "Nb cells"),
                           theme = custom_theme1)

# Cell family (merged cells)
custom_theme2 = gridExtra::ttheme_minimal(base_colour = c("black", "white", "white"),
                                          core = list(bg_params = list(fill = family_color,
                                                                       col = "black")),
                                          colhead = list(fg_params = list(col = "black")))
gb2 = data.frame(cell_family = c("immune cells", "stromal cells", "tumor cells")) %>%
  gridExtra::tableGrob(.,
                       row = NULL,
                       cols = c("Cell family"),
                       theme = custom_theme2)

# Merge
gb = gridExtra::gtable_combine(gb2,gb1, along = 1)
gb$layout[gb$layout$t != 1 & gb$layout$l == 1, c("t")] = c(2, 11, 15, 2, 11, 15)
gb$layout[gb$layout$b != 1  & gb$layout$l == 1, c("b")] = c(10, 14, 15, 10, 14, 15)

# Patchwork
title_gg = ggplot2::ggplot() +
  ggplot2::theme_minimal()

patchwork::wrap_plots(list(title_gg, gb)) +
  patchwork::plot_layout(design = "A\nB", heights = c(0.1,5)) &
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15))
```

We make this table only for fl/fl samples :

```{r figS2_E_flfl, fig.height = 5, fig.width = 8, class.source = "fold-hide"}
submetadata = subsobj@meta.data %>%
  dplyr::filter(genotype == "flfl")

df_quantities = table(submetadata$cluster_type,
                      submetadata$tumor_type) %>%
  prop.table(., margin = 2) %>%
  "*"(100) %>% round(., 2) %>%
  as.data.frame.array(stringAsFactors = FALSE) %>%
  dplyr::mutate(cluster_type = rownames(.)) %>%
  dplyr::relocate(cluster_type)
df_quantities$nb_cells = rowSums(table(submetadata$cluster_type, submetadata$tumor_type))
df_quantities = rbind(df_quantities,
                      c("Number of cells", as.numeric(table(submetadata$tumor_type)), nrow(submetadata)))

columns_color = unlist(color_markers[levels(submetadata$cluster_type)])
columns_color = c(columns_color, NA, rep(NA, times = nrow(df_quantities)*(ncol(df_quantities) - 1)))
font_color = ifelse(is.na(columns_color), yes = "black", no = "white")

# Cell type
custom_theme1 = gridExtra::ttheme_minimal(base_colour = font_color,
                                          core = list(bg_params = list(fill = columns_color,
                                                                       col = "black")),
                                          colhead = list(fg_params = list(col = "black")))
gb1 = gridExtra::tableGrob(df_quantities,
                           rows = NULL,
                           cols = c("Cell type", "pNF", "dyNF", "MPNST", "Nb cells"),
                           theme = custom_theme1)

# Cell family (merged cells)
custom_theme2 = gridExtra::ttheme_minimal(base_colour = c("black", "white", "white"),
                                          core = list(bg_params = list(fill = family_color,
                                                                       col = "black")),
                                          colhead = list(fg_params = list(col = "black")))
gb2 = data.frame(cell_family = c("immune cells", "stromal cells", "tumor cells")) %>%
  gridExtra::tableGrob(.,
                       row = NULL,
                       cols = c("Cell family"),
                       theme = custom_theme2)

# Merge
gb = gridExtra::gtable_combine(gb2,gb1, along = 1)
gb$layout[gb$layout$t != 1 & gb$layout$l == 1, c("t")] = c(2, 11, 15, 2, 11, 15)
gb$layout[gb$layout$b != 1  & gb$layout$l == 1, c("b")] = c(10, 14, 15, 10, 14, 15)

# Patchwork
title_gg = ggplot2::ggplot() +
  ggplot2::theme_minimal()

patchwork::wrap_plots(list(title_gg, gb)) +
  patchwork::plot_layout(design = "A\nB", heights = c(0.1,5)) &
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15))
```

and only for fl/- :

```{r figS2_E_flmn, fig.height = 5, fig.width = 8, class.source = "fold-hide"}
submetadata = subsobj@meta.data %>%
  dplyr::filter(genotype == "flmn")

df_quantities = table(submetadata$cluster_type,
                      submetadata$tumor_type) %>%
  prop.table(., margin = 2) %>%
  "*"(100) %>% round(., 2) %>%
  as.data.frame.array(stringAsFactors = FALSE) %>%
  dplyr::mutate(cluster_type = rownames(.)) %>%
  dplyr::relocate(cluster_type)
df_quantities$nb_cells = rowSums(table(submetadata$cluster_type, submetadata$tumor_type))
df_quantities = rbind(df_quantities,
                      c("Number of cells", as.numeric(table(submetadata$tumor_type)), nrow(submetadata)))

# Remove pNF column
df_quantities = df_quantities %>%
  dplyr::select(-pNF)

columns_color = unlist(color_markers[levels(submetadata$cluster_type)])
columns_color = c(columns_color, NA, rep(NA, times = nrow(df_quantities)*(ncol(df_quantities) - 1)))
font_color = ifelse(is.na(columns_color), yes = "black", no = "white")

# Cell type
custom_theme1 = gridExtra::ttheme_minimal(base_colour = font_color,
                                          core = list(bg_params = list(fill = columns_color,
                                                                       col = "black")),
                                          colhead = list(fg_params = list(col = "black")))
gb1 = gridExtra::tableGrob(df_quantities,
                           rows = NULL,
                           cols = c("Cell type", "dyNF", "MPNST", "Nb cells"),
                           theme = custom_theme1)

# Cell family (merged cells)
custom_theme2 = gridExtra::ttheme_minimal(base_colour = c("black", "white", "white"),
                                          core = list(bg_params = list(fill = family_color,
                                                                       col = "black")),
                                          colhead = list(fg_params = list(col = "black")))
gb2 = data.frame(cell_family = c("immune cells", "stromal cells", "tumor cells")) %>%
  gridExtra::tableGrob(.,
                       row = NULL,
                       cols = c("Cell family"),
                       theme = custom_theme2)

# Merge
gb = gridExtra::gtable_combine(gb2,gb1, along = 1)
gb$layout[gb$layout$t != 1 & gb$layout$l == 1, c("t")] = c(2, 11, 15, 2, 11, 15)
gb$layout[gb$layout$b != 1  & gb$layout$l == 1, c("b")] = c(10, 14, 15, 10, 14, 15)

# Patchwork
title_gg = ggplot2::ggplot() +
  ggplot2::theme_minimal()

patchwork::wrap_plots(list(title_gg, gb)) +
  patchwork::plot_layout(design = "A\nB", heights = c(0.1,5)) &
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15))
```



```{r clean_prop, echo = FALSE}
rm(df_quantities, custom_theme1, custom_theme2, gb, gb1, gb2,
   title_gg, font_color, columns_color, family_color)
```


## Zoom on tumor cells

We load the trajectory object :

```{r load_trajectory}
sobj = readRDS(paste0(mouse_dir, "/4_tumor_cells/mice_tumor_cells_subset_sobj_traj_slingshot_FE.rds"))
name2D = "mnn_dm_umap"

sobj
```


These are colors for the three groups for the functional enrichment analysis :

```{r group_colors}
group_colors = c("#77DEBF", "#DFA67A", "#A29EDA")
```

### UMAP

Where are those cells on the full TSNE ?

```{r fig4_A1, fig.width = 6, fig.height = 6}
sobj_full$is_tumor_cells = colnames(sobj_full) %in% colnames(sobj)

Seurat::DimPlot(sobj_full, group.by = "is_tumor_cells", reduction = reduction_tsne,
                cols = c("lightgray", color_markers[["TOM+ cells"]])) +
  ggplot2::labs(x = "tSNE 1", y = "tSNE 2", title = "Tumor cells",
                subtitle = paste0(length(colnames(sobj)), " cells")) +
  Seurat::NoLegend() + Seurat::NoAxes() +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
                 plot.subtitle = element_text(hjust = 0.5, face = "bold", size = 12),
                 aspect.ratio = 1)
```

This figure represents the sample of origin, to assess batch-effect correction of the projection. It shows cell cycle phase of each cell, and pseudotime values.

```{r fig4_A2, fig.width = 20, fig.height = 5, class.source = 'fold-hide'}
plot_list = list()

## Color by sample
plot_list[[2]] = Seurat::DimPlot(sobj,
                                 group.by = "sample_identifier",
                                 reduction = name2D,
                                 pt.size = 0.25) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$sample_identifier) +
  ggplot2::labs(title = paste(ncol(sobj), "tumor cells from", nrow(sample_info), "samples", sep = " ")) +
  ggplot2::guides(col = guide_legend(ncol = 1, title = "Sample name",
                                     override.aes = list(size = 2))) +
  ggplot2::theme(legend.position = "none")

## Color by cell cycle
plot_list[[3]] = Seurat::DimPlot(sobj,
                                 group.by = "Seurat.Phase",
                                 reduction = name2D,
                                 pt.size = 0.25) +
  ggplot2::scale_color_manual(values = aquarius::gg_color_hue(3),
                              labels = c("G0 / G1", "G2M", "S")) +
  ggplot2::labs(title = "Cell cycle phase") +
  ggplot2::theme(legend.position = c(0.05, 0.1),
                 legend.direction = "vertical") +
  ggplot2::guides(col = guide_legend(override.aes = list(size = 4)))

## Pseudotime
plot_list[[1]] = Seurat::FeaturePlot(sobj, pt.size = 0.25,
                                     reduction = name2D,
                                     features = "pseudotime") +
  ggplot2::scale_color_gradientn(colors = viridis::viridis(n = 100),
                                 breaks = range(sobj$pseudotime),
                                 labels = c("min", "max")) +
  ggplot2::labs(title = "Pseudotime") +
  ggplot2::theme(legend.position = c(0.01, 0.1),
                 legend.direction = "horizontal")

## Groups of cells
plot_list[[4]] = Seurat::DimPlot(sobj,
                                 reduction = name2D,
                                 group.by = "kmeans_group",
                                 cols = c(group_colors, "black")) +
  ggplot2::labs(title = "Cells groups") +
  ggplot2::theme(legend.position = c(0.05, 0.1)) +
  ggplot2::guides(color = guide_legend(override.aes = list(size = 4)))

## Genotype
plot_list[[5]] = Seurat::DimPlot(sobj, pt.size = 0.001,
                                 reduction = name2D,
                                 group.by = "genotype",
                                 cols = c("#C82721", "#2A71A6"),
                                 order = sample(colnames(sobj), size = ncol(sobj), replace = FALSE)) +
  ggplot2::labs(title = "Genotype") +
  ggplot2::theme(legend.position = c(0.05, 0.1)) +
  ggplot2::guides(color = guide_legend(override.aes = list(size = 4)))

## Same theme
plot_list = lapply(plot_list, FUN = function(one_plot) {
  one_plot +
    ggplot2::theme(aspect.ratio = 1,
                   plot.title = element_text(hjust = 0.5)) +
    Seurat::NoAxes()
})

## Patchwork
patchwork::wrap_plots(plot_list, nrow = 1) &
  ggplot2::theme(legend.text = element_text(size = 12))
```

This is a barplot depicted glial vs mesenchymal contribution to tumor cells population, in each sample. Warning : "high mito" cells are not there, and they are all glial cells present in dyNF samples from fl/fl samples, the proportion of mesenchymal cells is over-estimated. Moreover, pay attention to the number of cells in each sample. For some samples, 100 \% of cells correspond to 500 cells, and in some other, thousands of tumor cells were available.

```{r fig4_Z, class.source = "fold-hide", fig.width = 12, fig.height = 3.5}
sobj$Sox10_on_Sox9 = Seurat::FetchData(sobj, "Sox10")[, 1] / (Seurat::FetchData(sobj, "Sox9")[, 1] + Seurat::FetchData(sobj, "Sox10")[, 1] + 0.001)
Sox_group = sobj@meta.data %>%
  dplyr::select(seurat_clusters, Sox10_on_Sox9) %>%
  dplyr::group_by(seurat_clusters) %>%
  dplyr::summarise(Sox_group = mean(Sox10_on_Sox9)) %>%
  dplyr::mutate(Sox_group = ifelse(Sox_group > 0.6, yes = "glial", no = "mesenchymal")) %>%
  as.data.frame()
Sox_group = setNames(nm = Sox_group$seurat_clusters,
                     Sox_group$Sox_group)
# sobj$Sox_group = Sox_group[sobj$seurat_clusters]
sobj$Sox_group = ifelse(sobj$pseudotime < 0.65,
                        yes = "glial", no = "mesenchymal")

#### Plot
Sox_group = table(sobj$sample_identifier,
                  sobj$Sox_group) %>%
  prop.table(., margin = 1) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("sample_identifier", "Sox_group", "prop_cells"))

plot_list = list()

plot_list[[1]] = aquarius::plot_barplot(Sox_group,
                       x = "sample_identifier",
                       y = "prop_cells",
                       fill = "Sox_group",
                       position = ggplot2::position_fill()) +
  ggplot2::scale_fill_manual(values = c("#6aa84f", "#a61c00"),
                             breaks = c("glial", "mesenchymal"),
                             name = "Sox group") +
  ggplot2::scale_y_continuous(breaks = seq(0, 1, by = 0.1),
                              labels = paste0(seq(0, 100, by = 10), sep = " %"),
                              expand = c(0,0)) +
  ggplot2::theme(axis.title.y = ggplot2::element_blank(),
                 axis.line.x = ggplot2::element_line(colour = "lightgray"),
                 plot.title = ggtext::element_textbox(hjust = 0.5,
                                                      margin = margin(t = 5, b = 5))) +
  Seurat::NoLegend()

plot_list[[2]] = Seurat::DimPlot(sobj, reduction = name2D, group.by = "Sox_group") +
  ggplot2::scale_color_manual(values = c("#6aa84f", "#a61c00"),
                             breaks = c("glial", "mesenchymal"),
                             name = "Sox group") +
  ggplot2::theme(aspect.ratio = 1) +
  Seurat::NoAxes()

patchwork::wrap_plots(plot_list, nrow = 1, widths = c(1,0.4))
```


This is a representation of tumor cells for each sample of origin :

```{r fig4_A3, class.source = 'fold-hide', fig.width = 16.5, fig.height = 3.2}
plot_list = aquarius::plot_split_dimred(sobj = sobj,
                                        reduction = name2D,
                                        split_by = "sample_identifier",
                                        split_color = sample_info$color,
                                        bg_pt_size = 0.01,
                                        main_pt_size = 0.01)

# Patchwork
patchwork::wrap_plots(plot_list, nrow = 2) +
  patchwork::plot_layout(widths = 15, heights = 1)
```


This is a figure showing gene expression, for genes of interest.

```{r fig4_C, class.source = 'fold-hide', fig.width = 4, fig.height = 4}
plot_list = lapply(c(genes_tom_traj, "dtTomato"),
                   FUN = function(one_gene) {
                     one_plot = aquarius::fig_plot_feature(sobj,
                                                           myred = name2D,
                                                           gene = one_gene,
                                                           gene_expression_colors = aquarius::color_gene,
                                                           pt.size = 0.10) +
                       ggplot2::labs(title = one_gene) +
                       ggplot2::theme(axis.title = element_blank(),
                                      axis.line = element_blank(),
                                      legend.position = "left")
                     return(one_plot)
                   })
plot_list
```

For those tumor cells, we can also represent quality metrics :

```{r figS2_Z, fig.width = 12, fig.height = 3, class.source = 'fold-hide'}
plot_list = lapply(c("log_nCount_RNA", "nFeature_RNA", "percent.mt", "percent.rb"), FUN = function(metric) {
  Seurat::FeaturePlot(sobj, reduction = name2D,
                      features = metric) +
    ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
    ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
                   legend.text = element_text(size = 9),
                   aspect.ratio = 1) +
    Seurat::NoAxes()
})
plot_list[[1]] = plot_list[[1]] + ggplot2::ggtitle("Log number of UMI")
plot_list[[2]] = plot_list[[2]] + ggplot2::ggtitle("Number of genes with\n positive number of counts")
plot_list[[3]] = plot_list[[3]] + ggplot2::ggtitle("Proportion of transcripts\n from mitochondrial genes")
plot_list[[4]] = plot_list[[4]] + ggplot2::ggtitle("Proportion of transcripts\n from ribosomal genes")

patchwork::wrap_plots(plot_list, nrow = 1)
```

```{r clean_traj_mouse, echo = FALSE}
rm(plot_list, sobj_full)
```

### Heatmap

We prepare the genes and gene annotation.

```{r heatmap_genes, class.source = "fold-hide"}
features_group = sapply(c(1:length(genes_heatmap)),
                        FUN = function(i) {
                          rep(x = names(genes_heatmap)[i],
                              length = length(genes_heatmap[[i]]))}) %>% unlist() %>% as.vector()
features = unique(unlist(genes_heatmap))
```

We prepare the gene expression matrix :

```{r heatmap_matrix, fig.width = 12, fig.height = 30}
mat_expression = Seurat::GetAssayData(sobj, assay = "RNA", slot = "data")
mat_expression = mat_expression[features, ]
mat_expression = Matrix::t(mat_expression)
mat_expression = dynutils::scale_quantile(mat_expression) # between 0 and 1
mat_expression = Matrix::t(mat_expression)
dim(mat_expression)
```

We prepare all the colors for heatmap annotation :

```{r heatmap_colors_ann, class.source = 'fold-hide'}
list_colors = list()
list_colors[["expression"]] = aquarius::color_cnv
list_colors[["pseudotime"]] = circlize::colorRamp2(breaks = seq(min(sobj$pseudotime),
                                                                max(sobj$pseudotime),
                                                                length.out = 9),
                                                   colors = viridis::viridis(9))
list_colors[["sample_type"]] = c(flfl_pNF = "blue",
                                 flfl_dyNF = "green", flfl_MPNST = "brown",
                                 flmn_dyNF = "green", flmn_MPNST = "brown")
list_colors[["cell_cycle_phase"]] = aquarius::gg_color_hue(3)
names(list_colors[["cell_cycle_phase"]]) = c("G0 / G1", "G2M", "S")
list_colors[["percent_mito"]] = circlize::colorRamp2(breaks = seq(from = min(sobj$percent.mt),
                                                                  to = max(sobj$percent.mt),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
list_colors[["nFeature_RNA"]] = circlize::colorRamp2(breaks = seq(from = min(sobj$nFeature_RNA),
                                                                  to = max(sobj$nFeature_RNA),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
```

We prepare the heatmap annotation :

```{r heatmap_annotation, class.source = 'fold-hide'}
name_side = "right"
ha_top = HeatmapAnnotation(
  group = anno_block(gp = gpar(fill = group_colors),
                     labels = sort(unique(sobj$kmeans_group))),
  sample_type = sobj$sample_type,
  pseudotime = sobj$pseudotime,
  annotation_legend_param = list(pseudotime = list(at = range(sobj$pseudotime),
                                                   labels = c("min", "max"))),
  col = list(sample_type = list_colors[["sample_type"]],
             pseudotime = list_colors[["pseudotime"]]),
  annotation_name_side = name_side)

ha_bottom = HeatmapAnnotation(
  percent_mito = sobj$percent.mt,
  nFeature_RNA = sobj$nFeature_RNA,
  cell_cycle_phase = as.factor(sobj$Seurat.Phase) %>% `levels<-`(c("G0 / G1", "G2M", "S")),
  col = list(percent_mito = list_colors[["percent_mito"]],
             nFeature_RNA = list_colors[["nFeature_RNA"]],
             cell_cycle_phase = list_colors[["cell_cycle_phase"]]),
  annotation_name_side = name_side)
```

We build the heatmap :

```{r heatmap_ht, class.source = 'fold-hide'}
ht = ComplexHeatmap::Heatmap(
  as.matrix(mat_expression),
  # Expression colors
  col = list_colors[["expression"]],
  heatmap_legend_param = list(title = "expression", at = c(0, 1), 
                              labels = c("low", "high")),
  # Annotation
  top_annotation = ha_top,
  bottom_annotation = ha_bottom,
  # Cells grouping
  column_split = sobj$kmeans_group,
  column_gap = unit(2, "mm"),
  column_title = NULL,
  # Cells order
  column_order = order(sobj$pseudotime),
  cluster_columns = FALSE,
  show_column_dend = FALSE,
  # Cells label
  show_column_names = FALSE,
  # Gene grouping
  row_split = features_group,
  row_gap = unit(2, "mm"),
  row_title = NULL,
  cluster_rows = FALSE,
  show_row_dend = FALSE,
  # Gene labels
  row_labels = features,
  show_row_names = TRUE,
  row_names_side = name_side,
  row_names_gp = grid::gpar(fontsize = 13),
  # Visual aspect
  show_heatmap_legend = TRUE,
  border = TRUE,
  use_raster = FALSE)
```

```{r fig4_B, fig.width = 16.5, fig.height = 32}
ComplexHeatmap::draw(ht, # legend_grouping = "original",
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


### Functional analysis

For each group, we build the barplot from functional enrichment analysis. We load the pre-computed results :

```{r load_func_enr_list}
func_enr_list = readRDS(paste0(mouse_dir, "/4_tumor_cells/mice_tumor_cells_subset_list_results.rds"))
names(func_enr_list)
```

Then, we set the gene sets of interest for each group :

```{r gsea_gene_sets, class.source = "fold-hide"}
group1_gs_name = c("GOBP_GLIAL_CELL_DIFFERENTIATION",
                   "GOBP_ENSHEATHMENT_OF_NEURONS",
                   "GOBP_SCHWANN_CELL_DIFFERENTIATION",
                   "REACTOME_EGR2_AND_SOX10_MEDIATED_INITIATION_OF_SCHWANN_CELL_MYELINATION",
                   "REACTOME_L1CAM_INTERACTIONS",
                   "GOBP_GLIOGENESIS",
                   "GOMF_STRUCTURAL_CONSTITUENT_OF_CYTOSKELETON",
                   "REACTOME_RECYCLING_PATHWAY_OF_L1",
                   "GOCC_MYELIN_SHEATH",
                   "GOCC_GLIAL_CELL_PROJECTION")
group2_gs_name = c("GOBP_CYTOPLASMIC_TRANSLATION",
                   "GOBP_DNA_REPLICATION",
                   "KEGG_RIBOSOME",
                   "KEGG_SPLICEOSOME",
                   "HALLMARK_MYC_TARGETS_V1",
                   "HALLMARK_E2F_TARGETS",
                   "HALLMARK_G2M_CHECKPOINT",
                   "REACTOME_M_PHASE",
                   "REACTOME_CELL_CYCLE_CHECKPOINTS",
                   "REACTOME_DNA_REPLICATION")
group3_gs_name = c("GOBP_OSSIFICATION",
                   "GOBP_AMEBOIDAL_TYPE_CELL_MIGRATION",
                   "GOBP_RESPONSE_TO_TRANSFORMING_GROWTH_FACTOR_BETA",
                   "GOBP_MESENCHYME_DEVELOPMENT",
                   "GOBP_TRANSFORMING_GROWTH_FACTOR_BETA_RECEPTOR_SIGNALING_PATHWAY",
                   "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",
                   "HALLMARK_TNFA_SIGNALING_VIA_NFKB",
                   "HALLMARK_HYPOXIA",
                   "HALLMARK_APOPTOSIS",
                   "REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION")
```

We make a dataframe containing scores for each gene sets and each cells, for the three groups :

```{r figY_1, class.source = "fold-hide", fig.width = 13, fig.height = 6}
df_score_group1 = func_enr_list[["group 1_vs_group 3"]]$gsea_up[group1_gs_name, ]
df_score_group1$group = "group 1"
df_score_group1$NES = abs(df_score_group1$NES)
upper_bound = ceiling(max(df_score_group1$NES))

ggplot2::ggplot(df_score_group1, aes(x = stats::reorder(.data$ID, .data$NES),
                                     y = -.data$NES, fill = .data$group)) +
  ggplot2::geom_bar(stat = "identity", col = "black") +
  ggplot2::coord_flip(clip = "off") +
  ggplot2::geom_hline(yintercept = 0) +
  ggplot2::geom_label(mapping = aes(label = .data$ID, y = -0.05), fill = NA, label.size = NA,
                      hjust = 1, size = 4.5, fontface = "bold") +
  ggplot2::facet_grid(group ~ ., scales = "free", space = "free") +
  ggplot2::scale_x_discrete(position = "top") +
  ggplot2::scale_y_continuous(breaks = c(-upper_bound:0), labels = c(upper_bound:0)) +
  ggplot2::labs(y = "Normalized Enrichment Score (NES)", title = "") +
  ggplot2::geom_label(aes(label = round(.data$p.adjust, 4), y = -(.data$NES + 0.1)),
                      fill = NA, label.size = 0) +
  ggplot2::scale_fill_manual(values = group_colors[1],
                             guide = "none") +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "top",
                 legend.direction = "horizontal",
                 axis.title.y = element_blank(),
                 panel.border = element_blank(),
                 axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 strip.background = element_blank(),
                 strip.text = element_blank(),
                 panel.spacing = grid::unit(2, "lines"),
                 panel.grid.minor.x = element_blank(),
                 panel.grid.major.y = element_blank(),
                 axis.text.x = element_text(size = 15),
                 axis.title.x = element_text(size = 15),
                 axis.line.x = element_line(size = 0.5, color = "black"))
```


```{r figY_2, class.source = "fold-hide", fig.width = 13, fig.height = 6}
df_score_group2 = func_enr_list[["group 1_vs_group 2"]]$gsea_down[group2_gs_name, ]
df_score_group2$group = "group 2"
df_score_group2$NES = abs(df_score_group2$NES)
upper_bound = ceiling(max(df_score_group2$NES))

ggplot2::ggplot(df_score_group2, aes(x = stats::reorder(.data$ID, .data$NES),
                                     y = -.data$NES, fill = .data$group)) +
  ggplot2::geom_bar(stat = "identity", col = "black") +
  ggplot2::coord_flip(clip = "off") +
  ggplot2::geom_hline(yintercept = 0) +
  ggplot2::geom_label(mapping = aes(label = .data$ID, y = -0.05), fill = NA, label.size = NA,
                      hjust = 1, size = 4.5, fontface = "bold") +
  ggplot2::facet_grid(group ~ ., scales = "free", space = "free") +
  ggplot2::scale_x_discrete(position = "top") +
  ggplot2::scale_y_continuous(breaks = c(-upper_bound:0), labels = c(upper_bound:0)) +
  ggplot2::labs(y = "Normalized Enrichment Score (NES)", title = "") +
  ggplot2::geom_label(aes(label = round(.data$p.adjust, 4), y = -(.data$NES + 0.1)),
                      fill = NA, label.size = 0) +
  ggplot2::scale_fill_manual(values = group_colors[2],
                             guide = "none") +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "top",
                 legend.direction = "horizontal",
                 axis.title.y = element_blank(),
                 panel.border = element_blank(),
                 axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 strip.background = element_blank(),
                 strip.text = element_blank(),
                 panel.spacing = grid::unit(2, "lines"),
                 panel.grid.minor.x = element_blank(),
                 panel.grid.major.y = element_blank(),
                 axis.text.x = element_text(size = 15),
                 axis.title.x = element_text(size = 15),
                 axis.line.x = element_line(size = 0.5, color = "black"))
```


```{r figY_3, class.source = "fold-hide", fig.width = 13, fig.height = 6}
df_score_group3 = func_enr_list[["group 1_vs_group 3"]]$gsea_down[group3_gs_name, ]
df_score_group3$group = "group 3"
df_score_group3$NES = abs(df_score_group3$NES)
upper_bound = ceiling(max(df_score_group3$NES))

ggplot2::ggplot(df_score_group3, aes(x = stats::reorder(.data$ID, .data$NES),
                                     y = -.data$NES, fill = .data$group)) +
  ggplot2::geom_bar(stat = "identity", col = "black") +
  ggplot2::coord_flip(clip = "off") +
  ggplot2::geom_hline(yintercept = 0) +
  ggplot2::geom_label(mapping = aes(label = .data$ID, y = -0.05), fill = NA, label.size = NA,
                      hjust = 1, size = 4.5, fontface = "bold") +
  ggplot2::facet_grid(group ~ ., scales = "free", space = "free") +
  ggplot2::scale_x_discrete(position = "top") +
  ggplot2::scale_y_continuous(breaks = c(-upper_bound:0), labels = c(upper_bound:0)) +
  ggplot2::labs(y = "Normalized Enrichment Score (NES)", title = "") +
  ggplot2::geom_label(aes(label = round(.data$p.adjust, 4), y = -(.data$NES + 0.1)),
                      fill = NA, label.size = 0) +
  ggplot2::scale_fill_manual(values = group_colors[3],
                             guide = "none") +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "top",
                 legend.direction = "horizontal",
                 axis.title.y = element_blank(),
                 panel.border = element_blank(),
                 axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 strip.background = element_blank(),
                 strip.text = element_blank(),
                 panel.spacing = grid::unit(2, "lines"),
                 panel.grid.minor.x = element_blank(),
                 panel.grid.major.y = element_blank(),
                 axis.text.x = element_text(size = 15),
                 axis.title.x = element_text(size = 15),
                 axis.line.x = element_line(size = 0.5, color = "black"))
```


```{r clean_tumor_cells2, echo = FALSE}
rm(ha_top, ht, ha_bottom, list_colors, mat_expression, name2D, name_side,
   sobj, features, features_group, group_colors, sample_info, mouse_dir,
   reduction_tsne, df_score_group1, df_score_group2, df_score_group3,
   group1_gs_name, group2_gs_name, group3_gs_name,
   upper_bound, plot_content, func_enr_list)
```


# Grafting

In this section, we make the figures related to transplantation experiment.

```{r graft_settings}
sample_info = readRDS(paste0(graft_dir, "/1_metadata/graft_sample_info.rds"))
```


## Dataset prepration

We load the tumor cells. Trajectory has been inferred on tumor cells. We load the trajectory object :

```{r load_graft}
sobj = readRDS(paste0(graft_dir, "/4_tumor_cells/graft_tumor_cells_sobj_traj_slingshot.rds"))
name2D = "mnn_dm_umap"

sobj$sample_identifier = ifelse(is.na(sobj$sample_identifier),
                                yes = "donor", no = "recipient")

sobj
```

## Zoom on tumor cells

We build the figures showing Tomato expression levels, pseudotime, sample of origin and cell cycle phase.

```{r fig5_A, class.source = 'fold-hide', fig.width = 12, fig.height = 3.75}
plot_list = list()

## Tomato expression
plot_list[[1]] = Seurat::FeaturePlot(sobj, 
                                     reduction = name2D,
                                     features = "dtTomato",
                                     pt.size = 0.25) +
  ggplot2::labs(title = "tdTomato") +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
  ggplot2::theme(legend.position = c(0.01, 0.1),
                 legend.direction = "horizontal")

## Color by sample
plot_list[[2]] = Seurat::DimPlot(sobj,
                                 group.by = "sample_identifier",
                                 reduction = name2D,
                                 pt.size = 0.25) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$sample_identifier) +
  ggplot2::labs(title = "Sample of origin") +
  ggplot2::guides(col = guide_legend(override.aes = list(size = 3))) +
  ggplot2::theme(legend.position = c(0.05, 0.1),
                 legend.title = element_blank())

## Color by cell cycle
plot_list[[3]] = Seurat::DimPlot(sobj,
                                 group.by = "Seurat.Phase",
                                 reduction = name2D,
                                 pt.size = 0.25) +
  ggplot2::scale_color_manual(values = aquarius::gg_color_hue(3),
                              labels = c("G0 / G1", "G2M", "S")) +
  ggplot2::labs(title = "Cell cycle phase") +
  ggplot2::theme(legend.position = c(0.05, 0.1),
                 legend.direction = "vertical") +
  ggplot2::guides(col = guide_legend(override.aes = list(size = 3)))

## Pseudotime
plot_list[[4]] = Seurat::FeaturePlot(sobj, pt.size = 0.25,
                                     reduction = name2D,
                                     features = "pseudotime") +
  ggplot2::scale_color_gradientn(colors = viridis::viridis(n = 100),
                                 breaks = range(sobj$pseudotime, na.rm = TRUE),
                                 labels = c("min", "max")) +
  ggplot2::labs(title = "Pseudotime") +
  ggplot2::theme(legend.position = c(0.01, 0.1),
                 legend.direction = "horizontal")

## Same theme
plot_list = lapply(plot_list, FUN = function(one_plot) {
  one_plot +
    ggplot2::theme(aspect.ratio = 1,
                   plot.title = element_text(hjust = 0.5)) +
    Seurat::NoAxes()
})

## Patchwork
patchwork::wrap_plots(plot_list, nrow = 1) &
  ggplot2::theme(legend.text = element_text(size = 12))
```


We make the feature plots :

```{r fig5_C, class.source = 'fold-hide', fig.width = 2.66, fig.height = 26}
good_order = match(genes_tom_traj, unlist(genes_heatmap)) %>%
  as.factor() %>%
  `levels<-`(c(1:length(genes_tom_traj)))

plot_list = lapply(c("dtTomato", genes_tom_traj),
                   FUN = function(one_gene) {
                     one_plot = aquarius::fig_plot_feature(sobj,
                                                           myred = name2D,
                                                           gene = one_gene,
                                                           gene_expression_colors = aquarius::color_gene,
                                                           pt.size = 0.10) +
                       ggplot2::labs(title = one_gene) +
                       ggplot2::theme(axis.title = element_blank(),
                                      axis.line = element_blank(),
                                      legend.position = "left")
                     return(one_plot)
                   })
patchwork::wrap_plots(plot_list, ncol = 1)
```


## Heatmap

We can represent gene expression levels on a heatmap.

```{r graft_heatmap_genes, class.source = "fold-hide"}
features = unique(unlist(genes_heatmap))

features[!(features %in% rownames(sobj))]
features = features[features %in% rownames(sobj)]
```

We prepare the gene expression matrix :

```{r graft_heatmap_matrix, fig.width = 12, fig.height = 30}
mat_expression = Seurat::GetAssayData(sobj, assay = "RNA", slot = "data")
mat_expression = mat_expression[features, ]
mat_expression = Matrix::t(mat_expression)
mat_expression = dynutils::scale_quantile(mat_expression) # between 0 and 1
mat_expression = Matrix::t(mat_expression)
dim(mat_expression)
```

We prepare all the colors for heatmap annotation :

```{r graft_heatmap_colors_ann, class.source = 'fold-hide'}
list_colors = list()
list_colors[["expression"]] = c("#2166AC", "#F7F7F7", "#B2182B")
list_colors[["pseudotime"]] = circlize::colorRamp2(breaks = seq(min(sobj$pseudotime, na.rm = TRUE),
                                                                max(sobj$pseudotime, na.rm = TRUE),
                                                                length.out = 9),
                                                   colors = viridis::viridis(9))
list_colors[["sample"]] = setNames(nm = sample_info$sample_identifier,
                                   sample_info$color)
list_colors[["cell_cycle_phase"]] = aquarius::gg_color_hue(3)
names(list_colors[["cell_cycle_phase"]]) = c("G1", "G2M", "S")
list_colors[["percent_mito"]] = circlize::colorRamp2(breaks = seq(from = min(sobj$percent.mt),
                                                                  to = max(sobj$percent.mt),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
list_colors[["nFeature_RNA"]] = circlize::colorRamp2(breaks = seq(from = min(sobj$nFeature_RNA),
                                                                  to = max(sobj$nFeature_RNA),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
```


We prepare the heatmap annotation :

```{r graft_heatmap_annotation, class.source = 'fold-hide'}
name_side = "right"
ha_top = HeatmapAnnotation(sample = sobj$sample_identifier,
                           pseudotime = sobj$pseudotime,
                           col = list(sample = list_colors[["sample"]],
                                      pseudotime = list_colors[["pseudotime"]]),
                           annotation_legend_param = list(pseudotime = list(at = range(sobj$pseudotime, na.rm = TRUE),
                                                                            labels = c("min", "max"))),
                           annotation_name_side = name_side)
ha_bottom = HeatmapAnnotation(percent_mito = sobj$percent.mt,
                              nFeature_RNA = sobj$nFeature_RNA,
                              cell_cycle_phase = sobj$Seurat.Phase,
                              col = list(percent_mito = list_colors[["percent_mito"]],
                                         nFeature_RNA = list_colors[["nFeature_RNA"]],
                                         cell_cycle_phase = list_colors[["cell_cycle_phase"]]),
                              annotation_name_side = name_side)
```

We build the heatmap :

```{r graft_heatmap_ht, class.source = 'fold-hide'}
ht = ComplexHeatmap::Heatmap(as.matrix(mat_expression),
                             # Expression colors
                             col = list_colors[["expression"]],
                             heatmap_legend_param = list(title = "expression", at = c(0, 1), 
                                                         labels = c("low", "high")),
                             # Annotation
                             top_annotation = ha_top,
                             bottom_annotation = ha_bottom,
                             # Cells grouping
                             column_title = NULL,
                             # Cells order
                             column_order = order(sobj$pseudotime, na.last = FALSE),
                             cluster_columns = FALSE,
                             show_column_dend = FALSE,
                             # Cells label
                             show_column_names = FALSE,
                             # Gene grouping
                             # row_split = features_group,
                             # row_gap = unit(2, "mm"),
                             # row_title = NULL,
                             cluster_rows = FALSE,
                             show_row_dend = FALSE,
                             # Gene labels
                             show_row_names = TRUE,
                             # Visual aspect
                             show_heatmap_legend = TRUE,
                             border = TRUE,
                             use_raster = FALSE)
```

```{r fig5_B, fig.width = 9, fig.height = 22}
ComplexHeatmap::draw(ht, # legend_grouping = "original",
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


```{r clean_graft, echo = FALSE}
rm(ha_bottom, ha_top, ht, list_colors, mat_expression,
   plot_list, sobj, features, name_side, name2D, good_order,
   genes_tom_traj, graft_dir)
```


# Patients

In this section, we make the figure related to NF1 patients datasets.

## Dataset preparation

We load the human dataset.

```{r load_human_dataset}
sobj_full = readRDS(paste0(patient_dir, "/3_combined/all_patients_sobj.rds"))
sobj_full
```

We set the global settings for this analysis :

```{r variable_human}
sample_info = readRDS(paste0(patient_dir, "/1_metadata/patients_sample_info.rds"))
dotplot_markers = readRDS(paste0(patient_dir, "/1_metadata/patients_dotplot_markers.rds"))
color_markers = readRDS(paste0(patient_dir, "/1_metadata/patients_color_markers.rds"))

reduction_tsne = "harmony_80_tsne"
```

This is the correspondence between cell types and cell families :

```{r human_custom_order_cell_type}
custom_order_cell_type = data.frame(cell_type = c("cDC", "mast cells", "B cells", "NK cells",
                                                  "T cells", "M\U03A6/Mo", "mural cells",
                                                  "endothelial cells", "fibroblasts", "tumor cells"),
                                    cell_type_old = c("cDC", "mast cells", "B cells", "NK cells",
                                                      "T cells", "macrophages", "mural cells",
                                                      "endothelial cells", "fibroblasts", "tumor cells"),
                                    cell_family = c(rep("immune cells", 6),
                                                    rep("stromal cells", 3),
                                                    rep("tumor cells", 1)),
                                    stringsAsFactors = FALSE)
rownames(custom_order_cell_type) = custom_order_cell_type$cell_type
```


We add those information in the object :

```{r human_relevel_cell_type}
sobj_full$cell_type = factor(sobj_full$cell_type,
                             levels = levels(sobj_full$cell_type)[match(custom_order_cell_type$cell_type_old,
                                                                        levels(sobj_full$cell_type))]) # reorder cell type
levels(sobj_full$cell_type) = custom_order_cell_type$cell_type # change level names
```

We change also levels in `color_markers` and `dotplot_markers` :

```{r human_relevel_cell_type_markers}
names(color_markers)[names(color_markers) == "macrophages"] = "M\U03A6/Mo"
color_markers = color_markers[custom_order_cell_type$cell_type]

names(dotplot_markers)[names(dotplot_markers) == "macrophages"] = "M\U03A6/Mo"
dotplot_markers = dotplot_markers[custom_order_cell_type$cell_type]
```

We group cell type by cluster, to correct mis-annotation :

```{r human_cluster_type}
cluster_type = table(sobj_full$cell_type, sobj_full$seurat_clusters) %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
cluster_type = setNames(nm = names(cluster_type),
                        levels(sobj_full$cell_type)[cluster_type])

sobj_full$cluster_type = cluster_type[sobj_full$seurat_clusters]
sobj_full$cluster_type = factor(sobj_full$cluster_type,
                                levels = levels(sobj_full$cell_type))
```

We add a `cell_family` columns in metadata :

```{r human_cell_family}
sobj_full$cell_family = custom_order_cell_type[sobj_full$cluster_type, "cell_family"]
```

## Cell type composition

This figure represent cell types annotation from all patients datasets :

```{r fig6_B, class.source = 'fold-hide', fig.width = 7, fig.height = 5}
aquarius::fig_plot_dimplot(sobj_full, reduction_tsne,
                           "cluster_type", color_markers, title = "NF1 patients")
```

This figure shows that batch-effect is removed on this projection. It also show cell cycle phase, and a quality metric : the percentage of UMI by cell, for mitochondrial genes 

```{r fig6_Z, class.source = 'fold-hide', fig.width = 13, fig.height = 6.5}
aquarius::fig_plot_quality(sobj_full, reduction_tsne,
                           sample_info = sample_info,
                           gene_expression_colors = aquarius::color_gene) %>%
  patchwork::wrap_plots(., nrow = 1)
```

This dotplot assesses the cell type annotation :

```{r fig6_Y, class.source = 'fold-hide', fig.width = 8, fig.height = 8}
markers = dotplot_markers[levels(sobj_full$cluster_type)] # new order
markers = unique(unlist(markers)) # as vector
markers = markers[!(markers %in% c("PTPRC"))]
markers = c("PTPRC", markers)
markers = markers[markers %in% rownames(sobj_full)]

aquarius::fig_plot_dotplot(sobj_full,
                           groupby = "cluster_type",
                           markers = markers,
                           gene_expression_colors = aquarius::color_gene) +
  ggplot2::geom_vline(xintercept = 0.5 + c(4, 6, 7, 9, 11, 14, 15, 17, 18, 19, 20),
                      color = "gray92")
```

```{r clean_patients, echo = FALSE}
rm(dotplot_markers, markers)
```


## Zoom on tumor cells

We load the tumor cells object :

```{r tumor_cells_patients}
sobj_tum = readRDS(paste0(patient_dir, "/4_tumor_cells/patients_tumor_cells_sobj.rds"))
name2D = "mnn_dm_umap"

sobj_tum
```



Where are those cells on the full TSNE ?

```{r fig6_D1, fig.width = 6, fig.height = 6}
sobj_full$is_tumor_cells = colnames(sobj_full) %in% colnames(sobj_tum)

Seurat::DimPlot(sobj_full, group.by = "is_tumor_cells", reduction = reduction_tsne,
                cols = c("lightgray", color_markers[["tumor cells"]])) +
  ggplot2::labs(x = "tSNE 1", y = "tSNE 2", title = "Tumor cells",
                subtitle = paste0(length(colnames(sobj_tum)), " cells")) +
  Seurat::NoLegend() + Seurat::NoAxes() +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
                 plot.subtitle = element_text(hjust = 0.5, face = "bold", size = 12),
                 aspect.ratio = 1)
```

Where are samples of origin on the tumor object ?

```{r fig6_D2, fig.width = 6, fig.height = 6}
Seurat::DimPlot(sobj_tum,
                group.by = "orig.ident",
                pt.size = 0.1,
                reduction = name2D,
                cols = setNames(sample_info$color,
                                nm = sample_info$project_name)) +
  ggplot2::labs(title = paste(ncol(sobj_tum), "tumor cells from", nrow(sample_info), "samples", sep = " ")) +
  ggplot2::guides(col = guide_legend(ncol = 1,
                                     override.aes = list(size = 2))) +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.title = element_blank(),
                 legend.background = element_rect(fill = alpha(colour = "white", alpha = 0))) +
  ggplot2::guides(col = guide_legend(override.aes = list(size = 3))) +
  Seurat::NoAxes()
```

We represent expression level for some genes of interest :

```{r fig6_D3, class.source = 'fold-hide', fig.width = 9, fig.height = 9}
plot_list = lapply(genes_patients,
                   FUN = function(one_gene) {
                     one_plot = Seurat::FeaturePlot(sobj_tum,
                                                    reduction = name2D,
                                                    features = one_gene,
                                                    pt.size = 0.10,
                                                    order = TRUE) +
                       ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
                       ggplot2::labs(title = one_gene) +
                       Seurat::NoAxes() +
                       ggplot2::theme(legend.position = "left")
                     return(one_plot)
                   })
patchwork::wrap_plots(plot_list, ncol = 3)
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```

