---
title: "Mice"
subtitle: "Impact of TME heterozygosity"
author: "Audrey"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <<- Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


This file is used to compare tumor cells between fl/fl and fl/- mouse models.

```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
library(org.Mm.eg.db)

.libPaths()
```


# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "mice_tumor_cells_subset"
out_dir = "."
```


We load the dataset :

```{r load_sobj}
sobj = readRDS(paste0(out_dir, "/", save_name, "_sobj.rds"))
sobj
```

and sample information :

```{r load_sample_info}
sample_info = readRDS(paste0(out_dir, "/../1_metadata/mice_sample_info.rds")) 
sample_info
```


This is the projection of interest :

```{r name2D}
name2D = "mnn_41_tsne"
```


# Visualization

We change clustering and visualize clusters :

```{r clustering, fig.width = 6, fig.height = 4}
Seurat::DimPlot(sobj, reduction = name2D, label = TRUE) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1)
```


and split by origin :

```{r aquarius::plot_split_dimred, fig.width = 12, fig.height = 7}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "project_name",
                                        group_by = "seurat_clusters",
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name),
                                        group_color = aquarius::gg_color_hue(length(levels(sobj$seurat_clusters))))

patchwork::wrap_plots(plot_list, nrow = 3) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right") &
  guides(color = guide_legend(override.aes = list(size = 3), title = "Cluster"))
```

We represent the projection split by genotype, and colored by sample of origin :

```{r plot_split_genotype, fig.width = 10, fig.height = 7}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "genotype",
                                        split_color = rep("black", 2),
                                        group_by = "project_name",
                                        group_color = sample_info$color,
                                        bg_color = "gray70")

patchwork::wrap_plots(plot_list, nrow = 1) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "bottom",
                 legend.direction = "vertical") &
  ggplot2::guides(color = guide_legend(override.aes = list(size = 3), ncol = 7))
```

We represent genes of interest on the projection. Clusters without Sox10 and Sox9 expression express a high level of ribosomal genes.

```{r plot_genes, fig.width = 12, fig.height = 4}
plot_list = lapply(c("Sox10", "Birc5", "Sox9"), FUN = function(one_gene) {
  p = Seurat::FeaturePlot(sobj, features = one_gene,
                          reduction = name2D) +
    ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
    ggplot2::theme(aspect.ratio = 1) +
    Seurat::NoAxes()
  return(p)
})

patchwork::wrap_plots(plot_list, nrow = 1)
```

# Annotation preparation

In this section, we prepare the gene annotation. Gene annotation is stored in the object :

```{r gene_corresp}
gene_corresp = sobj@assays[["RNA"]]@meta.features[, c("gene_name", "Ensembl_ID")]
names(gene_corresp) = c("NAME", "ID")

head(gene_corresp)
```


We get signatures from MSigDB :

```{r msigdb_sets}
gene_sets = msigdbr::msigdbr(species = "Mus musculus") %>%
  dplyr::filter(gs_cat %in% c("H", "C2", "C5")) %>%
  dplyr::filter(gs_subcat %in% c("", "GO:BP", "GO:MF", "GO:CC", "CP:KEGG", "CP:BIOCARTA",
                                 "CP:WIKIPATHWAYS", "CP:REACTOME", "CP:PID")) %>%
  dplyr::select(gs_cat, gs_subcat, gs_name, ensembl_gene, gene_symbol)

head(gene_sets)

table(gene_sets$gs_subcat, gene_sets$gs_cat)
```


# Differential expression

What is the identity of specific clusters ? Results will be stored in the `list_results` list object.

```{r list_results}
list_results = list()
```


## Glial population

Glial populations are those clusters :

```{r glial_clusters}
clusters_oi = cbind(sobj@meta.data,
                    glial = Seurat::FetchData(sobj, "Sox10")[, 1]) %>%
  dplyr::select(seurat_clusters, glial) %>%
  dplyr::group_by(seurat_clusters) %>%
  dplyr::summarise(avg_score = mean(glial)) %>%
  as.data.frame() %>%
  dplyr::filter(avg_score > 0.75) %>%
  dplyr::pull(seurat_clusters) %>%
  as.character()

clusters_oi
```

We set specific groups to compare this population between fl/fl and fl/- :

```{r glial_ident}
Seurat::Idents(sobj) = ifelse(sobj$seurat_clusters %in% clusters_oi,
                              no = "other",
                              yes = ifelse(sobj$genotype == "flfl",
                                           yes = "glial_flfl", no = "glial_flmn"))
Seurat::Idents(sobj) = ifelse(sobj$sample_type == "flfl_pNF",
                              yes = "other",
                              no = as.character(Seurat::Idents(sobj)))

table(sobj$project_name, Seurat::Idents(sobj))
```

We will compare those groups :

```{r glial_red_blue, fig.width = 5, fig.height = 5}
group1 = "glial_flfl"
group2 = "glial_flmn"
group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We identify differentially expressed markers between both groups :

```{r glial_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```

### Up in fl/fl

```{r glial_up, fig.width = 5, fig.height = 5}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

list_results[[group_name]]$up_genes = genes_of_interest

dim(genes_of_interest)

head(genes_of_interest, n = 20)
```


We explore enrichment in gene sets :

```{r glial_ora_up, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(gene_names = rownames(genes_of_interest),
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$up_enrichr = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


### Up in fl/-

```{r glial_down, fig.width = 5, fig.height = 5}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, pct.2 - pct.1)

list_results[[group_name]]$dn_genes = genes_of_interest

dim(genes_of_interest)

head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r glial_ora_dn, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(gene_names = rownames(genes_of_interest),
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$dn_enrichr = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


### GSEA

We run a GSEA for each significant gene sets, using the full count matrix. First, we compute fold change between both groups :

```{r glial_foldchange}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```

From previous ORA, we extract significantly enriched gene sets :

```{r glial_gene_sets, fig.width = 12, fig.height = 25}
gene_sets_of_interest = c(
  list_results[[group_name]]$up_enrichr@result %>%
    dplyr::filter(p.adjust < 0.05) %>%
    dplyr::pull(ID),
  list_results[[group_name]]$dn_enrichr@result %>%
    dplyr::filter(p.adjust < 0.05) %>%
    dplyr::pull(ID)) %>% unique()

length(gene_sets_of_interest)
```

We run GSEA for each gene sets :

```{r glial_gsea, fig.width = 12, fig.height = 5}
gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gene_sets[gene_sets$gs_name %in%
                                                          gene_sets_of_interest, c("gs_name", "ensembl_gene")],
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```


We look at the GSEA plot for some gene sets :

```{r glial_gsea_plot, fig.width = 12, fig.height = 23}
# grep(gene_sets_of_interest, pattern = "RHO", value = TRUE)
gene_sets_plot = c("GOCC_COLLAGEN_CONTAINING_EXTRACELLULAR_MATRIX",
                   "REACTOME_SIGNALING_BY_ROBO_RECEPTORS",
                   "REACTOME_REGULATION_OF_EXPRESSION_OF_SLITS_AND_ROBOS",
                   "HALLMARK_INTERFERON_GAMMA_RESPONSE",
                   "REACTOME_INTERFERON_SIGNALING",
                   "GOCC_CELL_SUBSTRATE_JUNCTION",
                   "KEGG_FOCAL_ADHESION",
                   "GOMF_CADHERIN_BINDING",
                   "REACTOME_L1CAM_INTERACTIONS",
                   "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION")

plot_list = lapply(gene_sets_plot, FUN = function(gene_set) {
  gs_size = gene_sets %>%
    dplyr::filter(gs_name == gene_set) %>%
    dplyr::pull(ensembl_gene) %>%
    as.character() %>% unique() %>% length()
  
  p = enrichplot::gseaplot2(gsea_results, geneSetID = gene_set) +
    ggplot2::labs(title = gene_set,
                  subtitle = paste0(gs_size, " genes")) +
    ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold"),
                   plot.subtitle = element_text(hjust = 0.5))
  return(p)
})
patchwork::wrap_plots(plot_list, ncol = 2)
```


## Mesenchymal population

Mesenchymal populations are those clusters :

```{r mes_clusters}
clusters_oi = cbind(sobj@meta.data,
                    mes = Seurat::FetchData(sobj, "Dcn")[, 1]) %>%
  dplyr::select(seurat_clusters, mes) %>%
  dplyr::group_by(seurat_clusters) %>%
  dplyr::summarise(avg_score = mean(mes)) %>%
  as.data.frame() %>%
  dplyr::filter(avg_score > 0.85) %>%
  dplyr::pull(seurat_clusters) %>%
  as.character()

clusters_oi
```

We set specific groups to compare this population between fl/fl and fl/- :

```{r mes_ident}
Seurat::Idents(sobj) = ifelse(sobj$seurat_clusters %in% clusters_oi,
                              no = "other",
                              yes = ifelse(sobj$genotype == "flfl",
                                           yes = "mes_flfl", no = "mes_flmn"))

table(sobj$project_name, Seurat::Idents(sobj))
```

We will compare those groups :

```{r mes_red_blue, fig.width = 5, fig.height = 5}
group1 = "mes_flfl"
group2 = "mes_flmn"
group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We identify differentially expressed markers between both groups :

```{r mes_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```

### Up in fl/fl

```{r mes_up, fig.width = 5, fig.height = 5}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

list_results[[group_name]]$up_genes = genes_of_interest

dim(genes_of_interest)

head(genes_of_interest, n = 20)
```


We explore enrichment in gene sets :

```{r mes_ora_up, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(gene_names = rownames(genes_of_interest),
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$up_enrichr = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


### Up in fl/-

```{r mes_dn, fig.width = 5, fig.height = 5}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, pct.2 - pct.1)

list_results[[group_name]]$dn_genes = genes_of_interest

dim(genes_of_interest)

head(genes_of_interest, n = 20)
```

We explore enrichment in gene sets :

```{r mes_ora_dn, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(gene_names = rownames(genes_of_interest),
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$dn_enrichr = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


### GSEA

We run a GSEA for each significant gene sets, using the full count matrix. First, we compute fold change between both groups :

```{r mes_foldchange}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```

From previous ORA, we extract significantly enriched gene sets :

```{r mes_goi, fig.width = 12, fig.height = 25}
gene_sets_of_interest = c(
  list_results[[group_name]]$up_enrichr@result %>%
    dplyr::filter(p.adjust < 0.05) %>%
    dplyr::pull(ID),
  list_results[[group_name]]$dn_enrichr@result %>%
    dplyr::filter(p.adjust < 0.05) %>%
    dplyr::pull(ID)) %>% unique()

length(gene_sets_of_interest)
```

We run GSEA for each gene sets :

```{r mes_gsea, fig.width = 12, fig.height = 4}
gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gene_sets[gene_sets$gs_name %in%
                                                          gene_sets_of_interest, c("gs_name", "ensembl_gene")],
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

We look at the GSEA plot for some gene sets :

```{r mes_gsea_plot, fig.width = 12, fig.height = 25}
# grep(gene_sets_of_interest, pattern = "RHO", value = TRUE)
gene_sets_plot = c("WP_NEUROINFLAMMATION",
                   "GOCC_COLLAGEN_CONTAINING_EXTRACELLULAR_MATRIX",
                   "HALLMARK_TNFA_SIGNALING_VIA_NFKB",
                   "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",
                   "WP_IL18_SIGNALING_PATHWAY",
                   "HALLMARK_APOPTOSIS",
                   "HALLMARK_HYPOXIA",
                   "HALLMARK_P53_PATHWAY",
                   "GOBP_REGULATION_OF_SIGNAL_TRANSDUCTION_BY_P53_CLASS_MEDIATOR",
                   "PID_P53_DOWNSTREAM_PATHWAY",
                   "GOBP_RESPONSE_TO_OXIDATIVE_STRESS",
                   "HALLMARK_INTERFERON_ALPHA_RESPONSE",
                   "HALLMARK_INTERFERON_GAMMA_RESPONSE",
                   "GOMF_EXTRACELLULAR_MATRIX_BINDING",
                   "GOBP_REGULATION_OF_EXTRACELLULAR_MATRIX_ORGANIZATION",
                   "REACTOME_SIGNALING_BY_ROBO_RECEPTORS",
                   "REACTOME_REGULATION_OF_EXPRESSION_OF_SLITS_AND_ROBOS")

plot_list = lapply(gene_sets_plot, FUN = function(gene_set) {
  gs_size = gene_sets %>%
    dplyr::filter(gs_name == gene_set) %>%
    dplyr::pull(ensembl_gene) %>%
    as.character() %>% unique() %>% length()
  
  p = enrichplot::gseaplot2(gsea_results, geneSetID = gene_set) +
    ggplot2::labs(title = gene_set,
                  subtitle = paste0(gs_size, " genes")) +
    ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold"),
                   plot.subtitle = element_text(hjust = 0.5))
  return(p)
})
patchwork::wrap_plots(plot_list, ncol = 2)
```


## Other analyses

### Specific to `2022_23`

Are they specific enriched pathways in this sample, where Cdkn2a is still expressed, but Trp53 is lost ? First, we compare the glial population, then the mesenchymal. We compare gene expression only in fl/- datasets.

#### Glial

We extract glial clusters :

```{r 2022_23_glial}
clusters_oi = cbind(sobj@meta.data,
                    glial = Seurat::FetchData(sobj, "Sox10")[, 1]) %>%
  dplyr::select(seurat_clusters, glial) %>%
  dplyr::group_by(seurat_clusters) %>%
  dplyr::summarise(avg_score = mean(glial)) %>%
  as.data.frame() %>%
  dplyr::filter(avg_score > 0.85) %>%
  dplyr::pull(seurat_clusters) %>%
  as.character()

clusters_oi
```

We set specific groups to compare this population between `2022_23` and other fl/- :

```{r 2022_23_glial_ident}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" & sobj$project_name == "2022_23" ~ "A",
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" & sobj$project_name != "2022_23" ~ "B")

table(sobj$project_name, Seurat::Idents(sobj))
```


We will compare those groups :

```{r 2022_23_glial_redblue, fig.width = 5, fig.height = 5}
group1 = "A"
group2 = "B"
group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We identify differentially expressed markers between both groups :

```{r 2022_23_glial_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```



##### Up in `2022_23`

```{r 2022_23_glial_up, fig.width = 5, fig.height = 5}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

list_results[[group_name]]$up_genes = genes_of_interest

dim(genes_of_interest)

head(genes_of_interest, n = 20)
```


We explore enrichment in gene sets :

```{r 2022_23_glial_ora_up, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(gene_names = rownames(genes_of_interest),
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


##### Down in `2022_23`

```{r 2022_23_glial_dn, fig.width = 5, fig.height = 5}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

list_results[[group_name]]$dn_genes = genes_of_interest

dim(genes_of_interest)

head(genes_of_interest, n = 20)
```


We explore enrichment in gene sets :

```{r 2022_23_glial_ora_dn, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(gene_names = rownames(genes_of_interest),
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


#### Mesenchymal

We extract mesenchymal clusters :

```{r 2022_23_mesenchymal}
clusters_oi = cbind(sobj@meta.data,
                    mesenchymal = Seurat::FetchData(sobj, "Dcn")[, 1]) %>%
  dplyr::select(seurat_clusters, mesenchymal) %>%
  dplyr::group_by(seurat_clusters) %>%
  dplyr::summarise(avg_score = mean(mesenchymal)) %>%
  as.data.frame() %>%
  dplyr::filter(avg_score > 0.85) %>%
  dplyr::pull(seurat_clusters) %>%
  as.character()

clusters_oi = clusters_oi[clusters_oi != 15]
clusters_oi
```

We set specific groups to compare this population between `2022_23` and other fl/- :

```{r 2022_23_mesenchymal_ident}
Seurat::Idents(sobj) = case_when(
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" & sobj$project_name == "2022_23" ~ "A",
  sobj$seurat_clusters %in% clusters_oi & sobj$genotype == "flmn" & sobj$project_name != "2022_23" ~ "B")

table(sobj$project_name, Seurat::Idents(sobj))
```


We will compare those groups :

```{r 2022_23_mesenchymal_redblue, fig.width = 5, fig.height = 5}
group1 = "A"
group2 = "B"
group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We identify differentially expressed markers between both groups :

```{r 2022_23_mesenchymal_de, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1, ident.2 = group2)

mark %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC)) %>%
  dplyr::arrange(avg_logFC)

mark %>%
  dplyr::filter(abs(pct.1 - pct.2) > 0.3) %>%
  arrange(avg_logFC)
```



##### Up in `2022_23`

```{r 2022_23_mesenchymal_up, fig.width = 5, fig.height = 5}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

list_results[[group_name]]$up_genes = genes_of_interest

dim(genes_of_interest)

head(genes_of_interest, n = 20)
```


We explore enrichment in gene sets :

```{r 2022_23_mesenchymal_ora_up, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(gene_names = rownames(genes_of_interest),
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


##### Down in `2022_23`

```{r 2022_23_mesenchymal_dn, fig.width = 5, fig.height = 5}
genes_of_interest = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) %>%
  dplyr::arrange(avg_logFC, -(pct.1 - pct.2))

list_results[[group_name]]$dn_genes = genes_of_interest

dim(genes_of_interest)

head(genes_of_interest, n = 20)
```


We explore enrichment in gene sets :

```{r 2022_23_mesenchymal_ora_dn, fig.width = 12, fig.height = 8}
enrichr_results = aquarius::run_enrichr(gene_names = rownames(genes_of_interest),
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


# Genes of interest

```{r plot_kcna1, fig.width = 12, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj, reduction = name2D,
                                        split_by = "project_name",
                                        split_color = sample_info$color,
                                        color_by = "Kcna1")
patchwork::wrap_plots(plot_list, nrow = 3) + patchwork::plot_layout(guides = "collect")
```

```{r plot_jade1, fig.width = 12, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj, reduction = name2D,
                                        split_by = "project_name",
                                        split_color = sample_info$color,
                                        color_by = "Jade1")
patchwork::wrap_plots(plot_list, nrow = 3) + patchwork::plot_layout(guides = "collect")
```

```{r plot_postn, fig.width = 12, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj, reduction = name2D,
                                        split_by = "project_name",
                                        split_color = sample_info$color,
                                        color_by = "Postn")
patchwork::wrap_plots(plot_list, nrow = 3) + patchwork::plot_layout(guides = "collect")
```

```{r plot_cdkn2a, fig.width = 12, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj, reduction = name2D,
                                        split_by = "project_name",
                                        split_color = sample_info$color,
                                        color_by = "Cdkn2a")
patchwork::wrap_plots(plot_list, nrow = 3) + patchwork::plot_layout(guides = "collect")
```

```{r plot_trp53, fig.width = 12, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj, reduction = name2D,
                                        split_by = "project_name",
                                        split_color = sample_info$color,
                                        color_by = "Trp53")
patchwork::wrap_plots(plot_list, nrow = 3) + patchwork::plot_layout(guides = "collect")
```

```{r plot_eed, fig.width = 12, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj, reduction = name2D,
                                        split_by = "project_name",
                                        split_color = sample_info$color,
                                        color_by = "Eed")
patchwork::wrap_plots(plot_list, nrow = 3) + patchwork::plot_layout(guides = "collect")
```

```{r plot_suz12, fig.width = 12, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj, reduction = name2D,
                                        split_by = "project_name",
                                        split_color = sample_info$color,
                                        color_by = "Suz12")
patchwork::wrap_plots(plot_list, nrow = 3) + patchwork::plot_layout(guides = "collect")
```

```{r plot_ezh1, fig.width = 12, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj, reduction = name2D,
                                        split_by = "project_name",
                                        split_color = sample_info$color,
                                        color_by = "Ezh1")
patchwork::wrap_plots(plot_list, nrow = 3) + patchwork::plot_layout(guides = "collect")
```

```{r plot_ezh2, fig.width = 12, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj, reduction = name2D,
                                        split_by = "project_name",
                                        split_color = sample_info$color,
                                        color_by = "Ezh2")
patchwork::wrap_plots(plot_list, nrow = 3) + patchwork::plot_layout(guides = "collect")
```

```{r plot_mtap, fig.width = 12, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj, reduction = name2D,
                                        split_by = "project_name",
                                        split_color = sample_info$color,
                                        color_by = "Mtap")
patchwork::wrap_plots(plot_list, nrow = 3) + patchwork::plot_layout(guides = "collect")
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```

