---
title: "Mice"
subtitle: "Tumor cells functional enrichment"
author: "Audrey"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <<- Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


This script is used to :

* explore MSigDB gene sets : hallmark, GO BP, GO MF, GO CC, Kegg pathways, PID, Wiki pathways and Biocarta gene sets
* do functional enrichment and GSE analysis with `clusterProfiler`
* apply score to MSigDB gene sets with `AUCell::AUCell_calcAUC`

We load the Seurat object containing all tumor cells from Nf1-KO mice. Functional enrichment analysis needs genes. To do so, we make a clustering of tumor cells. We make two alternatives :

* keep a lot of clusters, even if they are not biologically meaningful, or ;
* group clusters into 3 main groups (glial, cycling, mesenchymal).

We made **differential expression** analysis (DEA) for each group, and for clusters along trajectory pseudotime. **Functional enrichment** analysis (FEA) is much faster than **Gene Set Enrichment Analysis** (GSEA). We used FEA to explore the full gene sets databases, and GSEA to better quantify the enrichment for the best ones.


The principle of FEA is that we make a DEA between two groups (A and B). We keep either upregulated genes in A compared to B, or downregulated genes in A compared to B. The comparison is symmetrical : upregulated genes in A are downregulated in B. We defined a gene as upregulated in A when fold change is positive and adjusted p-value is lower than 0.01. We gave to the FEA pipeline the list of upregulated or downregulated genes, and a database with gene sets. FEA will look at the overlaps between the list of genes we gave, and the genes in the gene set. For each gene set, FEA compute a `GeneRatio` : the number of genes in our list that are in the gene set ($k$) compared to the number of genes in the gene set ($K$). The enrichment p-value is computed from a hypergeometric distribution.

We have $N$ genes, the total number of genes. We know from the database that $K$ genes among the $N$ are our gene of interest, and represent a particular gene set. We draw $n$ genes among the $N$ : $n$ is the size of the differentially expressed genes. Among the $n$ genes, $k$ belong to the $K$ genes of interest. The enrichment p-value is computed as the p-value to obtain $k$ successes in $n$ draws, from a population of genes of size $N$ containing $K$ genes of interest.


There are several ways to run GSEA. Here, we provide to the GSEA pipeline the full list of genes, with their fold change. Information is : gene name, fold change between group A and group B. All genes are ordered according to fold change. Then, genes present in a gene set of interest are labeled : black sticks on the bottom of the GSEA figure. A **enrichment score** (ES) is computed. Enrichment curve initializes at 0. Then, we travel along the fold change axis. Each time we meet a labeled gene (gene belonging to the gene set), the enrichment increases. While traveling along the fold change axis without meeting any genes belonging the gene set, the enrichment decreases. The enrichment score corresponds to the maximum of the enrichment curve. The **normalized enrichment score** (NES) is computed by permuting genes along the fold change axis. Then, while ES is fixed, NES is subjected to randomness. NES is either positive (gene set in enriched) or negative (gene set is depleted). A good NES has a high absolute value, as -2 or 2. Finally, GSEA computed a **False Discovery Rate**. It generated a distribution of all NES generated by all permutations, and look where is our NES in this distribution. FDR belongs to the $[0,1]$ range. The interpretation is the same as the p-value : it should be close to 0 to be statistically significant.


In the following sections, we made three analysis :

1. **Functional enrichment** : We explore the 8 gene set databases with `clusterProfiler`. For each group's upregulated and downregulated, we keep top 30 enriched gene sets, ranked by the `GeneRatio` metric. The differentially expressed genes are kept the their adjusted p-value of differential expression in lower than 0.01.
2. **GSEA** : For each comparison, we keep significantly enriched gene sets from FEA (pvalue < 0.01). Then, on the barplot, only gene sets containing more than 5 genes, less than 500 genes, and found enriched in our comparison with a p-value lower than 0.05 are kept.
3. **AUCell** : GSEA and functional enrichment are well suited to compare bulk datasets. With single-cell datasets, we have gene expression for a lot of individual cells. We would like to visualize the progressive changes in gene sets expression, along pseudotime. To do this, we need to attribute a score for each cell and each gene set. We used the package `AUCell` to compute those score (note : `Seurat::AddModuleScore` gave similar results, but it longer to run). For each of the 8 database, the gene sets are taken as the union of all top 30 gene sets that came from all the functional enrichment analyses.

```{r library}
library(dplyr)
library(org.Mm.eg.db)
library(ComplexHeatmap)
library(ggplot2)

.libPaths()
```

# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "mice_tumor_cells_subset"
out_dir = "."
n_threads = 6 # for AUCell::AUCell_calcAUC
```


We load the sample information :

```{r custom_palette_sample, fig.width = 8, fig.height = 8}
sample_info = readRDS(paste0(out_dir, "/../1_metadata/mice_sample_info.rds"))
project_names_oi = sample_info$project_name

graphics::pie(rep(1, nrow(sample_info)),
              col = sample_info$color,
              labels = sample_info$project_name)
```

We load the Seurat object :

```{r load_sobj}
sobj = readRDS(paste0(out_dir, "/", save_name, "_sobj_traj_slingshot.rds"))
sobj
```

This is the projection name to represent cells :

```{r name2D}
name2D = "mnn_dm_umap"
```


# Clustering

We can visualize sample of origin of tumor cells, on the UMAP, and the pseudotime.

```{r dimplot, fig.width = 12, fig.height = 5}
sample_of_origin = Seurat::DimPlot(sobj,
                                   group.by = "project_name",
                                   reduction = name2D) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$sample_identifiant) +
  ggplot2::labs(title = "Sample of origin") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
  Seurat::NoAxes()

pseudotime = Seurat::FeaturePlot(sobj,
                                 features = "pseudotime",
                                 reduction = name2D) +
  ggplot2::scale_color_gradientn(colors = viridis::viridis(n = 100)) +
  ggplot2::theme(aspect.ratio = 1) +
  Seurat::NoAxes()

sample_of_origin | pseudotime
```

We cluster cells into three groups, using k-means :

```{r clustering_big, fig.width = 8, fig.height = 8}
# kmeans
kmeans_obj = sobj@reductions[["mnn"]]@cell.embeddings %>%
  as.data.frame()

set.seed(1337L)
kmeans_obj = stats::kmeans(kmeans_obj, centers = 3, nstart = 10, 
                           algorithm = "Hartigan-Wong", iter.max = 500)

# add column in method data
sobj$kmeans_group = kmeans_obj$cluster[colnames(sobj)]

# relevel according to pseudotime
kmeans_group_levels = sobj@meta.data[, c("kmeans_group", "pseudotime")] %>%
  dplyr::group_by(kmeans_group) %>%
  dplyr::summarise(avg_pdt = mean(pseudotime)) %>%
  dplyr::arrange(avg_pdt) %>%
  dplyr::pull(kmeans_group)

sobj$kmeans_group = factor(sobj$kmeans_group, levels = kmeans_group_levels)
levels(sobj$kmeans_group) = c("group 1", "group 2", "group 3")

# How many cells by group ?
table(sobj$kmeans_group)

# Visualization
Seurat::DimPlot(sobj,
                reduction = name2D,
                group.by = "kmeans_group",
                label = TRUE,
                cols = aquarius::gg_color_hue(3)) +
  Seurat::NoAxes() + Seurat::NoLegend() +
  ggplot2::theme(aspect.ratio = 1)
```

We can also have a lot of groups :

```{r clustering, fig.width = 8, fig.height = 8}
# kmeans
kmeans_obj = sobj@reductions[["mnn"]]@cell.embeddings %>%
  as.data.frame()

set.seed(1337L)
kmeans_obj = stats::kmeans(kmeans_obj, centers = 10, nstart = 10, 
                           algorithm = "Hartigan-Wong", iter.max = 500)

# add column in method data
sobj$small_clusters = kmeans_obj$cluster[colnames(sobj)]

# relevel according to pseudotime
small_clusters_levels = sobj@meta.data[, c("small_clusters", "pseudotime")] %>%
  dplyr::group_by(small_clusters) %>%
  dplyr::summarise(avg_pdt = mean(pseudotime)) %>%
  dplyr::arrange(avg_pdt) %>%
  dplyr::pull(small_clusters)

# replace plexiform as cluster 0
sobj$small_clusters = ifelse(sobj$sample_type == "flfl_pNF",
                             yes = 0,
                             no = sobj$small_clusters)

sobj$small_clusters = factor(sobj$small_clusters,
                             levels = c(0, small_clusters_levels))
levels(sobj$small_clusters) = c(0:length(small_clusters_levels))

# How many cells by group ?
table(sobj$small_clusters)

# Visualization
Seurat::DimPlot(sobj,
                reduction = name2D,
                group.by = "small_clusters",
                label = TRUE,
                cols = aquarius::gg_color_hue(11)) +
  Seurat::NoAxes() + Seurat::NoLegend() +
  ggplot2::theme(aspect.ratio = 1)
```

# Annotation preparation

In this section, we prepare the gene annotation. We load gene sets of interest from MSigDB.

Gene annotation is stored in the object :

```{r gene_corresp}
gene_corresp = sobj@assays[["RNA"]]@meta.features[, c("Ensembl_ID", "gene_name")]
names(gene_corresp) = c("ENTREZID", "NAME")

head(gene_corresp)
```


We get signatures from MSigDB :

```{r msigdb_sets}
gene_sets = msigdbr::msigdbr(species = "Mus musculus") %>%
  dplyr::filter(gs_cat %in% c("H", "C2", "C5")) %>%
  dplyr::filter(gs_subcat %in% c("", "GO:BP", "GO:MF", "GO:CC", "CP:KEGG", "CP:BIOCARTA",
                                 "CP:WIKIPATHWAYS", "CP:REACTOME", "CP:PID")) %>%
  dplyr::select(gs_cat, gs_subcat, gs_name, ensembl_gene, gene_symbol)

head(gene_sets)

table(gene_sets$gs_subcat, gene_sets$gs_cat)
```

# Between groups

In this section, we analyse the three groups. We change the active identifier :

```{r kmeans_group_kmeans_ident}
Seurat::Idents(sobj) = sobj$kmeans_group

table(sobj$sample_identifier,
      Seurat::Idents(sobj)) %>%
  rbind(., table(Seurat::Idents(sobj)))
```

We store results in the `list_result` list :

```{r initialization}
list_results = list()
```

## Group 1 vs Group 3

```{r see_blue_red_group1, fig.width = 5, fig.height = 5}
group1 = "group 1"
group2 = "group 3"

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r de_group1, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r foldchange_group1}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ENTREZID"]

head(ranked_gene_list)
```

To what extent Ensembl ID are available ?

```{r venn_ensembl, fig.width = 8, fig.height = 5}
ggvenn::ggvenn(list(our_ensembl_id = names(ranked_gene_list),
                    msigdb_ensembl_id = gene_sets$ensembl_gene))
```


### Up in group 1

We explore enrichment in gene sets :

```{r ora_group1, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r gsea_plot_full_group1, fig.width = 12, fig.height = 30}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

We make the GSEA plot for some signaling pathways :

```{r plot_gsea_group1, fig.width = 12, fig.height = 10}
gs_oi = c("GOBP_ENSHEATHMENT_OF_NEURONS",
          "GOBP_RESPONSE_TO_AXON_INJURY",
          "WP_NEUROINFLAMMATION_AND_GLUTAMATERGIC_SIGNALING",
          "GOBP_AXON_DEVELOPMENT")

plot_list = lapply(gs_oi, FUN = function(one_gene_set) {
  NES = gsea_results@result[one_gene_set, "NES"]
  p.adjust = gsea_results@result[one_gene_set, "p.adjust"]
  qvalues = gsea_results@result[one_gene_set, "qvalues"]
  
  my_subtitle = paste0("\nNES : ", round(NES, 2),
                       " | padj : ", round(p.adjust, 4),
                       " | qval : ", round(qvalues, 4))
  
  p = enrichplot::gseaplot2(gsea_results, one_gene_set) +
    ggplot2::labs(title = one_gene_set,
                  subtitle = my_subtitle) +
    ggplot2::theme(plot.title = element_text(hjust = 0.5),
                   plot.subtitle = element_text(hjust = 0.5, size = 10))
  return(p)
})

patchwork::wrap_plots(plot_list, ncol = 2)
```


### Up in group 3

We explore enrichment in gene sets :

```{r ora_group3, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r gsea_plot_full_group3, fig.width = 12, fig.height = 70}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

We make the GSEA plot for some signaling pathways :

```{r plot_gsea_group3, fig.width = 12, fig.height = 10}
gs_oi = c("GOBP_BONE_MINERALIZATION",
          "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",
          "HALLMARK_TGF_BETA_SIGNALING",
          "GOMF_EXTRACELLULAR_MATRIX_STRUCTURAL_CONSTITUENT")

plot_list = lapply(gs_oi, FUN = function(one_gene_set) {
  NES = gsea_results@result[one_gene_set, "NES"]
  p.adjust = gsea_results@result[one_gene_set, "p.adjust"]
  qvalues = gsea_results@result[one_gene_set, "qvalues"]
  
  my_subtitle = paste0("\nNES : ", round(NES, 2),
                       " | padj : ", round(p.adjust, 4),
                       " | qval : ", round(qvalues, 4))
  
  p = enrichplot::gseaplot2(gsea_results, one_gene_set) +
    ggplot2::labs(title = one_gene_set,
                  subtitle = my_subtitle) +
    ggplot2::theme(plot.title = element_text(hjust = 0.5),
                   plot.subtitle = element_text(hjust = 0.5, size = 10))
  return(p)
})

patchwork::wrap_plots(plot_list, ncol = 2)
```

## Group 1 vs Group 2

```{r see_blue_red_group1_2, fig.width = 5, fig.height = 5}
group1 = "group 1"
group2 = "group 2"

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r de_group1_2, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r foldchange_group1_2}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ENTREZID"]

head(ranked_gene_list)
```


### Up in group 1

We explore enrichment in gene sets :

```{r ora_group1_2_up, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r gsea_plot_full_group1_2_up, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in group 2

We explore enrichment in gene sets :

```{r ora_group1_2_down, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r gsea_plot_full_group1_2_down, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

## Group 3 vs Group 2

```{r see_blue_red_group3_2, fig.width = 5, fig.height = 5}
group1 = "group 3"
group2 = "group 2"

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r de_group3_2, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r foldchange_group3_2}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ENTREZID"]

head(ranked_gene_list)
```


### Up in group 3

We explore enrichment in gene sets :

```{r ora_group3_2_up, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r gsea_plot_full_group3_2_up, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in group 2

We explore enrichment in gene sets :

```{r ora_group3_2_down, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r gsea_plot_full_group3_2_down, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```


# Between clusters

We are going to perform functional analysis between clusters, along pseudotime. These are the clusters :

```{r see_all_clusters, fig.width = 8, fig.height = 8}
Seurat::Idents(sobj) = sobj$small_clusters

custom_cols = aquarius::gg_color_hue(length(levels(sobj$small_clusters)))
names(custom_cols) = levels(sobj$small_clusters)

Seurat::DimPlot(sobj,
                group.by = "small_clusters",
                reduction = name2D,
                label = TRUE,
                cols = custom_cols) +
  Seurat::NoLegend() + Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1)
```

We will compare all clusters, in the pseudotime order :

```{r clusters_in_pdt_order}
comparison = data.frame(id1 = c(0, 2, 3, 5, 6, 7, 8, 9),
                        id2 = c(2, 3, 5, 6, 7, 8, 9, 10),
                        stringsAsFactors = FALSE)
comparison
```


```{r init_id, echo = FALSE}
id = 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ENTREZID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ENTREZID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ENTREZID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ENTREZID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ENTREZID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ENTREZID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ENTREZID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ENTREZID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group1, 
                      " compared to clusters ",
                      paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(
  gene_names = genes_of_interest,
  gene_corresp = gene_corresp,
  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")] %>%
    `colnames<-`(c("gs_name", "entrez_gene")),
  make_plot = TRUE,
  plot_title = paste0("Up-regulated in cluster ", group2, 
                      " compared to clusters ",
                      paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```


# Save

We save the Seurat object, containing cluster ID.

```{r save_sobj}
saveRDS(sobj, file = paste0(out_dir, "/", save_name, "_sobj_traj_slingshot_FE.rds"))
```

We save the `list_results` object :

```{r save_list_results}
saveRDS(list_results, file = paste0(out_dir, "/", save_name, "_list_results.rds"))
```


# R session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```
