---
title: "Welcome"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE,
                      eval = FALSE)
```


This site gathers and describes all scripts used for the analysis of single-cell RNA sequencing (scRNA-Seq), from raw count matrices to figure presented in the publication.

<br>
<center>Radomska *et al.*, Glial-to-mesenchymal transition of tumor Schwann cells drives the genetic burden in MPNSTs from Neurofibromatosis type 1 mouse model, <br>
Science Advances; doi: *not yet available*</center>
<br>

All computational analyses were conducted using R language within RStudio. R Markdown (RMD) files were compiled to generate static HTML files, which are shared through this site. Once a file is open, a link in the top left corner enables to open it in a new tab. In the top right corner, the button "Code" allows to download the RMD file behind the HTML.

# Data availability

| Content               |      Availability      |
|:---------------------:|:----------------------:|
| scRNA-Seq data        |  ArrayExpress, accession number [E-MTAB-14222](https://www.ebi.ac.uk/biostudies/arrayexpress/studies/E-MTAB-14222) |
| R Markdown files      |  GitHub repository hosting this website |
| Singularity container |  Zenodo, record ID [15512051](https://doi.org/10.5281/zenodo.15512051) |

# Data presentation

Several datasets were analyzed in the article, whose main characteristics are described in the tables below.

```{r sample_info1, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 7}
library(dplyr)

data_dir = "./.."
sample_info = readRDS(paste0(data_dir, "/1_mice/1_metadata/mice_sample_info.rds"))

# patchwork
aquarius::fig_plot_gb(sample_info, title = "Primary mouse datasets") %>%
  patchwork::wrap_plots() +
  patchwork::plot_layout(design = "A\nB", heights = c(0.1,5)) &
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5, face = "bold", size = 15))
```

```{r sample_info2, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 2}
sample_info = readRDS(paste0(data_dir, "/3_graft/1_metadata/graft_sample_info.rds"))

# patchwork
aquarius::fig_plot_gb(sample_info, title = "Donor / Recipient mouse datasets") %>%
  patchwork::wrap_plots() +
  patchwork::plot_layout(design = "A\nB", heights = c(0.1,5)) &
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5, face = "bold", size = 15))
```

```{r sample_info3, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 7}
sample_info = readRDS(paste0(data_dir, "/2_patients/1_metadata/patients_sample_info.rds"))

# patchwork
aquarius::fig_plot_gb(sample_info, title = "Primary NF1 patients datasets") %>%
  patchwork::wrap_plots() +
  patchwork::plot_layout(design = "A\nB", heights = c(0.1,5)) &
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5, face = "bold", size = 15))
```


# Analysis framework

Analyses were conducted in specific directories, organised in a tree as shown below:

```{bash, eval = TRUE}
cat tree.txt
```

Note that the menu on the left follows almost the same structure.

# RMD knitting

R Markdown (RMD) notebooks were developed within R Studio. To ensure the reproducibility of all the analyses from the raw count matrices to the final figures and results, the RMD files were compiled using a Singularity container, to generate the HTML pages presented here. Excepting the notebooks focusing on individual dataset, all notebooks are parameter-free, so that they can be compiled using the same command. The following template was used:

```{bash}
file="FILE_NAME"
singularity exec \
--bind /home \
/path/to/singularity/image/singularity.simg \
Rscript -e "rmarkdown::render(input = '${file}.Rmd', output_file = '${file}.html')"
```

Since R libraries are not located at the same location within the Singularity container or on the local machine, we add the `.libPaths()` R command at the beginning of each notebook. In the compiled HTML file, we thus can easily check if the notebook was compiled through the container or not.

# Index page

The static website was built using OrganizeHTML (DOI: [10.5281/zenodo.14505634](https://doi.org/10.5281/zenodo.14505634))


<center>--end--</center>