---
title: "Figures for the paper"
subtitle: "Version 23"
author: "Audrey"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <<- Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE,
                      
                      # save figures in PDF in a separate folder
                      # https://stackoverflow.com/questions/27992239/knitr-include-figures-in-report-and-output-figures-to-separate-files
                      dev = c('png', 'pdf'), # tiff or pdf alone renders bad in html
                      # dpi = 300,
                      fig.path = "figures_detail/",
                      pdf.options(encoding = "ISOLatin9.enc"))
```


In this file, we make the scRNAseq figures for the paper.

```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
library(ComplexHeatmap)

.libPaths()
```

# Settings

Analyses are stored here :

```{r locations}
analysis_dir = "."

mouse_dir = paste0(analysis_dir, "/../1_mice/")
patient_dir = paste0(analysis_dir, "/../2_patients/")
graft_dir = paste0(analysis_dir, "/../3_graft/")
```

These are all the samples analysed :

```{r figZ_1, class.source = 'fold-hide', fig.width = 10, fig.height = 15}
# Mice
sample_info = readRDS(paste0(mouse_dir, "/1_metadata/mice_sample_info.rds"))
plot_list = aquarius::fig_plot_gb(sample_info, title = "Nf1 fl/fl mice datasets")

# Patients
sample_info = readRDS(paste0(patient_dir, "/1_metadata/patients_sample_info.rds"))
plot_list = c(plot_list, aquarius::fig_plot_gb(sample_info, title = "NF1 patients datasets"))

# Graft experiment
sample_info = readRDS(paste0(graft_dir, "/1_metadata/graft_sample_info.rds"))
plot_list = c(plot_list, aquarius::fig_plot_gb(sample_info, title = "Graft experiment"))

# patchwork
patchwork::wrap_plots(plot_list) +
  patchwork::plot_layout(nrow = length(plot_list), heights = c(0.1,4, 0.1, 4, 0.1, 0.5)) &
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15))
```

These are the custom colors for cell populations :

```{r figZ_2, fig.width = 12, fig.height = 1, class.source = "fold-hide"}
color_markers = readRDS(paste0(mouse_dir, "/1_metadata/mice_color_markers.rds"))

color_markers[["mural cells"]] = "gray60"
names(color_markers)[which(names(color_markers) == "tumor cells")] = "TOM+ cells"
names(color_markers)[which(names(color_markers) == "macrophages")] = "M\U03A6/Mo"

data.frame(cell_type = names(color_markers),
           color = unlist(color_markers)) %>%
  ggplot2::ggplot(., aes(x = cell_type, y = 0, fill = cell_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers), breaks = names(color_markers)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank())
```

```{r clean_settings, echo = FALSE}
rm(sample_info, plot_list)
```


# SH figure

We open the table :

```{r load_sh_tab}
sh_tab = openxlsx::read.xlsx(paste0(analysis_dir, "/data/Validation target SH.xlsx"), sheet = 3)
sh_tab$id = paste(sh_tab$sh_id, sh_tab$jour, sep = "_DAY_")
sh_tab = sh_tab %>%
  dplyr::filter(!(id %in% c("Sox9 clone 22-5_DAY_14", "Sox9 clone 22-6_DAY_14")))

head(sh_tab)
```

For each condition (`id`), we make a paired t-test.

```{r sh_ttest}
my_ttest = lapply(sort(unique(sh_tab$id)), FUN = function(one_id) {
  sub_tab = sh_tab %>%
    dplyr::filter(id == one_id)
  the_test = stats::t.test(x = sub_tab$mass_ctrl,
                           y = sub_tab$mass_sh,
                           paired = TRUE)
  p_value = the_test$p.value
  return(p_value)
})
my_ttest = data.frame(id = sort(unique(sh_tab$id)),
                      pvalue = unlist(my_ttest))
my_ttest$x = c(1:nrow(my_ttest))
head(my_ttest)
```

We make a long dataframe for ggplot :

```{r long_sh_tab}
long_sh_tab = rbind(sh_tab[, c("mouse_id", "sh_id", "sh_gene", "mass_ctrl", "jour", "mono_poly", "id")] %>%
                      `colnames<-`(c("mouse_id", "sh_id", "sh_gene", "mass", "jour", "mono_poly", "id")),
                    sh_tab[, c("mouse_id", "sh_id", "sh_gene", "mass_sh", "jour", "mono_poly", "id")] %>%
                      `colnames<-`(c("mouse_id", "sh_id", "sh_gene", "mass", "jour", "mono_poly", "id")))
long_sh_tab$sh_ctrl = c(rep("ctrl", nrow(long_sh_tab)/2),
                        rep("sh", nrow(long_sh_tab)/2))

head(long_sh_tab)
```

We make the figure for a subset of experiments :

```{r sh_figure_2, fig.width = 8, fig.height = 5, class.source = "fold-hide"}
clones_oi = c("Cxcl12_DAY_25",
              "CRABP1  clone 3_DAY_14",
              "MIF clone 1_DAY_14",
              "Pdgfr-alpha clone 1_DAY_14",
              "Pdgfr-béta clone 7_DAY_14",
              "Sox9 Clone 13_DAY_14")

# Subset clone of interest
long_sh_tab = long_sh_tab %>%
  dplyr::filter(id %in% clones_oi) %>%
  base::droplevels()

my_ttest = my_ttest %>%
  dplyr::filter(id %in% clones_oi) %>%
  dplyr::mutate(x = c(1:nrow(.))) %>%
  base::droplevels()

# Plot
ggplot2::ggplot(long_sh_tab, aes(x = sh_gene, y = mass, color = sh_ctrl)) +
  # Box plot and point inside
  ggplot2::geom_boxplot(alpha = 0.1, width = 0.75) +
  ggplot2::geom_point(position = position_dodge(width = 0.75)) +
  # Horizontal bar for paired t-test
  ggplot2::geom_segment(data = my_ttest, inherit.aes = FALSE,
                        aes(x = x - 0.4, xend = x + 0.4,
                            y = 1.15*max(long_sh_tab$mass),
                            yend = 1.15*max(long_sh_tab$mass))) +
  # p-value
  ggplot2::geom_label(data = my_ttest, inherit.aes = FALSE,
                      aes(x = x,
                          y = 1.10*max(long_sh_tab$mass),
                          label = format(round(pvalue, 2), nsmall = 2)),
                      fill = NA, label.size = 0) +
  # Significance
  ggplot2::geom_label(data = my_ttest, inherit.aes = FALSE,
                      aes(x = x,
                          y = case_when(pvalue > 0.05 ~ 1.21*max(long_sh_tab$mass),
                                        pvalue < 0.05 ~ 1.17*max(long_sh_tab$mass)),
                          size = case_when(pvalue > 0.05 ~ 1,
                                           pvalue < 0.05 ~ 2),
                          label = case_when(pvalue > 0.05 ~ "n.s.",
                                            pvalue > 0.01 & pvalue < 0.05 ~ "*",
                                            pvalue > 0.001 & pvalue < 0.01 ~ "**",
                                            pvalue < 0.001 ~ "***")),
                      fill = NA, label.size = 0, size = 5) +
  ggplot2::scale_radius(range = c(4, 7), guide = "none") +
  # Theme
  # ggplot2::scale_x_discrete(labels = long_sh_tab$id_simple) +
  ggplot2::scale_y_continuous(breaks = seq(0, 1.1*max(long_sh_tab$mass), by = 0.25),
                              labels = format(seq(0, 1.1*max(long_sh_tab$mass), by = 0.25), nsmall = 2),
                              limits = c(0, 1.23*max(long_sh_tab$mass)),
                              expand = c(0,0)) +
  ggplot2::labs(y = "Tumor mass (g)") +
  ggplot2::guides(color = guide_legend(title = "Tumor")) +
  ggplot2::theme_classic() +
  ggplot2::theme(plot.title = element_text(hjust = 0.5),
                 text = element_text(size = 20),
                 axis.line.x = element_line(colour = "lightgray"),
                 axis.title.x = element_blank())
```

```{r clean_sh, echo = FALSE}
rm(sh_tab, long_sh_tab, my_ttest, plot_list, clones_oi)
```

# SI figure

We open the table :

```{r si_read_table}
si_tab = read.csv(paste0(analysis_dir, "/data/SOX9-LSS06-Spike-Results-manuscript.csv")) %>%
  dplyr::mutate(Treatment = ifelse(Treatment == "siCTL",
                                   yes = "si-ctrl",
                                   no = "si-Sox9"))

head(si_tab)
```

We make the figure :

```{r si_figure, fig.width = 4, fig.height = 9}
ggplot2::ggplot(si_tab, aes(x = Treatment, y = X..of.siCTL.Nuc)) +
  # Box plot and point inside
  ggplot2::geom_boxplot(alpha = 0.1, width = 0.75, col = "gray50") +
  ggplot2::geom_jitter(width = 0.25, size = 2) +
  # Theme
  ggplot2::scale_y_continuous(breaks = seq(0, 1.1*max(si_tab$X..of.siCTL.Nuc), by = 25),
                              labels = format(seq(0, 1.1*max(si_tab$X..of.siCTL.Nuc), by = 25), nsmall = 2),
                              limits = c(0, 125),
                              expand = c(0,0)) +
  ggplot2::labs(y = "% of si-ctrl intensity",
                x = "siRNA") +
  ggplot2::guides(color = guide_legend(title = "Tumor")) +
  ggplot2::theme_classic() +
  ggplot2::theme(plot.title = element_text(hjust = 0.5),
                 text = element_text(size = 25),
                 axis.line.x = element_line(colour = "lightgray"),
                 panel.grid.major = element_line(colour = "lightgray", size = 0.25))
```

We clean the environment :

```{r clean_si}
rm(si_tab)
```

# Whole Exome Sequencing

In this section, we make the figure relative to the WES.

## CNV - primary

We make the figure relative to CNV in primary tumors.

### Load data

We load the CNV table :

```{r wes_load_cnv}
cnv_data = openxlsx::read.xlsx("./data/WES2023_Aggregated_annotated_Primaries.xlsx")

dim(cnv_data) ; colnames(cnv_data)
```

We arrange columns :

```{r wes_arrange_columns}
cnv_data = cnv_data %>%
  dplyr::select(Line, Chr, Start, End,
                `27187_PrimaryTumor_vs_27187_Spleen`,
                `83935_PrimaryTumor_vs_83935_Spleen`,
                CF153_PrimaryTumor_vs_CF153_Spleen,
                M23349_Tumor_vs_Spleen,
                M23465_Tumor_vs_Spleen,
                M46643_Tumor1_vs_M46643_Spleen,
                M46643_Tumor2_vs_M46643_Spleen,
                M46643_Tumor3_vs_M46643_Spleen,
                M62892_Tumor_vs_Spleen,
                M62908_Tumor_vs_M62908_Spleen,
                M63717_Tumor_vs_Spleen,
                M76339_Tumor_vs_M76339_Spleen,
                M83937_TumorP0_vs_M83937_Spleen,
                FREQ_LOSS, FREQ_GAIN, FREQ_TOTAL, Cyto.Start, Symbols) %>%
  dplyr::filter(FREQ_GAIN != 0 | FREQ_LOSS != 0) %>%
  dplyr::mutate(region = paste0(Chr, Cyto.Start)) %>%
  dplyr::mutate(region = stringr::str_replace_all(region, pattern = "\\.", replacement = "_"))

dim(cnv_data) ; colnames(cnv_data)
table(cnv_data$region) %>% head()
table(cnv_data$region) %>% dim()
```

We extract sample names :

```{r wes_sample_names}
sample_col_ids = c(5:17)
sample_names = colnames(cnv_data[sample_col_ids])
sample_names
```

We remove `NA`-containing rows, because it is better if the plot shows values for all samples :

```{r wes_remove_na}
cnv_data = na.omit(cnv_data)
dim(cnv_data)
```

### Annotation

We prepare everything to make the oncoplot :

```{r wes_prep_annotation, class.source = "fold-hide"}
binarize = function(value) {
  value = dplyr::case_when(is.na(value) ~ "na",
                           # loss of two copies
                           value <= -0.12 ~ "loss",
                           # no change (background)
                           value > -0.13 & value < 0.25 ~ "",
                           # gain
                           value >= 0.25 ~ "gain")
  return(value)
}

vec_colors_cnv = c("loss" = "#2166AC",
                   "background" = "#F7F7F7",
                   "gain" = "#B2182B",
                   na = "gray60")

alter_fun_cnv = list(
  background = alter_graphic("rect", fill = vec_colors_cnv[["background"]],
                             width = 0.9, height = 0.9),
  "loss" = alter_graphic("rect", fill = vec_colors_cnv[["loss"]],
                         width = 0.9, height = 0.9),
  "gain" = alter_graphic("rect", fill = vec_colors_cnv[["gain"]],
                         width = 0.9, height = 0.9),
  na = alter_graphic("rect", fill = vec_colors_cnv[["na"]],
                     width = 0.9, height = 0.9)
)

heatmap_legend_param = list(title = "Copy Number Variations",
                            at = c("loss", "gain", "na"), 
                            labels = c("loss", "gain", "na"))
```

### Binarize and subset

We binarize the CNV continuous data into gain, loss or background :

```{r wes_binarize}
cnv_data[, sample_names] = vapply(as.matrix(cnv_data[, sample_names]),
                                  FUN = binarize,
                                  FUN.VALUE = character(1))
```

We keep information when they are alterations in at least 4 samples :

```{r wes_subset_4_samples}
cnv_data$nb_samples = rowSums(cnv_data[, sample_names] != "", na.rm = TRUE)

cnv_data = cnv_data %>%
  dplyr::filter(nb_samples > 4)

dim(cnv_data)
```

### Regions of interest

We identify regions as the combined information of **chrosomome + cytoband + motif**, motif being as `___gain__loss___loss__loss__loss`.

```{r wes_motif}
cnv_data$motif = apply(cnv_data[, sample_names], 1, FUN = paste, collapse = "_")
head(cnv_data$motif)

cnv_data = cnv_data %>%
  dplyr::group_by(paste(region, motif))
```

We want to print the gene names of interest along with the chromosome, in left annotation of the oncoplot :

```{r wes_genes_oi, class.source = "fold-hide"}
# all_genes = (cnv_data$Symbols %>%
#                paste0(., collapse = ",") %>%
#                stringr::str_split(.,
#                                   pattern = ","))[[1]] %>%
#   unique() %>% sort()
# 
# not_of_interest = Rik
# 
# all_genes = grep(all_genes, pattern = "Rik$|^Gm[0-9]*$|^Mir[0-9][0-9]*",
#                  value = TRUE, invert = TRUE)
# 
# all_genes = grep(all_genes, pattern = "-ps$",
#                  value = TRUE, invert = TRUE)
# 
# all_genes = grep(all_genes, pattern = "^AA|^AU|^BC",
#                  value = TRUE, invert = TRUE)
# 
# all_genes = grep(all_genes, pattern = "^Snor",
#                  value = TRUE, invert = TRUE)
# 
# all_genes = grep(all_genes, pattern = "^Olfr[0-9][0-9]*",
#                  value = TRUE, invert = TRUE)
# 
# all_genes = grep(all_genes, pattern = "^Slc[0-9][0-9]*",
#                  value = TRUE, invert = TRUE)
# 
# all_genes = grep(all_genes, pattern = "^Rp[l|s][0-9]*$",
#                  value = TRUE, invert = TRUE)
# 
# 
# max_print = getOption("max.print")
# options(max.print = length(all_genes))
# print(all_genes, max.print = length(all_genes))
# options(max.print = max_print)

genes_oi = c("Acta2", "Btg1", "Dcn", "Lum",
             "Cdk1", "Cdk4", "Cdkn2a", "Igf1",
             "Lama2", "Lama4", "Mdm2", "Mif", "Mmp11", "Mmp19", "Nfkb2",
             "Pten", "Ptprk", "Timp3")
```

We arrange the grouped dataframe :

```{r wes_arrange_grouped, class.source = "fold-hide"}
cnv_group_fun = function(sub_dataframe) {
  grouped_row = list()
  
  ## Chr
  grouped_row[["Chr"]] = sub_dataframe[["Chr"]] %>%
    unique() %>%
    sort() %>%
    paste(., collapse = "_")
  
  ## Start
  grouped_row[["Start"]] = sub_dataframe[["Start"]] %>%
    min()
  
  ## End
  grouped_row[["End"]] = sub_dataframe[["End"]] %>%
    max()
  
  ## Width
  grouped_row[["Width"]] = grouped_row[["End"]] - grouped_row[["Start"]]
  
  ## Cyto
  grouped_row[["Cyto.Start"]] = sub_dataframe[["Cyto.Start"]] %>%
    unique() %>%
    sort() %>%
    paste(., collapse = "_")
  
  ## Region
  grouped_row[["Region"]] = paste0(grouped_row[["Chr"]],
                                   grouped_row[["Cyto.Start"]])#,
  #"_",
  #grouped_row[["Width"]]) 
  
  ## Symbols
  all_symbols = (sub_dataframe[["Symbols"]] %>%
                   paste(., collapse = ",") %>%
                   stringr::str_split(., pattern = ","))[[1]] %>%
    unique() %>% sort()
  
  grouped_row[["Symbols"]] = all_symbols[all_symbols %in% genes_oi] %>%
    paste(., collapse = ",")
  
  ## Add in symbols
  if (grouped_row[["Symbols"]] != "") {
    grouped_row[["Region"]] = paste0(grouped_row[["Region"]], " (",
                                     grouped_row[["Symbols"]], ")")
  }
  
  ## Unlist
  grouped_row = unlist(grouped_row)
  
  ## Samples
  sample_summary = sub_dataframe[sample_col_ids] %>%
    apply(., 2, FUN = unique) %>%
    unlist()
  
  ## Output
  grouped_row = c(grouped_row, sample_summary)
  
  return(grouped_row)
}

new_column_names = c("Chr", "Start", "End", "Width", "Cyto.Start", "Region", "Symbols",
                     colnames(cnv_data)[sample_col_ids])

cnv_data = cnv_data %>%
  dplyr::group_map(~ cnv_group_fun(.)) %>%
  do.call(rbind.data.frame, .) %>%
  `colnames<-`(new_column_names)

cnv_data[, c("Start", "End", "Width")] = apply(cnv_data[, c("Start", "End", "Width")], 2,
                                               FUN = function(x) {as.numeric(as.character(x))})

dim(cnv_data)
```

Still to many information. We keep region when they are longer than 500 nucleotides, and for each cytoband, keep the most affecting section.

```{r wes_subset_cnv}
cnv_data = cnv_data %>%
  dplyr::filter(Width > 500)

dim(cnv_data)

# Keep the most affecting section of the cytoband
cnv_data$prob_affected = apply(cnv_data, MARGIN = 1, FUN = function(one_row) {
  return(sum(one_row != ""))
})

cnv_data = cnv_data %>%
  dplyr::group_by(Chr, Cyto.Start) %>%
  dplyr::filter(prob_affected == max(prob_affected)) %>%
  dplyr::filter(Width == max(Width)) %>% # ensure only one
  as.data.frame()

dim(cnv_data)
```

### Oncoplot

We reshape the dataframe into a matrix compatible with the oncoplot function :

```{r wes_prep_matrix}
cnv_matrix = as.matrix(cnv_data[, sample_names])
rownames(cnv_matrix) = cnv_data$Region
```

We load sample annotation and re-order the matrix :

```{r wes_sample_annot}
# Sample annotation
sample_annot = openxlsx::read.xlsx("./data/Table_Genetic.xlsx", sheet = 3)

# Add id corresponding to scRNA-Seq
sample_annot$sc_ID = dplyr::case_when(
  sample_annot$ID == "#46643_T1" ~ "flmn_MPNST_1", # 2022_25 -- check
  sample_annot$ID == "#46643_T2" ~ "flmn_MPNST_2", # 2022_26 -- check
  sample_annot$ID == "#46643_T3" ~ "flmn_MPNST_0", # 2022_27 -- check
  sample_annot$ID == "#83935" ~ "flmn_MPNST_E",
  sample_annot$ID == "#27187" ~ "flmn_MPNST_F",
  sample_annot$ID == "#83937" ~ "flmn_MPNST_G",
  sample_annot$ID == "#CF153" ~ "flfl_MPNST_3", # 2019_G
  sample_annot$ID == "#62908" ~ "flfl_MPNST_2", # 2020_19
  sample_annot$ID == "#76339" ~ "flfl_MPNST_4", # 2021_33
  sample_annot$ID == "#63717" ~ "flfl_MPNST_A",
  sample_annot$ID == "#23249" ~ "flfl_MPNST_B",
  sample_annot$ID == "#23465" ~ "flfl_MPNST_C",
  sample_annot$ID == "#62982"  ~ "flfl_MPNST_D",
  TRUE ~ "PROBLEM")

# Reshape
sample_annot = sample_annot %>%
  `rownames<-`(c(.$sc_ID)) %>%
  dplyr::arrange(Genotype) %>%
  dplyr::select(Genotype, `ScRNA-Seq`, Exome_ID) %>%
  `colnames<-`(c("Genotype", "scRNA-Seq", "Exome_ID"))

## Re-order
cnv_matrix = cnv_matrix[, sample_annot$Exome_ID]
colnames(cnv_matrix) = rownames(sample_annot)
```

We define custom colors for the figure :

```{r wes_colors}
list_colors = list()
list_colors[["Genotype"]] = setNames(nm = c("fl/fl", "fl/-"),
                                     c("orange", "darkgreen"))
list_colors[["scRNA-Seq"]] = setNames(nm = c("Yes", "No"),
                                      c("black", "white"))
```

We define the annotation :

```{r wes_annot}
fontsize = 10

ha_top = HeatmapAnnotation(
  df = sample_annot[, c("Genotype", "scRNA-Seq")],
  # cbar = anno_oncoprint_barplot(height = unit(2, "cm")),
  col = list_colors,
  annotation_name_side = "left",
  annotation_name_gp = grid::gpar(fontsize = fontsize),
  na_col = "#F7F7F7")
```

and draw the oncoplot :

```{r wes_ht_cnv}
ht_cnv = ComplexHeatmap::oncoPrint(cnv_matrix,
                                   # Colors
                                   alter_fun = alter_fun_cnv,
                                   alter_fun_is_vectorized = TRUE,
                                   # Barplot and annotation
                                   col = vec_colors_cnv,
                                   top_annotation = ha_top,
                                   right_annotation = NULL,
                                   # Legend
                                   heatmap_legend_param = heatmap_legend_param,
                                   # Sample names
                                   show_column_names = TRUE,
                                   # column_order = colnames(cnv_matrix),
                                   column_names_gp = grid::gpar(fontsize = fontsize),
                                   # Region names
                                   row_names_side = "left",
                                   row_names_gp = grid::gpar(fontsize = fontsize),
                                   # row_order = rownames(cnv_matrix),
                                   # % side
                                   pct_side = "right",
                                   pct_digits = 0)
```

```{r figWES_CNV_primary, fig.width = 12, fig.height = 8}
ComplexHeatmap::draw(ht_cnv,
                     merge_legend = TRUE,
                     padding = unit(c(10,10,10,10), "mm"),
                     heatmap_legend_side = "left",
                     annotation_legend_side = "left")
```

## SNV - primary

We make the figure relative to SNV in primary tumors.

### Load data

We load the SNV table :

```{r wes_load_snv_primary}
snv_data = openxlsx::read.xlsx("./data/Figure_SNVs_MPNST-Paper.xlsx",
                               sheet = 1, startRow = 4)
snv_data = snv_data[, c(3:16)]
dim(snv_data)
```

We extract the information regarding number of SNV per sample :

```{r wes_nb_snv_data}
nb_snv_data = snv_data[c(8:10), ]
rownames(nb_snv_data) = nb_snv_data$X3
nb_snv_data$X3 = NULL
nb_snv_data = nb_snv_data %>%
  mutate(across(everything(), as.numeric)) %>%
  t()
nb_snv_data
```

We change the columns in `nb_snv_data` table :

```{r wes_nb_snv_data_change}
nb_snv_data = nb_snv_data %>%
  log10() %>%
  as.data.frame %>%
  `colnames<-`(c("total", "transcripts", "AA")) %>%
  dplyr::mutate(transcript_not_aa = transcripts - AA) %>%
  dplyr::mutate(total_not_transcripts = total - transcripts) %>%
  dplyr::select(AA, transcript_not_aa, total_not_transcripts)
```

We split the snv table to keep only information regarding genes :

```{r wes_keep_genes_info}
snv_data = snv_data[c(25:nrow(snv_data)), ]
rownames(snv_data) = snv_data$X3
snv_data$X3 = NULL
snv_data[is.na(snv_data) == TRUE] = ""
snv_data[snv_data == "*"] = "snv;*"
snv_data = as.matrix(snv_data)

snv_data[c(1:5), c(1:3)]
```

We re-order the dataframe in the same order as the CNV table 

```{r wes_reorder_snv}
snv_data = snv_data[, sample_annot$Exome_ID]
colnames(snv_data) = rownames(sample_annot)

nb_snv_data = nb_snv_data[sample_annot$Exome_ID,]
rownames(nb_snv_data) = rownames(sample_annot)
```

### Annotation

We define the oncoplot annotation :

```{r wes_snv_annot}
# Barplot annotation
breaks = c(1, 10, 100, 1000, 10000)
ha_top = ComplexHeatmap::HeatmapAnnotation(
  "Nb SNV" = anno_barplot(as.matrix(nb_snv_data),
                          gp = gpar(fill = c("black", "gray50", "lightgray")),
                          axis_param = list(at = log10(breaks),
                                            labels = format(breaks, scientific = FALSE)),
                          border = FALSE),
  height = unit(2, "cm"),
  annotation_name_side = "left")

lgd_list = list(
  ComplexHeatmap::Legend(
    labels = c("Total SNV", "SNV within transcripts", "SNV within transcripts, AA-changed"),
    title = "Nb of SNV",
    legend_gp = gpar(fill = c("lightgray", "gray50", "black"))))

# Oncoprint colors
vec_colors_snv = c("snv" = "mediumseagreen",
                   "background" = "#F7F7F7")

alter_fun_snv = list(
  "background" = alter_graphic("rect", fill = vec_colors_snv[["background"]],
                               width = 0.9, height = 0.9),
  "snv" = alter_graphic("rect", fill = vec_colors_snv[["snv"]],
                        width = 0.9, height = 0.9),
  "*" = function(x, y, w, h) {
    # grid.segments(x - w*0.4, y - h*0.4, x + w*0.4, y + h*0.4, gp = gpar(lwd = 2))
    # grid.segments(x + w*0.4, y - h*0.4, x - w*0.4, y + h*0.4, gp = gpar(lwd = 2))
    grid.points(x, y, pch = 16)
  })

heatmap_legend_param = list(title = "Single Nucleotide Variations",
                            at = c("snv", "*"), 
                            labels = c("SNV", "SNV AA-altered"))
```

and draw the oncoplot :

```{r wes_ht_snv}
ht_snv = ComplexHeatmap::oncoPrint(snv_data,
                                   # Colors
                                   alter_fun = alter_fun_snv,
                                   alter_fun_is_vectorized = TRUE,
                                   # Barplot and annotation
                                   col = vec_colors_snv,
                                   top_annotation = ha_top,
                                   # top_annotation = NULL,
                                   right_annotation = NULL,
                                   # Legend
                                   heatmap_legend_param = heatmap_legend_param,
                                   # Sample names
                                   show_column_names = TRUE,
                                   # column_order = colnames(snv_data),
                                   column_names_gp = grid::gpar(fontsize = fontsize),
                                   # Region names
                                   row_names_side = "left",
                                   row_names_gp = grid::gpar(fontsize = fontsize),
                                   row_order = rownames(snv_data),
                                   # % side
                                   pct_side = "right",
                                   pct_digits = 0)
```

```{r figWES_SNV_primary, fig.width = 12, fig.height = 10}
ComplexHeatmap::draw(ht_snv,
                     merge_legend = TRUE,
                     padding = unit(c(10,10,10,10), "mm"),
                     annotation_legend_list = lgd_list,
                     heatmap_legend_side = "left",
                     annotation_legend_side = "left")
```
## Primary

To force same columns order but automatic ordering, and same columns size, we merge both figures :

```{r figWES_primary, fig.width = 11, fig.height = 15}
ht_list = ht_cnv %v% ht_snv

ComplexHeatmap::draw(ht_list,
                     padding = unit(c(10,10,10,10), "mm"),
                     ht_gap = unit(1.5, "cm"),
                     merge_legend = TRUE,
                     annotation_legend_list = lgd_list,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```

We clean the environment :

```{r wes_clean_1}
rm(sample_col_ids, sample_names, new_column_names, genes_oi,
   cnv_group_fun, vec_colors_cnv, alter_fun_cnv, binarize,
   cnv_data, cnv_matrix, list_colors, heatmap_legend_param,
   snv_data, nb_snv_data, ha_top, vec_colors_snv, alter_fun_snv, 
   breaks, sample_annot,  ht_cnv, ht_snv, ht_list)
```

## SNV - transplanted

We make the figure relative to SNV in transplanted tumors.

### Load data

We load the SNV table :

```{r twes_load_snv_trans}
snv_data = openxlsx::read.xlsx("./data/A_Tableau_SNV_Transplanted.xlsx",
                               sheet = 1, startRow = 4)
snv_data = snv_data[c(8:nrow(snv_data)), c(3,8:11)]
dim(snv_data)
```

We extract the information regarding number of SNV per sample :

```{r twes_nb_snv_data}
nb_snv_data = snv_data[c(1:3), ]
rownames(nb_snv_data) = nb_snv_data$X3
nb_snv_data$X3 = NULL
nb_snv_data = nb_snv_data %>%
  mutate(across(everything(), as.numeric)) %>%
  t()
nb_snv_data
```

We split the snv table to keep only information regarding genes :

```{r twes_keep_genes_info}
snv_data = snv_data[c(4:nrow(snv_data)), ]
snv_data = snv_data[!duplicated(snv_data$X3), ]
rownames(snv_data) = snv_data$X3
snv_data$X3 = NULL
snv_data[is.na(snv_data) == TRUE] = ""
snv_data[snv_data == "*"] = "snv;*"
snv_data = as.matrix(snv_data)

head(snv_data)
```

We change columns in the `nb_snv_data` table :

```{r twes_nb_snv_data_change}
nb_snv_data = nb_snv_data %>%
  # log10() %>%
  as.data.frame %>%
  `colnames<-`(c("total", "transcripts", "AA")) %>%
  dplyr::mutate(transcript_not_aa = transcripts - AA) %>%
  dplyr::mutate(total_not_transcripts = total - transcripts) %>%
  dplyr::select(AA, transcript_not_aa, total_not_transcripts)
```

We add the loss of Cdkn2a, that was detected in CNV analysis :

```{r add_cdkn2a}
snv_data = rbind(rep("cnv", ncol(snv_data)),
                 snv_data)
rownames(snv_data)[1] = "Cdkn2a"

head(snv_data)
```


### Annotation

We define the oncoplot annotation :

```{r twes_snv_annot}
# Barplot annotation
breaks = c(1, 10, 100, 1000, 10000, 100000)
ha_top = ComplexHeatmap::HeatmapAnnotation(
  "Nb SNV\n" = anno_barplot(as.matrix(nb_snv_data),
                          gp = gpar(fill = c("black", "gray50", "lightgray")),
                          # axis_param = list(at = log10(breaks),
                          #                   labels = format(breaks, scientific = FALSE)),
                          border = FALSE),
  height = unit(2, "cm"),
  annotation_name_side = "left")

# Oncoprint colors
vec_colors_snv = c("cnv" = "#2166AC",
                   "snv" = "mediumseagreen",
                   "background" = "#F7F7F7")

lgd_list = list(
  ComplexHeatmap::Legend(
    labels = c("Total SNV", "SNV within transcripts", "SNV within transcripts, AA-changed"),
    title = "Nb of SNV",
    legend_gp = gpar(fill = c("lightgray", "gray50", "black"))),
  ComplexHeatmap::Legend(
    labels = c("loss"),
    title = "Copy Number Variations",
    legend_gp = gpar(fill = vec_colors_snv[["cnv"]])))

alter_fun_snv = list(
  "background" = alter_graphic("rect", fill = vec_colors_snv[["background"]],
                               width = 0.9, height = 0.9),
  "snv" = alter_graphic("rect", fill = vec_colors_snv[["snv"]],
                        width = 0.9, height = 0.9),
  "cnv" = alter_graphic("rect", fill = vec_colors_snv[["cnv"]],
                        width = 0.9, height = 0.9),
  "*" = function(x, y, w, h) {
    # grid.segments(x - w*0.4, y - h*0.4, x + w*0.4, y + h*0.4, gp = gpar(lwd = 2))
    # grid.segments(x + w*0.4, y - h*0.4, x - w*0.4, y + h*0.4, gp = gpar(lwd = 2))
    grid.points(x, y, pch = 16)
  })

heatmap_legend_param = list(title = "Single Nucleotide Variations",
                            at = c("snv", "*"), 
                            labels = c("SNV", "SNV AA-altered"))
```

and draw the oncoplot :

```{r twes_ht_snv}
ht_snv = ComplexHeatmap::oncoPrint(snv_data,
                                   # Colors
                                   alter_fun = alter_fun_snv,
                                   alter_fun_is_vectorized = TRUE,
                                   # Barplot and annotation
                                   col = vec_colors_snv,
                                   top_annotation = ha_top,
                                   # top_annotation = NULL,
                                   right_annotation = NULL,
                                   # Legend
                                   heatmap_legend_param = heatmap_legend_param,
                                   # Sample names
                                   show_column_names = TRUE,
                                   # column_order = colnames(snv_data),
                                   column_names_gp = grid::gpar(fontsize = fontsize),
                                   # Region names
                                   row_names_side = "left",
                                   row_names_gp = grid::gpar(fontsize = fontsize),
                                   row_order = rownames(snv_data),
                                   # % side
                                   pct_side = "right",
                                   pct_digits = 0)
```

```{r figWES_SNV_transplanted, fig.width = 7.5, fig.height = 9}
ComplexHeatmap::draw(ht_snv,
                     padding = unit(c(10,10,10,10), "mm"),
                     merge_legend = TRUE,
                     annotation_legend_list = lgd_list,
                     heatmap_legend_side = "right",
                     annotation_legend_side = "right")
```

# Mouse

In this section, we make the figures related to spontaneous MPNST development in mice. There are three sections, related to specific analyses :

* MPNST only in both fl/fl and fl/- mouse models
* pNF to MPNST atlas in fl/fl mouse model
* pNF to MPNST tumor cells trajectory in fl/fl mouse model

Here are common settings :

```{r sample_info_mouse}
sample_info = readRDS(paste0(mouse_dir, "/1_metadata/mice_sample_info.rds")) # sample info
dotplot_markers = readRDS(paste0(mouse_dir, "/1_metadata/mice_dotplot_markers.rds")) # dotplot markers
names(dotplot_markers) = names(color_markers)
```

For the dotplot, we merge the three DC populations :

```{r merge_dotplot_markers}
dotplot_markers[["DC"]] = dotplot_markers[c("pDC", "mDC", "cDC")] %>%
  unlist() %>% unname()
dotplot_markers
```


We are going to make barplot with cell type and associated cell family. This are settings for cell families representation :

```{r custom_order_cell_type}
custom_order_cell_type = data.frame(
  cell_type = c("DC", "DC", "DC", "mast cells", "neutrophils",
                "B cells", "NK cells", "T cells", "M\U03A6/Mo", "skeletal muscle cells",
                "mural cells", "endothelial cells", "fibroblasts", "TOM+ cells"),
  cell_type_simple = c(rep("other immune cells", 7), "T cells", "M\U03A6/Mo",
                       rep("other stromal cells", 3), "fibroblasts", "TOM+ cells"),
  cell_type_old = c("pDC", "mDC", "cDC", "mast cells", "neutrophils",
                    "B cells", "NK cells", "T cells", "macrophages", "skeletal muscle cells",
                    "mural cells", "endothelial cells", "fibroblasts", "tumor cells"),
  cell_family = c(rep("immune cells", 9),
                  rep("stromal cells", 4),
                  rep("tumor cells", 1)),
  stringsAsFactors = FALSE)
rownames(custom_order_cell_type) = custom_order_cell_type$cell_type_old

family_color = c("immune cells" = "gray90",
                 "stromal cells" = "gray50",
                 "tumor cells" = "gray0")
```

We define colors for simple cell type :

```{r color_markers_simple}
color_markers_simple = setNames(
  nm = c("other immune cells", "T cells", "M\U03A6/Mo",
         "other stromal cells", "fibroblasts", "TOM+ cells"),
  c("pink", unlist(color_markers[c("T cells", "M\U03A6/Mo")]),
    "gray70", unlist(color_markers[c("fibroblasts", "TOM+ cells")])))
color_markers_simple
```


We change fl/- MPNST tumors colors to purple shades :

```{r flmn_colors}
bool_flmn = sample_info$genotype == "flmn"
sample_info[bool_flmn, "color"] = colorRampPalette(c("violetred1", "violetred3"))(sum(bool_flmn))
rm(bool_flmn)
```

## MPNST atlas

### Dataset preparation

We load the MPNST dataset :

```{r load_mpnst_dataset}
sobj = readRDS(paste0(mouse_dir, "/3_combined_mpnst/mpnst_sobj.rds"))
reduction_tsne = "harmony_38_tsne"

sobj
```


This is the number of cells in each dataset :

```{r fig_supp_mpnst_1, fig.width = 5, fig.height = 4}
table(sobj$sample_identifier) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("sample_id", "nb_cells")) %>%
  dplyr::left_join(.,
                   y = sample_info[, c("sample_identifier", "color")],
                   by = c("sample_id" = "sample_identifier")) %>%
  aquarius::fig_plot_gb(column_names = c("Sample ID", "Nb cells", "Color")) %>%
  patchwork::wrap_plots(ncol = 1) +
  patchwork::plot_layout(heights = c(0,5))
```


There is a problem with skeletal muscle cells annotation. We change annotation :

```{r main_cell_type_by_cluster_mpnst}
main_cell_type_by_cluster = table(sobj$seurat_clusters,
                                  sobj$cell_type) %>%
  prop.table(., margin = 1) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("clusters", "cell_type", "prop")) %>%
  dplyr::group_by(clusters) %>%
  dplyr::filter(prop == max(prop)) %>%
  dplyr::arrange(clusters) %>%
  as.data.frame()
main_cell_type_by_cluster = setNames(
  nm = main_cell_type_by_cluster$clusters,
  as.character(main_cell_type_by_cluster$cell_type))

sobj$cell_type = ifelse(sobj$cell_type == "skeletal muscle cells",
                        yes = main_cell_type_by_cluster[sobj$seurat_clusters] %>%
                          unname() %>% as.character(),
                        no = sobj$cell_type %>%
                          as.character()) %>%
  as.factor()
```

We change the levels of `cell_type`, for the figures legend to be in a nice order.

```{r change_levels_mpnst}
sobj$cell_type_old = sobj$cell_type
sobj$cell_type = sobj$cell_type %>%
  as.character() %>%
  factor(., levels = custom_order_cell_type$cell_type_old) %>%  # reorder cell type
  `levels<-`(custom_order_cell_type$cell_type)                  # change level names

sobj$cell_type_simple = custom_order_cell_type[as.character(sobj$cell_type_old), "cell_type_simple"]
sobj$cell_type_simple = factor(sobj$cell_type_simple,
                               levels = names(color_markers_simple))
```

We modify color markers to merge DC :

```{r color_markers}
color_markers = color_markers[!(names(color_markers) %in% c("cDC", "pDC"))]
names(color_markers)[names(color_markers) == "mDC"] = "DC"
```


### Cell type composition

This figure represent cell types annotation from all MPNST datasets.

```{r fig3_A1, class.source = 'fold-hide', fig.width = 7, fig.height = 5}
aquarius::fig_plot_dimplot(sobj, reduction_tsne, "cell_type",
                           color_markers, title = "MPNST atlas", order = "TOM+ cells")
```

This figure represent cell types annotation split by genotype :

```{r fig3_A1_split, class.source = 'fold-hide', fig.width = 7, fig.height = 5}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = reduction_tsne,
                                        split_by = "genotype",
                                        group_by = "cell_type",
                                        group_color = unlist(color_markers),
                                        main_pt_size = 0.1,
                                        bg_pt_size = 0.1)

plot_list
```

This figure represent Tomato expression to justify tumor cells annotation.

```{r fig3_A2, class.source = 'fold-hide', fig.width = 6, fig.height = 5}
aquarius::fig_plot_feature(sobj, reduction_tsne, gene = "dtTomato", gene_expression_colors = aquarius::color_gene)
```

This figure represents Prss56 expression :

```{r fig_supp_mpnst_prss56, class.source = 'fold-hide', fig.width = 6, fig.height = 5}
aquarius::fig_plot_feature(sobj, reduction_tsne, gene = "Prss56",
                           gene_expression_colors = aquarius::color_gene, order = TRUE)
```

This barplot represents the composition of each dataset in cell populations :

```{r fig_supp_mpnst_barplot, class.source = 'fold-hide', fig.width = 7, fig.height = 6}
df_proportion = table(sobj$sample_identifier,
                      sobj$cell_type_simple) %>%
  prop.table(., margin = 1) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("sample_identifier", "cell_type_simple", "prop_cells")) %>%
  dplyr::left_join(., y = sample_info, by = "sample_identifier")

plot_content = sample_info %>%
  dplyr::pull(sample_type) %>% table()
plot_content = setNames(nm = names(plot_content), as.vector(plot_content))

plot_list = lapply(names(plot_content), FUN = function(st) {
  p = df_proportion %>%
    dplyr::filter(.data$sample_type == .env$st) %>%
    aquarius::plot_barplot(.,
                           x = "sample_identifier",
                           y = "prop_cells",
                           fill = "cell_type_simple",
                           position = ggplot2::position_fill()) +
    ggplot2::scale_fill_manual(values = color_markers_simple,
                               breaks = names(color_markers_simple),
                               name = "Cell type") +
    ggplot2::scale_y_continuous(breaks = seq(0, 1, by = 0.1),
                                labels = c("0 %", rep("", 4), "50 %", rep("", 4), "100 %"),
                                expand = c(0,0)) +
    ggplot2::theme(axis.title.y = ggplot2::element_blank(),
                   axis.line.x = ggplot2::element_line(colour = "lightgray"),
                   plot.title = ggtext::element_textbox(hjust = 0.5,
                                                        margin = margin(t = 5, b = 5)))
  
  return(p)
})

# Reorder
names(plot_list) = names(plot_content)
plot_list = plot_list[c("flfl_MPNST", "flmn_MPNST")]
plot_content = plot_content[names(plot_list)]

# Remove legend except for the first one
plot_list[c(2:length(plot_list))] = lapply(plot_list[c(2:length(plot_list))], FUN = function(p) {
  p = p +
    ggplot2::theme(axis.line.y = element_blank(),
                   axis.text.y = element_blank(),
                   axis.ticks.y = element_blank())
  return(p)
})

# Patchwork
patchwork::wrap_plots(plot_list, widths = plot_content, nrow = 1) +
  patchwork::plot_layout(guides = "collect")
```

Same plot group by genotype :

```{r fig3_C, class.source = 'fold-hide', fig.width = 4.5, fig.height = 5.5}
df_proportion = table(sobj$sample_type,
                      sobj$cell_type_simple) %>%
  prop.table(., margin = 1) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("sample_type", "cell_type_simple", "prop_cells"))

plot_content = sample_info %>%
  dplyr::pull(sample_type) %>% table()
plot_content = setNames(nm = names(plot_content), as.vector(plot_content))

plot_list = lapply(names(plot_content), FUN = function(st) {
  p = df_proportion %>%
    dplyr::filter(.data$sample_type == .env$st) %>%
    aquarius::plot_barplot(.,
                           x = "sample_type",
                           y = "prop_cells",
                           fill = "cell_type_simple",
                           position = ggplot2::position_fill()) +
    ggplot2::scale_fill_manual(values = color_markers_simple,
                               breaks = names(color_markers_simple),
                               name = "Cell type") +
    ggplot2::scale_y_continuous(breaks = seq(0, 1, by = 0.1),
                                labels = c("0 %", rep("", 4), "50 %", rep("", 4), "100 %"),
                                expand = c(0,0)) +
    ggplot2::theme(axis.title.y = ggplot2::element_blank(),
                   axis.line.x = ggplot2::element_line(colour = "lightgray"),
                   plot.title = ggtext::element_textbox(hjust = 0.5,
                                                        margin = margin(t = 5, b = 5)))
  
  return(p)
})

# Reorder
names(plot_list) = names(plot_content)
plot_list = plot_list[c("flfl_MPNST", "flmn_MPNST")]
plot_content = plot_content[names(plot_list)]

# Remove legend except for the first one
plot_list[c(2:length(plot_list))] = lapply(plot_list[c(2:length(plot_list))], FUN = function(p) {
  p = p +
    ggplot2::theme(axis.line.y = element_blank(),
                   axis.text.y = element_blank(),
                   axis.ticks.y = element_blank())
  return(p)
})

# Patchwork
patchwork::wrap_plots(plot_list, nrow = 1) +
  patchwork::plot_layout(guides = "collect")
```

This dotplot assesses the cell type annotation :

```{r fig_supp_mpnst_dotplot, class.source = 'fold-hide', fig.width = 8, fig.height = 12}
markers = dotplot_markers[levels(sobj$cell_type)] # new order
markers = unique(unlist(markers)) # as vector
markers = markers[!(markers %in% c("Ptprc", "dtTomato"))]
markers = c("Ptprc", markers, "dtTomato")

aquarius::fig_plot_dotplot(sobj,
                           groupby = "cell_type",
                           markers = markers,
                           gene_expression_colors = aquarius::color_gene) +
  ggplot2::geom_vline(xintercept = 0.5 + c(5, 7, 9, 11, 13, 15,
                                           18, 20, 22, 24, 26, 28, 30, 32),
                      color = "gray92") +
  ggplot2::theme(legend.direction = "vertical",
                 legend.box = "horizontal",
                 legend.position = "bottom")
```

How many cells for each cell type, by tumor type ?

```{r fig_supp_mpnst_table, fig.height = 5, fig.width = 8, class.source = "fold-hide"}
df_quantities = table(sobj$cell_type,
                      sobj$genotype) %>%
  prop.table(., margin = 2) %>%
  "*"(100) %>% round(., 2) %>%
  as.data.frame.array(stringAsFactors = FALSE) %>%
  dplyr::mutate(cell_type = rownames(.)) %>%
  dplyr::relocate(cell_type)
df_quantities$nb_cells = rowSums(table(sobj$cell_type, sobj$genotype))
df_quantities = rbind(df_quantities,
                      c("Number of cells", as.numeric(table(sobj$genotype)), ncol(sobj)))

columns_color = unlist(color_markers[levels(sobj$cell_type)])
columns_color = c(columns_color, NA, rep(NA, times = nrow(df_quantities)*(ncol(df_quantities) - 1)))
font_color = ifelse(is.na(columns_color), yes = "black", no = "white")

# Cell type
custom_theme1 = gridExtra::ttheme_minimal(base_colour = font_color,
                                          core = list(bg_params = list(fill = columns_color,
                                                                       col = "black")),
                                          colhead = list(fg_params = list(col = "black")))
gb1 = gridExtra::tableGrob(df_quantities,
                           rows = NULL,
                           cols = c("Cell type", "fl/fl", "fl/-", "Nb cells"),
                           theme = custom_theme1)

# Cell family (merged cells)
custom_theme2 = gridExtra::ttheme_minimal(base_colour = c("black", "white", "white"),
                                          core = list(bg_params = list(fill = family_color,
                                                                       col = "black")),
                                          colhead = list(fg_params = list(col = "black")))
gb2 = data.frame(cell_family = c("immune cells", "stromal cells", "tumor cells")) %>%
  gridExtra::tableGrob(.,
                       row = NULL,
                       cols = c("Cell family"),
                       theme = custom_theme2)

# Merge
gb = gridExtra::gtable_combine(gb2,gb1, along = 1)
gb$layout[gb$layout$t != 1 & gb$layout$l == 1, c("t")] = c(2, 9, 13, 2, 9, 13)
gb$layout[gb$layout$b != 1  & gb$layout$l == 1, c("b")] = c(8, 12, 13, 8, 12, 13)

# Patchwork
title_gg = ggplot2::ggplot() +
  ggplot2::theme_minimal()

patchwork::wrap_plots(list(title_gg, gb)) +
  patchwork::plot_layout(design = "A\nB", heights = c(0.1,5)) &
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15))
```

We save an Excel file with the proportion of cells by cell type and by sample :

```{r excel_df_quantities_mpnst}
df_quantities = table(sobj$cell_type,
                      sobj$sample_identifier) %>%
  prop.table(., margin = 2) %>%
  "*"(100) %>% round(., 2) %>%
  as.data.frame.array(stringAsFactors = FALSE) %>%
  dplyr::mutate(cell_type = rownames(.)) %>%
  dplyr::relocate(cell_type)
df_quantities$nb_cells = rowSums(table(sobj$cell_type, sobj$sample_identifier))
df_quantities = rbind(df_quantities,
                      c("Number of cells", as.numeric(table(sobj$sample_identifier)), ncol(sobj)))

openxlsx::write.xlsx(df_quantities, file = "./data/mpnst_cell_quantities.xlsx")
```


## fl/fl atlas

These are the two samples sorted for Tomato expression, by FACS. Associated datasets contain tumor cells. They will not be represented on the cell type composition figures :

```{r mouse_sample_tomato}
sample_tomato = c("2020_14", "2020_02")
sample_tomato = sample_info %>%
  dplyr::filter(project_name %in% sample_tomato) %>%
  dplyr::pull(sample_identifier) %>%
  as.character()
```

### Dataset preparation

We load the mouse dataset :

```{r load_mouse_dataset}
sobj_full = readRDS(paste0(mouse_dir, "/4_combined_flfl/flfl_sobj.rds"))
reduction_tsne = "harmony_37_tsne"

sobj_full
```


This is the number of cells in each dataset :

```{r fig_supp_flfl_1, fig.width = 5, fig.height = 6}
table(sobj_full$sample_identifier) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("sample_id", "nb_cells")) %>%
  dplyr::left_join(.,
                   y = sample_info[, c("sample_identifier", "color")],
                   by = c("sample_id" = "sample_identifier")) %>%
  aquarius::fig_plot_gb(column_names = c("Sample ID", "Nb cells", "Color")) %>%
  patchwork::wrap_plots(ncol = 1) +
  patchwork::plot_layout(heights = c(0,5))
```

There is a problem with skeletal muscle cells annotation. We change annotation :

```{r main_cell_type_by_cluster_mouse}
main_cell_type_by_cluster = table(sobj_full$seurat_clusters,
                                  sobj_full$cell_type) %>%
  prop.table(., margin = 1) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("clusters", "cell_type", "prop")) %>%
  dplyr::group_by(clusters) %>%
  dplyr::filter(prop == max(prop)) %>%
  dplyr::arrange(clusters) %>%
  as.data.frame()
main_cell_type_by_cluster = setNames(
  nm = main_cell_type_by_cluster$clusters,
  as.character(main_cell_type_by_cluster$cell_type))

sobj_full$cell_type = ifelse(sobj_full$cell_type == "skeletal muscle cells",
                             yes = main_cell_type_by_cluster[sobj_full$seurat_clusters] %>%
                               unname() %>% as.character(),
                             no = sobj_full$cell_type %>%
                               as.character()) %>%
  as.factor()
```

We change the levels of `cell_type`, for the figures legend to be in a nice order.

```{r change_levels_mouse}
sobj_full$cell_type_old = sobj_full$cell_type
sobj_full$cell_type = sobj_full$cell_type %>%
  as.character() %>%
  factor(., levels = custom_order_cell_type$cell_type_old) %>%  # reorder cell type
  `levels<-`(custom_order_cell_type$cell_type)                  # change level names

sobj_full$cell_type_simple = custom_order_cell_type[as.character(sobj_full$cell_type_old), "cell_type_simple"]
sobj_full$cell_type_simple = factor(sobj_full$cell_type_simple,
                                    levels = names(color_markers_simple))
```

We add a `cell_family` columns in metadata :

```{r sobj_cell_family}
sobj_full$cell_family = custom_order_cell_type[as.character(sobj_full$cell_type_old), "cell_family"]
```

### Cell type composition

This figure represent cell types annotation from all mouse datasets.

```{r fig4_A1, class.source = 'fold-hide', fig.width = 7, fig.height = 5}
aquarius::fig_plot_dimplot(sobj_full, reduction_tsne, "cell_type", color_markers, title = "PNST atlas")
```

This figure represent cell types annotation split by tumor type :

```{r fig4_A1_split, class.source = 'fold-hide', fig.width = 7, fig.height = 5}
plot_list = aquarius::plot_split_dimred(sobj_full,
                                        reduction = reduction_tsne,
                                        split_by = "tumor_type",
                                        group_by = "cell_type",
                                        group_color = unlist(color_markers),
                                        main_pt_size = 0.1,
                                        bg_pt_size = 0.1)

plot_list
```

This figure represents Tomato expression to justify tumor cells annotation.

```{r fig4_A2, class.source = 'fold-hide', fig.width = 6, fig.height = 5}
aquarius::fig_plot_feature(sobj_full, reduction_tsne, gene = "dtTomato", gene_expression_colors = aquarius::color_gene)
```

This figure represents Prss56 expression :

```{r fig_supp_flfl_prss56, class.source = 'fold-hide', fig.width = 6, fig.height = 5}
aquarius::fig_plot_feature(sobj_full, reduction_tsne, gene = "Prss56",
                           gene_expression_colors = aquarius::color_gene, order = TRUE)
```
This figure represents Top2a expression :

```{r fig_supp_flfl_top2a, class.source = 'fold-hide', fig.width = 6, fig.height = 5}
aquarius::fig_plot_feature(sobj_full, reduction_tsne, gene = "Top2a",
                           gene_expression_colors = aquarius::color_gene, order = TRUE)
```

This figure represents Mki67 expression :

```{r fig_supp_flfl_mki67, class.source = 'fold-hide', fig.width = 6, fig.height = 5}
aquarius::fig_plot_feature(sobj_full, reduction_tsne, gene = "Mki67",
                           gene_expression_colors = aquarius::color_gene, order = TRUE)
```

This figure represents Ptprc expression :

```{r fig_supp_flfl_ptprc, class.source = 'fold-hide', fig.width = 6, fig.height = 5}
aquarius::fig_plot_feature(sobj_full, reduction_tsne, gene = "Ptprc",
                           gene_expression_colors = aquarius::color_gene, order = TRUE)
```

We represent also Tomato expression split by genotype and tumor type. Cells annotated as fibroblasts in "MPNST" zone, in pNF samples, are Tom-, thus true fibroblasts.

```{r fig_supp_flfl_tomato_split, class.source = 'fold-hide', fig.width = 7, fig.height = 5}
plot_list = aquarius::plot_split_dimred(sobj_full,
                                        reduction = reduction_tsne,
                                        split_by = "tumor_type",
                                        color_by = "dtTomato")

plot_list
```


This figure shows that batch-effect is removed on this projection. It also show cell cycle phase, and a quality metric : the percentage of UMI by cell, for mitochondrial genes 

```{r fig_supp_flfl_qc, class.source = 'fold-hide', fig.width = 7, fig.height = 5}
aquarius::fig_plot_quality(sobj_full, reduction_tsne,
                           gene_expression_colors = aquarius::color_gene,
                           sample_info = sample_info)
```

To quantify cell types in each dataset, we remove the two datasets corresponding to samples that have been sorted for Tomato expression (flfl_dyNF_4, flfl_dyNF_6).

```{r subsobj_flfl}
bool = !(sobj_full$sample_identifier %in% sample_tomato)
subsobj = subset(sobj_full, cells = colnames(sobj_full)[bool])
subsobj$sample_identifier = factor(subsobj$sample_identifier,
                                   levels = intersect(levels(subsobj$sample_identifier), subsobj$sample_identifier))
subsobj
```


This barplot represents the composition of each dataset in cell populations :

```{r prep_barplot_flfl, class.source = 'fold-hide'}
df_proportion = table(subsobj$sample_identifier,
                      subsobj$cell_type_simple) %>%
  prop.table(., margin = 1) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("sample_identifier", "cell_type_simple", "prop_cells")) %>%
  dplyr::left_join(., y = sample_info, by = "sample_identifier")

plot_content = sample_info %>%
  dplyr::filter(!(sample_identifier %in% sample_tomato)) %>%
  dplyr::pull(sample_type) %>% table()
plot_content = setNames(nm = names(plot_content), as.vector(plot_content))

plot_list = lapply(names(plot_content), FUN = function(st) {
  p = df_proportion %>%
    dplyr::filter(.data$sample_type == .env$st) %>%
    aquarius::plot_barplot(.,
                           x = "sample_identifier",
                           y = "prop_cells",
                           fill = "cell_type_simple",
                           position = ggplot2::position_fill()) +
    ggplot2::scale_fill_manual(values = color_markers_simple,
                               breaks = names(color_markers_simple),
                               name = "Cell type") +
    ggplot2::scale_y_continuous(breaks = seq(0, 1, by = 0.1),
                                labels = c("0 %", rep("", 4), "50 %", rep("", 4), "100 %"),
                                expand = c(0,0)) +
    ggplot2::theme(axis.title.y = ggplot2::element_blank(),
                   axis.line.x = ggplot2::element_line(colour = "lightgray"),
                   plot.title = ggtext::element_textbox(hjust = 0.5,
                                                        margin = margin(t = 5, b = 5)))
  
  return(p)
})

# Reorder
names(plot_list) = names(plot_content)
plot_list = plot_list[c("flfl_pNF", "flfl_dyNF", "flfl_MPNST")]
plot_content = plot_content[names(plot_list)]

plot_list_all = plot_list
```

Barplot for all samples :

```{r fig_supp_flfl_barplot, fig.width = 10, fig.height = 6}
# Remove legend except for the first one
plot_list[c(2:length(plot_list))] = lapply(plot_list[c(2:length(plot_list))], FUN = function(p) {
  p = p +
    ggplot2::theme(axis.line.y = element_blank(),
                   axis.text.y = element_blank(),
                   axis.ticks.y = element_blank())
  return(p)
})

# Patchwork
patchwork::wrap_plots(plot_list, widths = plot_content, nrow = 1) +
  patchwork::plot_layout(guides = "collect")
```

This barplot represents the proportion of cell populations, by tumor type for all mice, except the two samples filtered for Tomato expression by FACS :

```{r fig4_C, class.source = 'fold-hide', fig.width = 4.5, fig.height = 4}
df_proportion = table(subsobj$tumor_type,
                      subsobj$cell_family) %>%
  prop.table(., margin = 1) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("sample_type", "cell_family", "prop_cells"))

plot_content = sample_info %>%
  dplyr::filter(project_name %in% as.character(unique(subsobj$project_name))) %>%
  dplyr::pull(tumor_type) %>% table()
plot_content = setNames(nm = names(plot_content), as.vector(plot_content))

plot_list = lapply(names(plot_content), FUN = function(st) {
  p = df_proportion %>%
    dplyr::filter(.data$sample_type == .env$st) %>%
    aquarius::plot_barplot(.,
                           x = "sample_type",
                           y = "prop_cells",
                           fill = "cell_family",
                           position = ggplot2::position_fill()) +
    ggplot2::scale_fill_manual(values = family_color,
                               breaks = names(family_color),
                               name = "Cell family") +
    ggplot2::scale_y_continuous(breaks = seq(0, 1, by = 0.1),
                                labels = c("0 %", rep("", 4), "50 %", rep("", 4), "100 %"),
                                expand = c(0,0)) +
    ggplot2::theme(axis.title.y = ggplot2::element_blank(),
                   axis.line.x = ggplot2::element_line(colour = "lightgray"),
                   plot.title = ggtext::element_textbox(hjust = 0.5,
                                                        margin = margin(t = 5, b = 5)))
  
  return(p)
})

# Reorder
names(plot_list) = names(plot_content)
plot_list = plot_list[c("pNF", "dyNF", "MPNST")]
plot_content = plot_content[names(plot_list)]

# Remove legend except for the first one
plot_list[c(2:length(plot_list))] = lapply(plot_list[c(2:length(plot_list))], FUN = function(p) {
  p = p +
    ggplot2::theme(axis.line.y = element_blank(),
                   axis.text.y = element_blank(),
                   axis.ticks.y = element_blank())
  return(p)
})

# Patchwork
patchwork::wrap_plots(plot_list, nrow = 1) +
  patchwork::plot_layout(guides = "collect")
```


This dotplot assesses the cell type annotation :

```{r fig_supp_flfl_dotplot, class.source = 'fold-hide', fig.width = 8, fig.height = 12}
markers = dotplot_markers[levels(sobj_full$cell_type)] # new order
markers = unique(unlist(markers)) # as vector
markers = markers[!(markers %in% c("Ptprc", "dtTomato"))]
markers = c("Ptprc", markers, "dtTomato")

aquarius::fig_plot_dotplot(sobj_full,
                           groupby = "cell_type",
                           markers = markers,
                           gene_expression_colors = aquarius::color_gene) +
  ggplot2::geom_vline(xintercept = 0.5 + c(5, 7, 9, 11, 13, 15,
                                           18, 20, 22, 24, 26, 28, 30, 32),
                      color = "gray92") +
  ggplot2::theme(legend.direction = "vertical",
                 legend.box = "horizontal",
                 legend.position = "bottom")
```

```{r clean_mouse, echo = FALSE}
rm(dotplot_markers, bool, markers, sample_tomato, df_proportion)
```


How many cells for each cell type, by tumor type ?

```{r fig_supp_flfl_table, fig.height = 5, fig.width = 8, class.source = "fold-hide"}
df_quantities = table(subsobj$cell_type,
                      subsobj$tumor_type) %>%
  prop.table(., margin = 2) %>%
  "*"(100) %>% round(., 2) %>%
  as.data.frame.array(stringAsFactors = FALSE) %>%
  dplyr::mutate(cell_type = rownames(.)) %>%
  dplyr::relocate(cell_type)
df_quantities$nb_cells = rowSums(table(subsobj$cell_type, subsobj$tumor_type))
df_quantities = rbind(df_quantities,
                      c("Number of cells", as.numeric(table(subsobj$tumor_type)), ncol(subsobj)))

columns_color = unlist(color_markers[levels(subsobj$cell_type)])
columns_color = c(columns_color, NA, rep(NA, times = nrow(df_quantities)*(ncol(df_quantities) - 1)))
font_color = ifelse(is.na(columns_color), yes = "black", no = "white")

# Cell type
custom_theme1 = gridExtra::ttheme_minimal(base_colour = font_color,
                                          core = list(bg_params = list(fill = columns_color,
                                                                       col = "black")),
                                          colhead = list(fg_params = list(col = "black")))
gb1 = gridExtra::tableGrob(df_quantities,
                           rows = NULL,
                           cols = c("Cell type", "pNF", "dyNF", "MPNST", "Nb cells"),
                           theme = custom_theme1)

# Cell family (merged cells)
custom_theme2 = gridExtra::ttheme_minimal(base_colour = c("black", "white", "white"),
                                          core = list(bg_params = list(fill = family_color,
                                                                       col = "black")),
                                          colhead = list(fg_params = list(col = "black")))
gb2 = data.frame(cell_family = c("immune cells", "stromal cells", "tumor cells")) %>%
  gridExtra::tableGrob(.,
                       row = NULL,
                       cols = c("Cell family"),
                       theme = custom_theme2)

# Merge
gb = gridExtra::gtable_combine(gb2,gb1, along = 1)
gb$layout[gb$layout$t != 1 & gb$layout$l == 1, c("t")] = c(2, 9, 13, 2, 9, 13)
gb$layout[gb$layout$b != 1  & gb$layout$l == 1, c("b")] = c(8, 12, 13, 8, 12, 13)

# Patchwork
title_gg = ggplot2::ggplot() +
  ggplot2::theme_minimal()

patchwork::wrap_plots(list(title_gg, gb)) +
  patchwork::plot_layout(design = "A\nB", heights = c(0.1,5)) &
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15))
```

We save an Excel file with the proportion of cells by cell type and by sample :

```{r excel_df_quantities_flfl}
df_quantities = table(sobj_full$cell_type,
                      sobj_full$sample_identifier) %>%
  prop.table(., margin = 2) %>%
  "*"(100) %>% round(., 2) %>%
  as.data.frame.array(stringAsFactors = FALSE) %>%
  dplyr::mutate(cell_type = rownames(.)) %>%
  dplyr::relocate(cell_type)
df_quantities$nb_cells = rowSums(table(sobj_full$cell_type, sobj_full$sample_identifier))
df_quantities = rbind(df_quantities,
                      c("Number of cells", as.numeric(table(sobj_full$sample_identifier)), ncol(sobj_full)))

openxlsx::write.xlsx(df_quantities, file = "./data/flfl_cell_quantities.xlsx")
```

```{r clean_prop, echo = FALSE}
rm(df_quantities, custom_theme1, custom_theme2, gb, gb1, gb2,
   title_gg, font_color, columns_color, family_color)
```


## Zoom in fibroblasts

We load the fibroblasts object :

```{r load_fibro}
sobj = readRDS(paste0(mouse_dir, "/6_fibroblasts/mice_fibroblasts_annotated_sobj.rds"))
name2D = "mnn_44_umap"

sobj
```

This is a quantification table, to know where come from the fibroblasts, and how many fibroblasts each dataset contains :

```{r quantif_fibro}
quantif = table(sobj$sample_identifier) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("tumor_type", "nb_cells"))
quantif
```

We set custom colors for fibroblast subtypes :

```{r color_fibro}
color_fibro = setNames(aquarius::c30_palette[c(4, 19, 18, 23) + 1],
                       nm = c("endoFb", "epiFb", "periFb", "CAF"))
color_fibro
```


### UMAP

Where are those cells on the full TSNE ?

```{r figFibro_A, fig.width = 6, fig.height = 6}
sobj_full$is_fibroblasts = colnames(sobj_full) %in% colnames(sobj)

Seurat::DimPlot(sobj_full, group.by = "is_fibroblasts", reduction = reduction_tsne,
                cols = c("lightgray", color_markers[["fibroblasts"]]), order = TRUE) +
  ggplot2::labs(x = "tSNE 1", y = "tSNE 2", title = "Fibroblasts",
                subtitle = paste0(length(colnames(sobj)), " cells")) +
  Seurat::NoLegend() + Seurat::NoAxes() +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
                 plot.subtitle = element_text(hjust = 0.5, face = "bold", size = 12),
                 aspect.ratio = 1)
```

This figure represents the fibroblasts annotation for all datasets, and the proportion of those populations by tumor type.

```{r figFibro_B, fig.width = 10, fig.height = 5}
p1 = aquarius::fig_plot_dimplot(sobj, myred = name2D, groupby = "FB_type",
                                title = "Fibroblasts", colors = color_fibro) +
  ggplot2::theme(legend.position = "right") +
  Seurat::NoAxes()
p2 = aquarius::fig_plot_bp_tt(sobj, groupby = "FB_type", colors = color_fibro,
                              with_cell_family = FALSE, quantif = TRUE)

patchwork::wrap_plots(p1, p2) +
  patchwork::plot_layout(width = c(4, 1.5))
```

This is a figure showing gene expression, for genes of interest.

```{r figFibro_D, class.source = 'fold-hide', fig.width = 4, fig.height = 4}
genes_fibro = c("Cilp", "Thy1", "Gsn", "Efemp1", "Apoe",
                "Vwa1", "Itgb4", "Itga6", "Clec3b", "Gpnmb", "Postn")
plot_list = lapply(genes_fibro,
                   FUN = function(one_gene) {
                     one_plot = aquarius::fig_plot_feature(sobj,
                                                           myred = name2D,
                                                           gene = one_gene,
                                                           gene_expression_colors = aquarius::color_gene,
                                                           pt.size = 0.10) +
                       ggplot2::labs(title = one_gene) +
                       ggplot2::theme(axis.title = element_blank(),
                                      axis.line = element_blank())
                     return(one_plot)
                   })
names(plot_list) = genes_fibro
plot_list
```

### Heatmap

This is a heatmap. Genes in bold are cited in this paper : [https://doi.org/10.1016/j.stem.2018.10.024](https://doi.org/10.1016/j.stem.2018.10.024)

```{r prep_heatmap_fibro, class.source = "fold-hide"}
features_list = list("caf" = c("Postn", "Loxl1", "Thy1", "Col14a1", "Col16a1", "Lsp1", "Pi16", "Ptprs", "Arap1",
                               "Cilp", "Ptn", "Mest", "Tnc", "Plagl1"),
                     "endo" = c("Apod", "Apoe", "Cp", "Gpc3", "Ccl11", "Tgfbi", "Cpe", "Crispld2", "Vwa1", "Vit"),
                     "peri" = c("Lrp4", "Hspb1", "Itga6", "Cav1", "Wnt6", "Cav2", "Sdpr", "Itgb4", "Sfrp5", "Ildr2"),
                     "epi" = c("Clec3b", "Ifi205", "Adamts5", "Gdf10", "Hspb8", "Gda", "Lgals3", "C7", "Fst", "Gpnmb"))
features_group = sapply(c(1:length(features_list)),
                        FUN = function(i) {
                          rep(x = names(features_list)[i],
                              length = length(features_list[[i]]))}) %>% unlist()
features = unlist(features_list)
fibro_from_sciatic_nerves = c("Itga6", "Itgb4", "Sfrp5", "Clec3b", "Gdf10", "Gpnmb",
                              "Crispld2", "Adamts5", "Apod", "Cp",
                              "Cilp", "Ptn", "Mest", "Tnc", "Plagl1")
fibro_from_sciatic_nerves = ifelse(features %in% fibro_from_sciatic_nerves,
                                   yes = "bold",
                                   no = "plain")
cell_annot_1 = "FB_type"
cell_annot_1_levels = c("endoFb", "periFb", "epiFb", "CAF")
sobj@meta.data[, cell_annot_1] = factor(sobj@meta.data[, cell_annot_1],
                                        levels = cell_annot_1_levels)
cell_annot_2 = "seurat_clusters"
cell_annot_2_levels = unique(sobj@meta.data[, cell_annot_2])
sobj@meta.data[, cell_annot_2] = factor(sobj@meta.data[, cell_annot_2],
                                        levels = cell_annot_2_levels)
ht = aquarius::fig_make_heatmap(sobj,
                                features = features,
                                cell_annot_1 = cell_annot_1,
                                cell_annot_1_color = color_fibro,
                                cell_annot_2 = NULL,
                                features_group = features_group,
                                # column_title_rot = 15,
                                # features_fontface = fibro_from_sciatic_nerves
)
```

```{r figFibro_heatmap, fig.width = 8, fig.height = 10}
ComplexHeatmap::draw(ht, # legend_grouping = "original",
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


## Zoom in tumor cells

These are genes of interest for tumor cells :

```{r tumor_genes}
# Genes with IHC validation or of interest in Exome-Seq
genes_tom_traj = c("Sox10", "S100b", "Ngfr", "L1cam", "Chl1", "Pdgfra",
                   "Pdgfrb", "Sox9", "Cd34", "Spp1", "Gas1", "Crabp1",
                   "dtTomato", "Birc5", "Top2a", "Mki67",
                   "Acta2", "Tagln", "Nes",
                   "Cdkn2a", "Trp53", "Eed", "Suz12", "Rbbp4", "Ezh1", "Ezh2",
                   "Mbp", "Mpz", "Ncmap", "Cldn19", "Pmp22", "Scn7a")
genes_patients = c("SOX10", "S100B", "NGFR", "L1CAM", "CHL1", "PDGFRA",
                   "PDGFRB", "SOX9", "CD34", "SPP1", "GAS1", "CRABP1",
                   "CDKN2A", "TP53", "EED", "SUZ12", "RBBP4", "EZH1", "EZH2")

# Genes for the heatmap
genes_heatmap = list(group1 = c("Mpz", "Mbp", "S100b", "Rxrg", "Kcna1", "Rarres2", "Ptn",
                                "C4b", "Vwa1", "Scn7a", "Adam23", "Art3", "Fxyd1", "Cadm1",
                                "Cd59a", "Chl1", "Cmtm5", "Itga1", "Cdh19", "Aatk",
                                "Lgi4", "Plp1", "Erbb3", "L1cam", "Cadm4", "Sox10", "Foxd3",
                                "Mcam", "Ngfr", "Cdkn2a", "Moxd1", "Uchl1", "Pdgfb",
                                "Tagln", "Acta2", "Nes", "Pdgfa", "Postn"),
                     group2 = c("Ccnd1",  "Tnfrsf12a", "Snrpd3", "Tgfbi", "Actr3",
                                "Nme1", "Mif", "Pfn1", "Ybx1", "Ran", "Lgals1",
                                "Snrpf", "Nudc", "Gapdh", "Snrpd1",
                                "S100a6", "Nucks1", "Ube2s", "H2afz", "Banf1",
                                "Hmga2", "Crabp1", "Cd34", "Cks1b",
                                "Dut", "Tubb6", "Lyar", "Tyms", "Dctpp1", "Mcm6",
                                "Top2a", "Birc5", "Mki67"),
                     group3 = c("Nfkbia", "Bgn", "Spp1", "Cxcl12", "Pdlim2", "Fn1", "Cpe", "Tcf4", "Cst3",
                                "Errfi1", "Grn", "Nr2f1",  "Trps1", "Nav1",
                                "Prrx1", "Pbx1", "Wls", "Pcolce",
                                "Loxl1", "Twist1",  "Igfbp4", "Serpinf1", "Ddr2", "Mgp",
                                "Dcn", "Gas1", "Pdgfra", "Sesn3", "Mmp2", "Cd302", "Cdh11", 
                                "Mpzl1", "Itgb5", "Nbl1", "Pdgfrb", "Nrep",
                                "Sox9", "Cdh13", "Spon2", "Bicc1", "Rora"))
```

We load the trajectory object :

```{r load_trajectory}
sobj = readRDS(paste0(mouse_dir, "/5_tumor_cells/mice_tumor_cells_subset_sobj_traj_tinga_FE.rds"))
name2D = "mnn_dm_umap"

sobj
```


These are colors for the three groups for the functional enrichment analysis :

```{r group_colors}
group_colors = c("#77DEBF", "#DFA67A", "#A29EDA")
```

For by-sample representation, we merged samples *flfl_dyNF_3* and *flfl_dyNF_4* because they come from the same mouse :

```{r mouse_info}
row_id = which(sample_info$project_name == "2020_13")

mouse_info = sample_info %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), as.character))
mouse_info = rbind(mouse_info[c(1:(row_id-1)), ],
                   c("2020_13_14", "flfl", "dyNF", "M", "here",
                     "flfl_dyNF_3_4", "flfl_dyNF", "#77DC77"),
                   mouse_info[c((row_id+2):nrow(mouse_info)), ]) %>%
  dplyr::filter(!(project_name %in% c("2020_13", "2020_14"))) %>%
  dplyr::mutate(project_name = factor(project_name, levels = project_name)) %>%
  dplyr::mutate(sample_identifier = factor(sample_identifier, levels = sample_identifier)) %>%
  dplyr::rename(mouse_identifier = sample_identifier)

mouse_info
```

We add a new column in metadata, corresponding to `mouse_identifier` :

```{r mouse_identifier}
sobj$mouse_identifier = ifelse(sobj$sample_identifier %in% c("flfl_dyNF_3", "flfl_dyNF_4"),
                               yes = "flfl_dyNF_3_4",
                               no = as.character(sobj$sample_identifier)) %>%
  factor(., levels = levels(mouse_info$mouse_identifier)) %>%
  base::droplevels()

table(sobj$mouse_identifier)
```


### UMAP

Where are those cells on the full TSNE ?

```{r fig5_A1, fig.width = 6, fig.height = 6}
sobj_full$is_tumor_cells = colnames(sobj_full) %in% colnames(sobj)

Seurat::DimPlot(sobj_full, group.by = "is_tumor_cells", reduction = reduction_tsne,
                cols = c("lightgray", color_markers[["TOM+ cells"]]), order = TRUE) +
  ggplot2::labs(x = "tSNE 1", y = "tSNE 2", title = "Tumor cells",
                subtitle = paste0(length(colnames(sobj)), " cells")) +
  Seurat::NoLegend() + Seurat::NoAxes() +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
                 plot.subtitle = element_text(hjust = 0.5, face = "bold", size = 12),
                 aspect.ratio = 1)
```

This figure represents the sample of origin, to assess batch-effect correction of the projection. It shows cell cycle phase of each cell, and pseudotime values.

```{r fig5_A2, fig.width = 5, fig.height = 6, class.source = 'fold-hide'}
plot_list = list()

## Color by sample
plot_list[[2]] = Seurat::DimPlot(sobj,
                                 group.by = "sample_identifier",
                                 reduction = name2D,
                                 pt.size = 0.25) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$sample_identifier) +
  ggplot2::labs(title = paste(ncol(sobj), "tumor cells from 16 samples", sep = " ")) +
  ggplot2::guides(col = guide_legend(ncol = 1, title = "Sample name",
                                     override.aes = list(size = 2))) +
  ggplot2::theme(legend.position = "none")

## Color by cell cycle
plot_list[[3]] = Seurat::DimPlot(sobj,
                                 group.by = "Seurat.Phase",
                                 reduction = name2D,
                                 pt.size = 0.25) +
  ggplot2::scale_color_manual(values = aquarius::gg_color_hue(3),
                              labels = c("G0 / G1", "G2M", "S")) +
  ggplot2::labs(title = "Cell cycle phase") +
  ggplot2::theme(legend.position = c(0.01, 0.08),
                 legend.direction = "vertical") +
  ggplot2::guides(col = guide_legend(override.aes = list(size = 4)))

## Pseudotime
plot_list[[1]] = Seurat::FeaturePlot(sobj, pt.size = 0.25,
                                     reduction = name2D,
                                     features = "pseudotime") +
  ggplot2::scale_color_gradientn(colors = viridis::viridis(n = 100),
                                 breaks = range(sobj$pseudotime),
                                 labels = c("min", "max")) +
  ggplot2::labs(title = "Pseudotime") +
  ggplot2::theme(legend.position = c(0.01, 0.05),
                 legend.direction = "horizontal")

## Groups of cells
plot_list[[4]] = Seurat::DimPlot(sobj,
                                 reduction = name2D,
                                 group.by = "kmeans_group",
                                 cols = c(group_colors, "black")) +
  ggplot2::labs(title = "Cells groups") +
  ggplot2::theme(legend.position = c(0.05, 0.1)) +
  ggplot2::guides(color = guide_legend(override.aes = list(size = 4)))

## Same theme
plot_list = lapply(plot_list, FUN = function(one_plot) {
  one_plot +
    ggplot2::theme(aspect.ratio = 1,
                   plot.title = element_text(hjust = 0.5)) +
    Seurat::NoAxes()
})

## Patchwork
plot_list
```


This is a representation of tumor cells for each sample of origin :

```{r fig5_A3, class.source = 'fold-hide', fig.width = 1.5, fig.height = 1.7}
plot_list = aquarius::plot_split_dimred(sobj = sobj,
                                        reduction = name2D,
                                        split_by = "mouse_identifier",
                                        split_color = mouse_info$color,
                                        bg_pt_size = 0.01,
                                        main_pt_size = 0.01)

# Patchwork
plot_list
```


This is a figure showing gene expression, for genes of interest.

```{r fig5_C, class.source = 'fold-hide', fig.width = 4, fig.height = 4}
plot_list = lapply(c(genes_tom_traj, "dtTomato"),
                   FUN = function(one_gene) {
                     one_plot = aquarius::fig_plot_feature(sobj,
                                                           myred = name2D,
                                                           gene = one_gene,
                                                           gene_expression_colors = aquarius::color_gene,
                                                           pt.size = 0.10,
                                                           order = TRUE) +
                       ggplot2::labs(title = one_gene) +
                       ggplot2::theme(axis.title = element_blank(),
                                      axis.line = element_blank(),
                                      legend.position = "left")
                     return(one_plot)
                   })
names(plot_list) = c(genes_tom_traj, "dtTomato")
plot_list
```

```{r clean_traj_mouse, echo = FALSE}
rm(plot_list, sobj_full)
```

### Heatmap

We prepare the genes and gene annotation.

```{r heatmap_genes, class.source = "fold-hide"}
features_group = sapply(c(1:length(genes_heatmap)),
                        FUN = function(i) {
                          rep(x = names(genes_heatmap)[i],
                              length = length(genes_heatmap[[i]]))}) %>% unlist() %>% as.vector()
features = unique(unlist(genes_heatmap))
```

We prepare the gene expression matrix :

```{r heatmap_matrix, fig.width = 12, fig.height = 30}
mat_expression = Seurat::GetAssayData(sobj, assay = "RNA", slot = "data")
mat_expression = mat_expression[features, ]
mat_expression = Matrix::t(mat_expression)
mat_expression = dynutils::scale_quantile(mat_expression) # between 0 and 1
mat_expression = Matrix::t(mat_expression)
dim(mat_expression)
```

We prepare all the colors for heatmap annotation :

```{r heatmap_colors_ann, class.source = 'fold-hide'}
list_colors = list()
list_colors[["expression"]] = aquarius::color_cnv
list_colors[["pseudotime"]] = circlize::colorRamp2(breaks = seq(min(sobj$pseudotime),
                                                                max(sobj$pseudotime),
                                                                length.out = 9),
                                                   colors = viridis::viridis(9))
list_colors[["sample_type"]] = c(flfl_pNF = "blue",
                                 flfl_dyNF = "green", flfl_MPNST = "brown")
list_colors[["cell_cycle_phase"]] = aquarius::gg_color_hue(3)
names(list_colors[["cell_cycle_phase"]]) = c("G0 / G1", "G2M", "S")
list_colors[["percent_mito"]] = circlize::colorRamp2(breaks = seq(from = min(sobj$percent.mt),
                                                                  to = max(sobj$percent.mt),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
list_colors[["nFeature_RNA"]] = circlize::colorRamp2(breaks = seq(from = min(sobj$nFeature_RNA),
                                                                  to = max(sobj$nFeature_RNA),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
```

We prepare the heatmap annotation :

```{r heatmap_annotation, class.source = 'fold-hide'}
name_side = "right"
ha_top = HeatmapAnnotation(
  group = anno_block(gp = gpar(fill = group_colors),
                     labels = sort(unique(sobj$kmeans_group))),
  sample_type = sobj$sample_type,
  pseudotime = sobj$pseudotime,
  annotation_legend_param = list(pseudotime = list(at = range(sobj$pseudotime),
                                                   labels = c("min", "max"))),
  col = list(sample_type = list_colors[["sample_type"]],
             pseudotime = list_colors[["pseudotime"]]),
  annotation_name_side = name_side)

ha_bottom = HeatmapAnnotation(
  percent_mito = sobj$percent.mt,
  nFeature_RNA = sobj$nFeature_RNA,
  cell_cycle_phase = as.factor(sobj$Seurat.Phase) %>% `levels<-`(c("G0 / G1", "G2M", "S")),
  col = list(percent_mito = list_colors[["percent_mito"]],
             nFeature_RNA = list_colors[["nFeature_RNA"]],
             cell_cycle_phase = list_colors[["cell_cycle_phase"]]),
  annotation_name_side = name_side)
```

We build the heatmap :

```{r heatmap_ht, class.source = 'fold-hide'}
ht = ComplexHeatmap::Heatmap(
  as.matrix(mat_expression),
  # Expression colors
  col = list_colors[["expression"]],
  heatmap_legend_param = list(title = "expression", at = c(0, 1), 
                              labels = c("low", "high")),
  # Annotation
  top_annotation = ha_top,
  bottom_annotation = ha_bottom,
  # Cells grouping
  column_split = sobj$kmeans_group,
  column_gap = unit(2, "mm"),
  column_title = NULL,
  # Cells order
  column_order = order(sobj$pseudotime),
  cluster_columns = FALSE,
  show_column_dend = FALSE,
  # Cells label
  show_column_names = FALSE,
  # Gene grouping
  row_split = features_group,
  row_gap = unit(2, "mm"),
  row_title = NULL,
  cluster_rows = FALSE,
  show_row_dend = FALSE,
  # Gene labels
  row_labels = features,
  show_row_names = TRUE,
  row_names_side = name_side,
  row_names_gp = grid::gpar(fontsize = 20),
  # Visual aspect
  show_heatmap_legend = TRUE,
  border = TRUE,
  use_raster = FALSE)
```

```{r fig5_B, fig.width = 16.5, fig.height = 44}
ComplexHeatmap::draw(ht, # legend_grouping = "original",
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


### Functional analysis

For each group, we build the barplot from functional enrichment analysis. We load the pre-computed results :

```{r load_func_enr_list}
func_enr_list = readRDS(paste0(mouse_dir, "/5_tumor_cells/mice_tumor_cells_subset_list_results.rds"))
names(func_enr_list)
```

Then, we set the gene sets of interest for each group :

```{r gsea_gene_sets, class.source = "fold-hide"}
group1_gs_name = c("GOBP_GLIAL_CELL_DIFFERENTIATION",
                   "GOBP_ENSHEATHMENT_OF_NEURONS",
                   "GOBP_SCHWANN_CELL_DIFFERENTIATION",
                   "REACTOME_EGR2_AND_SOX10_MEDIATED_INITIATION_OF_SCHWANN_CELL_MYELINATION",
                   "REACTOME_L1CAM_INTERACTIONS",
                   "GOBP_GLIOGENESIS",
                   "GOMF_STRUCTURAL_CONSTITUENT_OF_CYTOSKELETON",
                   "REACTOME_RECYCLING_PATHWAY_OF_L1",
                   "GOCC_MYELIN_SHEATH",
                   "GOCC_GLIAL_CELL_PROJECTION")
group2_gs_name = c("GOBP_CYTOPLASMIC_TRANSLATION",
                   "GOBP_DNA_REPLICATION",
                   "KEGG_RIBOSOME",
                   "KEGG_SPLICEOSOME",
                   "HALLMARK_MYC_TARGETS_V1",
                   "HALLMARK_E2F_TARGETS",
                   "HALLMARK_G2M_CHECKPOINT",
                   "REACTOME_M_PHASE",
                   "REACTOME_CELL_CYCLE_CHECKPOINTS",
                   "REACTOME_DNA_REPLICATION")
group3_gs_name = c("GOBP_OSSIFICATION",
                   "GOBP_AMEBOIDAL_TYPE_CELL_MIGRATION",
                   "GOBP_RESPONSE_TO_TRANSFORMING_GROWTH_FACTOR_BETA",
                   "GOBP_MESENCHYME_DEVELOPMENT",
                   "GOBP_TRANSFORMING_GROWTH_FACTOR_BETA_RECEPTOR_SIGNALING_PATHWAY",
                   "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",
                   "HALLMARK_TNFA_SIGNALING_VIA_NFKB",
                   "HALLMARK_HYPOXIA",
                   "HALLMARK_APOPTOSIS",
                   "REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION")
```

We make a dataframe containing scores for each gene sets and each cells, for the three groups :

```{r figY_1, class.source = "fold-hide", fig.width = 15, fig.height = 6}
df_score_group1 = func_enr_list[["group 1_vs_group 3"]]$gsea_up[group1_gs_name, ]
df_score_group1$group = "group 1"
df_score_group1$NES = abs(df_score_group1$NES)

ggplot2::ggplot(df_score_group1, aes(x = stats::reorder(.data$ID, .data$NES),
                                     y = .data$NES, fill = .data$group)) +
  ggplot2::geom_bar(stat = "identity", col = "black") +
  ggplot2::coord_flip(clip = "off") +
  ggplot2::geom_hline(yintercept = 0) +
  ggplot2::geom_label(mapping = aes(label = .data$ID, y = 0.05), fill = NA, label.size = NA,
                      hjust = 0, size = 4.5, fontface = "bold") +
  ggplot2::facet_grid(group ~ ., scales = "free", space = "free") +
  ggplot2::scale_x_discrete(position = "top") +
  ggplot2::labs(y = "Normalized Enrichment Score (NES)", title = "") +
  ggplot2::geom_label(aes(label = round(.data$p.adjust, 4), y = (.data$NES + 0.1)),
                      fill = NA, label.size = 0, size = 5) +
  ggplot2::scale_fill_manual(values = group_colors[1],
                             guide = "none") +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "top",
                 legend.direction = "horizontal",
                 axis.title.y = element_blank(),
                 panel.border = element_blank(),
                 axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 strip.background = element_blank(),
                 strip.text = element_blank(),
                 panel.spacing = grid::unit(2, "lines"),
                 panel.grid.minor.x = element_blank(),
                 panel.grid.major.y = element_blank(),
                 axis.text.x = element_text(size = 15),
                 axis.title.x = element_text(size = 15),
                 axis.line.x = element_line(size = 0.5, color = "black"))
```


```{r figY_2, class.source = "fold-hide", fig.width = 15, fig.height = 6}
df_score_group2 = func_enr_list[["group 1_vs_group 2"]]$gsea_down[group2_gs_name, ]
df_score_group2$group = "group 2"
df_score_group2$NES = abs(df_score_group2$NES)

ggplot2::ggplot(df_score_group2, aes(x = stats::reorder(.data$ID, .data$NES),
                                     y = .data$NES, fill = .data$group)) +
  ggplot2::geom_bar(stat = "identity", col = "black") +
  ggplot2::coord_flip(clip = "off") +
  ggplot2::geom_hline(yintercept = 0) +
  ggplot2::geom_label(mapping = aes(label = .data$ID, y = 0.05), fill = NA, label.size = NA,
                      hjust = 0, size = 4.5, fontface = "bold") +
  ggplot2::facet_grid(group ~ ., scales = "free", space = "free") +
  ggplot2::scale_x_discrete(position = "top") +
  ggplot2::labs(y = "Normalized Enrichment Score (NES)", title = "") +
  ggplot2::geom_label(aes(label = round(.data$p.adjust, 4), y = (.data$NES + 0.1)),
                      fill = NA, label.size = 0, size = 5) +
  ggplot2::scale_fill_manual(values = group_colors[2],
                             guide = "none") +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "top",
                 legend.direction = "horizontal",
                 axis.title.y = element_blank(),
                 panel.border = element_blank(),
                 axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 strip.background = element_blank(),
                 strip.text = element_blank(),
                 panel.spacing = grid::unit(2, "lines"),
                 panel.grid.minor.x = element_blank(),
                 panel.grid.major.y = element_blank(),
                 axis.text.x = element_text(size = 15),
                 axis.title.x = element_text(size = 15),
                 axis.line.x = element_line(size = 0.5, color = "black"))
```


```{r figY_3, class.source = "fold-hide", fig.width = 15, fig.height = 6}
df_score_group3 = func_enr_list[["group 1_vs_group 3"]]$gsea_down[group3_gs_name, ]
df_score_group3$group = "group 3"
df_score_group3$NES = abs(df_score_group3$NES)

ggplot2::ggplot(df_score_group3, aes(x = stats::reorder(.data$ID, .data$NES),
                                     y = .data$NES, fill = .data$group)) +
  ggplot2::geom_bar(stat = "identity", col = "black") +
  ggplot2::coord_flip(clip = "off") +
  ggplot2::geom_hline(yintercept = 0) +
  ggplot2::geom_label(mapping = aes(label = .data$ID, y = 0.05), fill = NA, label.size = NA,
                      hjust = 0, size = 4.5, fontface = "bold") +
  ggplot2::facet_grid(group ~ ., scales = "free", space = "free") +
  ggplot2::scale_x_discrete(position = "top") +
  ggplot2::labs(y = "Normalized Enrichment Score (NES)", title = "") +
  ggplot2::geom_label(aes(label = round(.data$p.adjust, 4), y = (.data$NES + 0.1)),
                      fill = NA, label.size = 0, size = 5) +
  ggplot2::scale_fill_manual(values = group_colors[3],
                             guide = "none") +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "top",
                 legend.direction = "horizontal",
                 axis.title.y = element_blank(),
                 panel.border = element_blank(),
                 axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 strip.background = element_blank(),
                 strip.text = element_blank(),
                 panel.spacing = grid::unit(2, "lines"),
                 panel.grid.minor.x = element_blank(),
                 panel.grid.major.y = element_blank(),
                 axis.text.x = element_text(size = 15),
                 axis.title.x = element_text(size = 15),
                 axis.line.x = element_line(size = 0.5, color = "black"))
```


```{r clean_tumor_cells2, echo = FALSE}
rm(ha_top, ht, ha_bottom, list_colors, mat_expression, name2D, name_side,
   sobj, features, group_colors, sample_info, mouse_dir,
   reduction_tsne, df_score_group1, df_score_group2, df_score_group3,
   group1_gs_name, group2_gs_name, group3_gs_name,
   plot_content, func_enr_list)
```


# Grafting

In this section, we make the figures related to transplantation experiment.

```{r graft_settings}
sample_info = readRDS(paste0(graft_dir, "/1_metadata/graft_sample_info.rds")) %>%
  dplyr::mutate(sample_identifier = ifelse(sample_identifier == "donor",
                                           yes = "primary",
                                           no = "transplanted"))
sample_info
```


## Dataset prepration

We load the tumor cells. Trajectory has been inferred on tumor cells. We load the trajectory object :

```{r load_graft}
sobj = readRDS(paste0(graft_dir, "/4_tumor_cells/graft_tumor_cells_sobj_traj_slingshot.rds"))
name2D = "mnn_dm_umap"

sobj$sample_identifier = ifelse(is.na(sobj$sample_identifier),
                                yes = "primary", no = "transplanted")

sobj
```

## Zoom on tumor cells

We build the figures showing Tomato expression levels, pseudotime, sample of origin and cell cycle phase.

```{r figGraft_A, class.source = 'fold-hide', fig.width = 4, fig.height = 3.75}
plot_list = list()

## Tomato expression
plot_list[[1]] = Seurat::FeaturePlot(sobj, 
                                     reduction = name2D,
                                     features = "dtTomato",
                                     pt.size = 0.25) +
  ggplot2::labs(title = "tdTomato") +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
  ggplot2::theme(legend.position = c(0.01, 0.1),
                 legend.direction = "horizontal")

## Color by sample
plot_list[[2]] = Seurat::DimPlot(sobj,
                                 group.by = "sample_identifier",
                                 reduction = name2D,
                                 pt.size = 0.25) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$sample_identifier) +
  ggplot2::labs(title = "Sample of origin") +
  ggplot2::guides(col = guide_legend(override.aes = list(size = 3))) +
  ggplot2::theme(legend.position = c(0.05, 0.1),
                 legend.title = element_blank())

## Color by cell cycle
plot_list[[3]] = Seurat::DimPlot(sobj,
                                 group.by = "Seurat.Phase",
                                 reduction = name2D,
                                 pt.size = 0.25) +
  ggplot2::scale_color_manual(values = aquarius::gg_color_hue(3),
                              labels = c("G0 / G1", "G2M", "S")) +
  ggplot2::labs(title = "Cell cycle phase") +
  ggplot2::theme(legend.position = c(0.05, 0.1),
                 legend.direction = "vertical") +
  ggplot2::guides(col = guide_legend(override.aes = list(size = 3)))

## Pseudotime
plot_list[[4]] = Seurat::FeaturePlot(sobj, pt.size = 0.25,
                                     reduction = name2D,
                                     features = "pseudotime") +
  ggplot2::scale_color_gradientn(colors = viridis::viridis(n = 100),
                                 breaks = range(sobj$pseudotime, na.rm = TRUE),
                                 labels = c("min", "max")) +
  ggplot2::labs(title = "Pseudotime") +
  ggplot2::theme(legend.position = c(0.01, 0.1),
                 legend.direction = "horizontal")

## Same theme
plot_list = lapply(plot_list, FUN = function(one_plot) {
  one_plot +
    ggplot2::theme(aspect.ratio = 1,
                   plot.title = element_text(hjust = 0.5),
                   legend.text = element_text(size = 12)) +
    Seurat::NoAxes()
})

## Patchwork
plot_list
```


We make the feature plots :

```{r figGraft_C, class.source = 'fold-hide', fig.width = 3, fig.height = 2.5}
genes_oi = c("dtTomato", genes_tom_traj)
genes_oi = genes_oi[genes_oi %in% rownames(sobj)]

plot_list = lapply(genes_oi,
                   FUN = function(one_gene) {
                     one_plot = aquarius::fig_plot_feature(sobj,
                                                           myred = name2D,
                                                           gene = one_gene,
                                                           gene_expression_colors = aquarius::color_gene,
                                                           pt.size = 0.10, order = TRUE) +
                       ggplot2::labs(title = one_gene) +
                       ggplot2::theme(axis.title = element_blank(),
                                      axis.line = element_blank(),
                                      legend.position = "left")
                     return(one_plot)
                   })

names(plot_list) = genes_oi
plot_list
```

Cdkn2a split by condition :

```{r figGraft_D, class.source = 'fold-hide', fig.width = 3, fig.height = 2.5}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "project_name",
                                        color_by = "Cdkn2a",
                                        main_pt_size = 0.5,
                                        bg_pt_size = 0.5,
                                        bg_color = "gray98",
                                        order = TRUE)

plot_list
```

## Heatmap

We can represent gene expression levels on a heatmap.

```{r graft_heatmap_genes, class.source = "fold-hide"}
features_group = sapply(c(1:length(genes_heatmap)),
                        FUN = function(i) {
                          rep(x = names(genes_heatmap)[i],
                              length = length(genes_heatmap[[i]]))}) %>% unlist() %>% as.vector()
features = unique(unlist(genes_heatmap))

features[!(features %in% rownames(sobj))]
features = features[features %in% rownames(sobj)]
features_group = features_group[features]
```

We prepare the gene expression matrix :

```{r graft_heatmap_matrix, fig.width = 12, fig.height = 30}
mat_expression = Seurat::GetAssayData(sobj, assay = "RNA", slot = "data")
mat_expression = mat_expression[features, ]
mat_expression = Matrix::t(mat_expression)
mat_expression = dynutils::scale_quantile(mat_expression) # between 0 and 1
mat_expression = Matrix::t(mat_expression)
dim(mat_expression)
```

We prepare all the colors for heatmap annotation :

```{r graft_heatmap_colors_ann, class.source = 'fold-hide'}
list_colors = list()
list_colors[["expression"]] = c("#2166AC", "#F7F7F7", "#B2182B")
list_colors[["pseudotime"]] = circlize::colorRamp2(breaks = seq(min(sobj$pseudotime, na.rm = TRUE),
                                                                max(sobj$pseudotime, na.rm = TRUE),
                                                                length.out = 9),
                                                   colors = viridis::viridis(9))
list_colors[["sample"]] = setNames(nm = sample_info$sample_identifier,
                                   sample_info$color)
list_colors[["cell_cycle_phase"]] = aquarius::gg_color_hue(3)
names(list_colors[["cell_cycle_phase"]]) = c("G1", "G2M", "S")
list_colors[["percent_mito"]] = circlize::colorRamp2(breaks = seq(from = min(sobj$percent.mt),
                                                                  to = max(sobj$percent.mt),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
list_colors[["nFeature_RNA"]] = circlize::colorRamp2(breaks = seq(from = min(sobj$nFeature_RNA),
                                                                  to = max(sobj$nFeature_RNA),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
```


We prepare the heatmap annotation :

```{r graft_heatmap_annotation, class.source = 'fold-hide'}
name_side = "right"
ha_top = HeatmapAnnotation(sample = sobj$sample_identifier,
                           pseudotime = sobj$pseudotime,
                           col = list(sample = list_colors[["sample"]],
                                      pseudotime = list_colors[["pseudotime"]]),
                           annotation_legend_param = list(pseudotime = list(at = range(sobj$pseudotime, na.rm = TRUE),
                                                                            labels = c("min", "max"))),
                           annotation_name_side = name_side)
ha_bottom = HeatmapAnnotation(percent_mito = sobj$percent.mt,
                              nFeature_RNA = sobj$nFeature_RNA,
                              cell_cycle_phase = sobj$Seurat.Phase,
                              col = list(percent_mito = list_colors[["percent_mito"]],
                                         nFeature_RNA = list_colors[["nFeature_RNA"]],
                                         cell_cycle_phase = list_colors[["cell_cycle_phase"]]),
                              annotation_name_side = name_side)
```

We build the heatmap :

```{r graft_heatmap_ht, class.source = 'fold-hide'}
ht = ComplexHeatmap::Heatmap(as.matrix(mat_expression),
                             # Expression colors
                             col = list_colors[["expression"]],
                             heatmap_legend_param = list(title = "expression", at = c(0, 1), 
                                                         labels = c("low", "high")),
                             # Annotation
                             top_annotation = ha_top,
                             bottom_annotation = ha_bottom,
                             # Cells grouping
                             column_split = !is.na(sobj$pseudotime),
                             column_gap = unit(2, "mm"),
                             column_title = NULL,
                             # Cells order
                             column_order = order(sobj$pseudotime, na.last = FALSE),
                             cluster_columns = FALSE,
                             show_column_dend = FALSE,
                             # Cells label
                             show_column_names = FALSE,
                             # Gene grouping
                             row_split = features_group,
                             row_gap = unit(2, "mm"),
                             row_title = NULL,
                             # row_title = NULL,
                             cluster_rows = FALSE,
                             show_row_dend = FALSE,
                             # Gene labels
                             show_row_names = TRUE,
                             # Visual aspect
                             show_heatmap_legend = TRUE,
                             border = TRUE,
                             use_raster = FALSE)
```

```{r figGraft_B, fig.width = 9, fig.height = 22}
ComplexHeatmap::draw(ht, # legend_grouping = "original",
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


```{r clean_graft, echo = FALSE}
rm(ha_bottom, ha_top, ht, list_colors, mat_expression,
   plot_list, sobj, features, name_side, name2D,
   genes_tom_traj, graft_dir)
```


# Patients

In this section, we make the figure related to NF1 patients datasets.

## Dataset preparation

We load the human dataset.

```{r load_human_dataset}
sobj_full = readRDS(paste0(patient_dir, "/3_combined/all_patients_sobj.rds"))
sobj_full
```

We set the global settings for this analysis :

```{r variable_human}
sample_info = readRDS(paste0(patient_dir, "/1_metadata/patients_sample_info.rds"))
dotplot_markers = readRDS(paste0(patient_dir, "/1_metadata/patients_dotplot_markers.rds"))
color_markers = readRDS(paste0(patient_dir, "/1_metadata/patients_color_markers.rds"))
color_markers[["mural cells"]] = "gray60"

reduction_tsne = "harmony_80_tsne"
```

This is the correspondence between cell types and cell families :

```{r human_custom_order_cell_type}
custom_order_cell_type = data.frame(cell_type = c("cDC", "mast cells", "B cells", "NK cells",
                                                  "T cells", "M\U03A6/Mo", "mural cells",
                                                  "endothelial cells", "fibroblasts", "tumor cells"),
                                    cell_type_old = c("cDC", "mast cells", "B cells", "NK cells",
                                                      "T cells", "macrophages", "mural cells",
                                                      "endothelial cells", "fibroblasts", "tumor cells"),
                                    cell_family = c(rep("immune cells", 6),
                                                    rep("stromal cells", 3),
                                                    rep("tumor cells", 1)),
                                    stringsAsFactors = FALSE)
rownames(custom_order_cell_type) = custom_order_cell_type$cell_type
```


We add those information in the object :

```{r human_relevel_cell_type}
sobj_full$cell_type = factor(sobj_full$cell_type,
                             levels = levels(sobj_full$cell_type)[match(custom_order_cell_type$cell_type_old,
                                                                        levels(sobj_full$cell_type))]) # reorder cell type
levels(sobj_full$cell_type) = custom_order_cell_type$cell_type # change level names
```

We change also levels in `color_markers` and `dotplot_markers` :

```{r human_relevel_cell_type_markers}
names(color_markers)[names(color_markers) == "macrophages"] = "M\U03A6/Mo"
color_markers = color_markers[custom_order_cell_type$cell_type]

names(dotplot_markers)[names(dotplot_markers) == "macrophages"] = "M\U03A6/Mo"
dotplot_markers = dotplot_markers[custom_order_cell_type$cell_type]
```

We group cell type by cluster, to correct mis-annotation :

```{r human_cluster_type}
cluster_type = table(sobj_full$cell_type, sobj_full$seurat_clusters) %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
cluster_type = setNames(nm = names(cluster_type),
                        levels(sobj_full$cell_type)[cluster_type])

sobj_full$cluster_type = cluster_type[sobj_full$seurat_clusters]
sobj_full$cluster_type = factor(sobj_full$cluster_type,
                                levels = levels(sobj_full$cell_type))
```

We add a `cell_family` columns in metadata :

```{r human_cell_family}
sobj_full$cell_family = custom_order_cell_type[sobj_full$cluster_type, "cell_family"]
```

## Cell type composition

This figure represent cell types annotation from all patients datasets :

```{r figPatients_B, class.source = 'fold-hide', fig.width = 7, fig.height = 5}
aquarius::fig_plot_dimplot(sobj_full, reduction_tsne,
                           "cluster_type", color_markers, title = "NF1 patients")
```

Same figure split by tumor type :

```{r figPatients_C, class.source = 'fold-hide', fig.width = 7, fig.height = 5}
plot_list = aquarius::plot_split_dimred(sobj_full,
                                        reduction = reduction_tsne,
                                        split_by = "tumor_type",
                                        group_by = "cluster_type",
                                        group_color = unlist(color_markers),
                                        main_pt_size = 0.1,
                                        bg_pt_size = 0.1)

plot_list
```

This figure shows that batch-effect is removed on this projection. It also show cell cycle phase, and a quality metric : the percentage of UMI by cell, for mitochondrial genes 

```{r figPatients_Z, class.source = 'fold-hide', fig.width = 7, fig.height = 6.5}
aquarius::fig_plot_quality(sobj_full, reduction_tsne,
                           sample_info = sample_info,
                           gene_expression_colors = aquarius::color_gene)
```

Genes of interest :

```{r figPatients_genes, class.source = 'fold-hide', fig.width = 6, fig.height = 5}
plot_list = lapply(c("PTPRC", "SOX10", "NGFR", "L1CAM", "CHL1",
                     "PDGFRA", "PDGFRB", "CD34", "SPP1", "CRABP1", "SOX9", "DLK1"), FUN = function(one_gene) {
                       p = aquarius::fig_plot_feature(sobj_full, reduction_tsne, gene = one_gene,
                                                      gene_expression_colors = aquarius::color_gene,
                                                      order = TRUE)
                       return(p)
                     })
plot_list
```

This dotplot assesses the cell type annotation :

```{r figPatients_Y, class.source = 'fold-hide', fig.width = 8, fig.height = 9}
markers = dotplot_markers[levels(sobj_full$cluster_type)] # new order
markers = unique(unlist(markers)) # as vector
markers = markers[!(markers %in% c("PTPRC"))]
markers = c("PTPRC", markers)
markers = markers[markers %in% rownames(sobj_full)]

aquarius::fig_plot_dotplot(sobj_full,
                           groupby = "cluster_type",
                           markers = markers,
                           gene_expression_colors = aquarius::color_gene) +
  ggplot2::geom_vline(xintercept = 0.5 + c(4, 6, 7, 9, 11, 14, 15, 17, 18, 19, 20),
                      color = "gray92") +
  ggplot2::theme(legend.direction = "vertical",
                 legend.box = "horizontal",
                 legend.position = "bottom")
```

```{r clean_markers, echo = FALSE}
rm(dotplot_markers, markers)
```

## Zoom on tumor cells

We load the tumor cells object :

```{r tumor_cells_patients}
sobj_tum = readRDS(paste0(patient_dir, "/4_tumor_cells/patients_tumor_cells_sobj.rds"))
name2D = "mnn_39_tsne"

sobj_tum
```


Where are those cells on the full TSNE ?

```{r figPatients_D1, fig.width = 6, fig.height = 6}
sobj_full$is_tumor_cells = colnames(sobj_full) %in% colnames(sobj_tum)

Seurat::DimPlot(sobj_full, group.by = "is_tumor_cells", reduction = reduction_tsne,
                cols = c("lightgray", color_markers[["tumor cells"]])) +
  ggplot2::labs(x = "tSNE 1", y = "tSNE 2", title = "Tumor cells",
                subtitle = paste0(length(colnames(sobj_tum)), " cells")) +
  Seurat::NoLegend() + Seurat::NoAxes() +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
                 plot.subtitle = element_text(hjust = 0.5, face = "bold", size = 12),
                 aspect.ratio = 1)
```

Where are samples of origin on the tumor object ?

```{r figPatients_D2, fig.width = 6, fig.height = 6}
Seurat::DimPlot(sobj_tum,
                group.by = "orig.ident",
                pt.size = 0.1,
                reduction = name2D,
                cols = setNames(sample_info$color,
                                nm = sample_info$project_name)) +
  ggplot2::labs(title = paste(ncol(sobj_tum), "tumor cells from", nrow(sample_info), "samples", sep = " ")) +
  ggplot2::guides(col = guide_legend(ncol = 1,
                                     override.aes = list(size = 2))) +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.title = element_blank(),
                 legend.background = element_rect(fill = alpha(colour = "white", alpha = 0))) +
  ggplot2::guides(col = guide_legend(override.aes = list(size = 3))) +
  Seurat::NoAxes()
```

We represent expression level for some genes of interest :

```{r figPatients_D3, class.source = 'fold-hide', fig.width = 3, fig.height = 3}
plot_list = lapply(genes_patients,
                   FUN = function(one_gene) {
                     one_plot = Seurat::FeaturePlot(sobj_tum,
                                                    reduction = name2D,
                                                    features = one_gene,
                                                    pt.size = 0.10,
                                                    order = TRUE) +
                       ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
                       ggplot2::labs(title = one_gene) +
                       Seurat::NoAxes() +
                       ggplot2::theme(legend.position = "left")
                     return(one_plot)
                   })
plot_list
```

## Heatmap

We make the same heatmap as for mouse :

```{r patients_heatmap_genes, class.source = "fold-hide"}
genes_heatmap = lapply(genes_heatmap, FUN = toupper)
features_group = sapply(c(1:length(genes_heatmap)),
                        FUN = function(i) {
                          rep(x = names(genes_heatmap)[i],
                              length = length(genes_heatmap[[i]]))}) %>% unlist() %>% as.vector()
features = unique(unlist(genes_heatmap))

features[which(features == "CD59A")] = "CD59"

features[!(features %in% rownames(sobj_tum))]
features = features[features %in% rownames(sobj_tum)]
features_group = features_group[features]
```

We prepare the gene expression matrix :

```{r patients_heatmap_matrix, fig.width = 12, fig.height = 30}
mat_expression = Seurat::GetAssayData(sobj_tum, assay = "RNA", slot = "data")
mat_expression = mat_expression[features, ]
mat_expression = Matrix::t(mat_expression)
mat_expression = dynutils::scale_quantile(mat_expression) # between 0 and 1
mat_expression = Matrix::t(mat_expression)
dim(mat_expression)
```

We prepare all the colors for heatmap annotation :

```{r patients_heatmap_colors_ann, class.source = 'fold-hide'}
list_colors = list()
list_colors[["expression"]] = c("#2166AC", "#F7F7F7", "#B2182B")
list_colors[["sample"]] = setNames(nm = sample_info$project_name,
                                   sample_info$color)
list_colors[["sample_type"]] = setNames(nm = c("pNF", "dyNF", "MPNST"),
                                        c("blue", "green", "purple"))
list_colors[["percent_mito"]] = circlize::colorRamp2(breaks = seq(from = min(sobj_tum$percent.mt),
                                                                  to = max(sobj_tum$percent.mt),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
list_colors[["nFeature_RNA"]] = circlize::colorRamp2(breaks = seq(from = min(sobj_tum$nFeature_RNA),
                                                                  to = max(sobj_tum$nFeature_RNA),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
```


We prepare the heatmap annotation :

```{r patients_heatmap_annotation, class.source = 'fold-hide'}
name_side = "right"
ha_top = HeatmapAnnotation(sample_type = sobj_tum$tumor_type,
                           sample = sobj_tum$project_name,
                           col = list(sample = list_colors[["sample"]],
                                      sample_type = list_colors[["sample_type"]],
                                      pseudotime = list_colors[["pseudotime"]]),
                           annotation_name_side = name_side)
ha_bottom = HeatmapAnnotation(percent_mito = sobj_tum$percent.mt,
                              nFeature_RNA = sobj_tum$nFeature_RNA,
                              col = list(percent_mito = list_colors[["percent_mito"]],
                                         nFeature_RNA = list_colors[["nFeature_RNA"]]),
                              annotation_name_side = name_side)
```

We build the heatmap :

```{r patients_heatmap_ht, class.source = 'fold-hide'}
ht = ComplexHeatmap::Heatmap(as.matrix(mat_expression),
                             # Expression colors
                             col = list_colors[["expression"]],
                             heatmap_legend_param = list(title = "expression", at = c(0, 1), 
                                                         labels = c("low", "high")),
                             # Annotation
                             top_annotation = ha_top,
                             bottom_annotation = ha_bottom,
                             # Cells grouping
                             column_split = sobj_tum$tumor_type,
                             column_gap = unit(1, "mm"),
                             column_title = NULL,
                             # Cells order
                             column_order = order(sobj_tum$project_name),
                             cluster_columns = FALSE,
                             show_column_dend = FALSE,
                             # Cells label
                             show_column_names = FALSE,
                             # Gene grouping
                             row_split = features_group,
                             row_gap = unit(1, "mm"),
                             row_title = NULL,
                             # row_title = NULL,
                             cluster_rows = FALSE,
                             show_row_dend = FALSE,
                             # Gene labels
                             show_row_names = TRUE,
                             # Visual aspect
                             show_heatmap_legend = TRUE,
                             border = TRUE,
                             use_raster = FALSE)
```

```{r figPatients_heatmap, fig.width = 10, fig.height = 20}
ComplexHeatmap::draw(ht, # legend_grouping = "original",
                     merge_legend = TRUE,
                     heatmap_legend_side = "left",
                     annotation_legend_side = "left")
```


```{r clean_patients, echo = FALSE}
rm(ha_bottom, ha_top, ht, list_colors, mat_expression,
   plot_list, sobj_full, features, name_side, name2D, patient_dir)
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```

