---
title: "Mice"
subtitle: "Tumor cells functional enrichment"
author: "Audrey"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <<- Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


This script is used to :

* explore MSigDB gene sets : hallmark, GO BP, GO MF, GO CC, Kegg pathways, PID, Wiki pathways and Biocarta gene sets
* do functional enrichment and GSE analysis with `clusterProfiler`

```{r library}
library(dplyr)
library(org.Mm.eg.db)
library(ComplexHeatmap)
library(ggplot2)

.libPaths()
```

# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "mice_tumor_cells_subset"
out_dir = "."
```


We load the sample information :

```{r custom_palette_sample, fig.width = 8, fig.height = 8}
sample_info = readRDS(paste0(out_dir, "/../1_metadata/mice_sample_info.rds"))
sample_info = sample_info %>%
  dplyr::filter(genotype == "flfl") %>%
  base::droplevels()
project_names_oi = sample_info$project_name

graphics::pie(rep(1, nrow(sample_info)),
              col = sample_info$color,
              labels = sample_info$project_name)
```

We load the Seurat object :

```{r load_sobj}
sobj = readRDS(paste0(out_dir, "/", save_name, "_sobj_traj_tinga.rds"))
sobj
```

This is the projection name to represent cells :

```{r name2D}
name2D = "mnn_dm_umap"
```


# Clustering

We can visualize sample of origin of tumor cells, on the UMAP, and the pseudotime.

```{r dimplot, fig.width = 12, fig.height = 5}
sample_of_origin = Seurat::DimPlot(sobj,
                                   group.by = "project_name",
                                   reduction = name2D) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$sample_identifiant) +
  ggplot2::labs(title = "Sample of origin") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
  Seurat::NoAxes()

pseudotime = Seurat::FeaturePlot(sobj,
                                 features = "pseudotime",
                                 reduction = name2D) +
  ggplot2::scale_color_gradientn(colors = viridis::viridis(n = 100)) +
  ggplot2::theme(aspect.ratio = 1) +
  Seurat::NoAxes()

sample_of_origin | pseudotime
```

We visualize pre-existing clusters :

```{r seurat_clusters, fig.width = 12, fig.height = 8}
plot_list = lapply(c("Mbp", "Mpz", "Mal"), FUN = function(one_gene) {
  p = Seurat::FeaturePlot(sobj, features = one_gene,
                          reduction = name2D) +
    ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
    ggplot2::theme(aspect.ratio = 1) +
    Seurat::NoAxes()
  return(p)
  
})
plot_list[[length(plot_list) + 1]] = Seurat::DimPlot(sobj,
                reduction = name2D,
                group.by = "seurat_clusters",
                label = TRUE,
                cols = aquarius::gg_color_hue(length(levels(sobj$seurat_clusters)))) +
  Seurat::NoAxes() + Seurat::NoLegend() +
  ggplot2::theme(aspect.ratio = 1)

patchwork::wrap_plots(plot_list) +
  patchwork::plot_layout(design = "DDDA\nDDDB\nDDDC")
```

There is a contamination cluster (myelinating Schwann cells) :

```{r contam_clust}
contam_clust = 14
```


We cluster cells into three groups, using k-means :

```{r clustering_big, fig.width = 8, fig.height = 8}
# kmeans
kmeans_obj = sobj@reductions[["mnn"]]@cell.embeddings %>%
  as.data.frame()

set.seed(1337L)
kmeans_obj = stats::kmeans(kmeans_obj, centers = 3, nstart = 10, 
                           algorithm = "Hartigan-Wong", iter.max = 500)

# add column in method data
sobj$kmeans_group = kmeans_obj$cluster[colnames(sobj)]

# relevel according to pseudotime
kmeans_group_levels = sobj@meta.data[, c("kmeans_group", "pseudotime")] %>%
  dplyr::group_by(kmeans_group) %>%
  dplyr::summarise(avg_pdt = mean(pseudotime)) %>%
  dplyr::arrange(avg_pdt) %>%
  dplyr::pull(kmeans_group)

sobj$kmeans_group = factor(sobj$kmeans_group, levels = kmeans_group_levels)
levels(sobj$kmeans_group) = c("group 1", "group 2", "group 3")

# How many cells by group ?
table(sobj$kmeans_group)

# Visualization
Seurat::DimPlot(sobj,
                reduction = name2D,
                group.by = "kmeans_group",
                label = TRUE,
                cols = aquarius::gg_color_hue(3)) +
  Seurat::NoAxes() + Seurat::NoLegend() +
  ggplot2::theme(aspect.ratio = 1)
```

We can also define a lot of groups :

```{r clustering, fig.width = 8, fig.height = 8}
# kmeans
kmeans_obj = sobj@reductions[["mnn"]]@cell.embeddings %>%
  as.data.frame()

set.seed(1337L)
kmeans_obj = stats::kmeans(kmeans_obj, centers = 10, nstart = 10, 
                           algorithm = "Hartigan-Wong", iter.max = 500)

# add column in method data
sobj$small_clusters = kmeans_obj$cluster[colnames(sobj)]

# we remove a contaminating cluster
sobj$small_clusters = ifelse(sobj$seurat_clusters == contam_clust,
                             yes = NA, no = sobj$small_clusters)

# relevel according to pseudotime
small_clusters_levels = sobj@meta.data[, c("small_clusters", "pseudotime")] %>%
  dplyr::group_by(small_clusters) %>%
  dplyr::summarise(avg_pdt = mean(pseudotime)) %>%
  dplyr::arrange(avg_pdt) %>%
  dplyr::pull(small_clusters)

# replace plexiform as cluster 0
sobj$small_clusters = ifelse(sobj$sample_type == "flfl_pNF" & sobj$seurat_clusters != contam_clust,
                             yes = 0,
                             no = sobj$small_clusters)

sobj$small_clusters = factor(sobj$small_clusters,
                             levels = c(0, small_clusters_levels))
levels(sobj$small_clusters) = c(0:(length(small_clusters_levels)-1))

# How many cells by group ?
table(sobj$small_clusters, useNA = "ifany")

# Visualization
Seurat::DimPlot(sobj,
                reduction = name2D,
                group.by = "small_clusters",
                label = TRUE,
                cols = aquarius::gg_color_hue(11)) +
  Seurat::NoAxes() + Seurat::NoLegend() +
  ggplot2::theme(aspect.ratio = 1)
```

# Annotation preparation

In this section, we prepare the gene annotation. We load gene sets of interest from MSigDB.

Gene annotation is stored in the object :

```{r gene_corresp}
gene_corresp = sobj@assays[["RNA"]]@meta.features[, c("gene_name", "Ensembl_ID")]
names(gene_corresp) = c("NAME", "ID")

head(gene_corresp)
```


We get signatures from MSigDB :

```{r msigdb_sets}
gene_sets = msigdbr::msigdbr(species = "Mus musculus") %>%
  dplyr::filter(gs_cat %in% c("H", "C2", "C5")) %>%
  dplyr::filter(gs_subcat %in% c("", "GO:BP", "GO:MF", "GO:CC", "CP:KEGG", "CP:BIOCARTA",
                                 "CP:WIKIPATHWAYS", "CP:REACTOME", "CP:PID")) %>%
  dplyr::select(gs_cat, gs_subcat, gs_name, ensembl_gene, gene_symbol)

head(gene_sets)

table(gene_sets$gs_subcat, gene_sets$gs_cat)
```

# Between groups

In this section, we analyse the three groups. We change the active identifier :

```{r kmeans_group_kmeans_ident}
Seurat::Idents(sobj) = sobj$kmeans_group

table(sobj$sample_identifier,
      Seurat::Idents(sobj)) %>%
  rbind(., table(Seurat::Idents(sobj)))
```

We store results in the `list_result` list :

```{r initialization}
list_results = list()
```

## Group 1 vs Group 3

```{r see_blue_red_group1, fig.width = 5, fig.height = 5}
group1 = "group 1"
group2 = "group 3"

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r de_group1, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r foldchange_group1}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```

To what extent Ensembl ID are available ?

```{r venn_ensembl, fig.width = 8, fig.height = 5}
ggvenn::ggvenn(list(our_ensembl_id = names(ranked_gene_list),
                    msigdb_ensembl_id = gene_sets$ensembl_gene))
```


### Up in group 1

We explore enrichment in gene sets :

```{r ora_group1, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r gsea_plot_full_group1, fig.width = 12, fig.height = 30}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

We make the GSEA plot for some signaling pathways :

```{r plot_gsea_group1, fig.width = 12, fig.height = 10}
gs_oi = c("GOBP_ENSHEATHMENT_OF_NEURONS",
          "GOBP_RESPONSE_TO_AXON_INJURY",
          "WP_NEUROINFLAMMATION_AND_GLUTAMATERGIC_SIGNALING",
          "GOBP_AXON_DEVELOPMENT")

plot_list = lapply(gs_oi, FUN = function(one_gene_set) {
  NES = gsea_results@result[one_gene_set, "NES"]
  p.adjust = gsea_results@result[one_gene_set, "p.adjust"]
  qvalues = gsea_results@result[one_gene_set, "qvalues"]
  
  my_subtitle = paste0("\nNES : ", round(NES, 2),
                       " | padj : ", round(p.adjust, 4),
                       " | qval : ", round(qvalues, 4))
  
  p = enrichplot::gseaplot2(gsea_results, one_gene_set) +
    ggplot2::labs(title = one_gene_set,
                  subtitle = my_subtitle) +
    ggplot2::theme(plot.title = element_text(hjust = 0.5),
                   plot.subtitle = element_text(hjust = 0.5, size = 10))
  return(p)
})

patchwork::wrap_plots(plot_list, ncol = 2)
```


### Up in group 3

We explore enrichment in gene sets :

```{r ora_group3, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r gsea_plot_full_group3, fig.width = 12, fig.height = 70}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

We make the GSEA plot for some signaling pathways :

```{r plot_gsea_group3, fig.width = 12, fig.height = 10}
gs_oi = c("GOBP_BONE_MINERALIZATION",
          "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",
          "HALLMARK_TGF_BETA_SIGNALING",
          "GOMF_EXTRACELLULAR_MATRIX_STRUCTURAL_CONSTITUENT")

plot_list = lapply(gs_oi, FUN = function(one_gene_set) {
  NES = gsea_results@result[one_gene_set, "NES"]
  p.adjust = gsea_results@result[one_gene_set, "p.adjust"]
  qvalues = gsea_results@result[one_gene_set, "qvalues"]
  
  my_subtitle = paste0("\nNES : ", round(NES, 2),
                       " | padj : ", round(p.adjust, 4),
                       " | qval : ", round(qvalues, 4))
  
  p = enrichplot::gseaplot2(gsea_results, one_gene_set) +
    ggplot2::labs(title = one_gene_set,
                  subtitle = my_subtitle) +
    ggplot2::theme(plot.title = element_text(hjust = 0.5),
                   plot.subtitle = element_text(hjust = 0.5, size = 10))
  return(p)
})

patchwork::wrap_plots(plot_list, ncol = 2)
```

## Group 1 vs Group 2

```{r see_blue_red_group1_2, fig.width = 5, fig.height = 5}
group1 = "group 1"
group2 = "group 2"

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r de_group1_2, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r foldchange_group1_2}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```


### Up in group 1

We explore enrichment in gene sets :

```{r ora_group1_2_up, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r gsea_plot_full_group1_2_up, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in group 2

We explore enrichment in gene sets :

```{r ora_group1_2_down, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r gsea_plot_full_group1_2_down, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

## Group 3 vs Group 2

```{r see_blue_red_group3_2, fig.width = 5, fig.height = 5}
group1 = "group 3"
group2 = "group 2"

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r de_group3_2, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r foldchange_group3_2}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```


### Up in group 3

We explore enrichment in gene sets :

```{r ora_group3_2_up, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r gsea_plot_full_group3_2_up, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in group 2

We explore enrichment in gene sets :

```{r ora_group3_2_down, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r gsea_plot_full_group3_2_down, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```


# Between clusters

We are going to perform functional analysis between clusters, along pseudotime. These are the clusters :

```{r see_all_clusters, fig.width = 8, fig.height = 8}
Seurat::Idents(sobj) = sobj$small_clusters

custom_cols = aquarius::gg_color_hue(length(levels(sobj$small_clusters)))
names(custom_cols) = levels(sobj$small_clusters)

Seurat::DimPlot(sobj,
                group.by = "small_clusters",
                reduction = name2D,
                label = TRUE,
                cols = custom_cols) +
  Seurat::NoLegend() + Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1)
```

We will compare all clusters, in the pseudotime order :

```{r clusters_in_pdt_order}
comparison = data.frame(id1 = c(0, 1, 3, 5, 3, 6, 7, 8, 9, 8),
                        id2 = c(1, 3, 5, 6, 6, 7, 8, 9, 10, 10),
                        stringsAsFactors = FALSE)
comparison
```


```{r init_id, echo = FALSE}
id = 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

```{r, echo = FALSE}
id = id + 1
```

## Between `r comparison[id, 1]` and `r comparison[id, 2]`

```{r, fig.width = 5, fig.height = 5}
group1 = comparison[id, 1]
group2 = comparison[id, 2]

group_name = paste0(paste0(group1, collapse = "_"), "_vs_",
                    paste0(group2, collapse = "_"))

aquarius::plot_red_and_blue(sobj, group1, group2, reduction = name2D)
```

We run the differential expression analysis :

```{r, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj,
                           ident.1 = group1,
                           ident.2 = group2)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)
list_results[[group_name]]$de_genes = mark

dim(mark)

mark %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::top_n(n = 20, wt = abs(avg_logFC))
```

For GSEA, we compute the fold change between both groups :

```{r}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% group1],
                                            group2 = colnames(sobj)[sobj@active.ident %in% group2])
names(ranked_gene_list) = gene_corresp[names(ranked_gene_list), "ID"]

head(ranked_gene_list)
```


### Up in `r group1`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group1, 
                                                            " compared to clusters ",
                                                            paste0(group2, collapse = ", ")))

list_results[[group_name]]$enrichr_up = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_up = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

### Up in `r group2`

We explore enrichment in gene sets :

```{r, fig.width = 12, fig.height = 8}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = paste0("Up-regulated in cluster ", group2, 
                                                            " compared to clusters ",
                                                            paste0(group1, collapse = ", ")))

list_results[[group_name]]$enrichr_down = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


We run a GSEA for each significant gene sets :

```{r, fig.width = 12, fig.height = 5}
gene_sets_of_interest = enrichr_results$ego@result %>%
  dplyr::filter(p.adjust < 0.05) %>%
  dplyr::pull(ID)
gs_content = gene_sets[gene_sets$gs_name %in% gene_sets_of_interest, ] %>%
  dplyr::select(gs_name, ensembl_gene) %>%
  `colnames<-`(c("gs_name", "entrez_gene"))

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gs_content,
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea_down = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(n = 20, wt = -pvalue) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

# Save

We save the Seurat object, containing cluster ID.

```{r save_sobj}
saveRDS(sobj, file = paste0(out_dir, "/", save_name, "_sobj_traj_tinga_FE.rds"))
```

We save the `list_results` object :

```{r save_list_results}
saveRDS(list_results, file = paste0(out_dir, "/", save_name, "_list_results.rds"))
```

We also save it in XLSX format, for only DE genes :

```{r save_list_results_xlsx}
list_results = lapply(list_results, FUN = function(elem) {
  the_df = as.data.frame(elem[1])
  colnames(the_df) = c("p_val", "avg_logFC", "pct.1", "pct.2", "p_val_adj")
  
  return(the_df)})

openxlsx::write.xlsx(x = list_results,
                     file = paste0(out_dir, "/", save_name, "_list_results.xlsx"),
                     col.names  = TRUE,
                     row.names = TRUE)
```


# R session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```
