---
title: "Mice"
subtitle: "Tumor cells GSEA heatmap"
author: "Audrey"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <<- Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


In the previous file, we explored gene sets enrichment in tumor cells. In this file, we attribute a score to each cell, for each "significant" gene set, and we visualize the result on a heatmap.


```{r library}
library(dplyr)
library(org.Mm.eg.db)
library(ComplexHeatmap)
library(ggplot2)

.libPaths()
```

# Preparation

In this section, we set the global settings of the analysis. Data are stored there :

```{r out_dir}
out_dir = "."
n_threads = 1 # for AUCell::AUCell_buildRankings
```

Note : Not more than `nCores = 2` because it will reaches the maximum of memory and makes RStudio crash :(

We load the sample information :

```{r custom_palette_sample, fig.width = 8, fig.height = 8}
sample_info = readRDS(paste0(out_dir, "/../1_metadata/mice_sample_info.rds"))
project_names_oi = sample_info$project_name

graphics::pie(rep(1, nrow(sample_info)),
              col = sample_info$color,
              labels = sample_info$project_name)
```

We load the Seurat object :

```{r load_sobj}
save_name = "mice_tumor_cells_subset"
sobj = readRDS(paste0(out_dir, "/", save_name, "_sobj_traj_tinga_FE.rds"))
sobj
```

We load the list containing all enrichment results :

```{r load_results}
list_results = readRDS(paste0(out_dir, "/", save_name, "_list_results.rds"))

names(list_results)
```

We get signatures from MSigDB :

```{r msigdb_sets}
gene_sets = msigdbr::msigdbr(species = "Mus musculus") %>%
  dplyr::filter(gs_cat %in% c("H", "C2", "C5")) %>%
  dplyr::filter(gs_subcat %in% c("", "GO:BP", "GO:MF", "GO:CC", "CP:KEGG", "CP:BIOCARTA",
                                 "CP:WIKIPATHWAYS", "CP:REACTOME", "CP:PID")) %>%
  dplyr::select(gs_cat, gs_subcat, gs_name, ensembl_gene, gene_symbol)

head(gene_sets)

gene_sets %>%
  dplyr::select(gs_subcat, gs_cat, gs_name) %>%
  dplyr::distinct() %>%
  dplyr::select(gs_subcat, gs_cat) %>%
  table()
```

# Heatmap of gene sets expression score

## Gene sets of interest

We keep maximum 200 gene sets significant by GSEA in each comparison.

```{r choosen_gs_name_heatmap}
choosen_gs_name = lapply(names(list_results), FUN = function(comparison) {
  comparison = list_results[[comparison]]
  
  gs_names_up = comparison$gsea_up@result %>%
    dplyr::filter(pvalue < 0.05) %>%
    dplyr::top_n(., n = 100, wt = abs(NES)) %>%
    dplyr::pull(ID)
  
  gs_names_down = comparison$gsea_down@result %>%
    dplyr::filter(pvalue < 0.05) %>%
    dplyr::top_n(., n = 100, wt = abs(NES)) %>%
    dplyr::pull(ID)
  
  gs_names = c(gs_names_up, gs_names_down)
  return(gs_names)
}) %>% unlist() %>% unique()

length(choosen_gs_name)
```

How many in each category ?

```{r bycategory}
length(grep(choosen_gs_name, pattern = "^HALLMARK"))
length(grep(choosen_gs_name, pattern = "^GOBP"))
length(grep(choosen_gs_name, pattern = "^GOCC"))
length(grep(choosen_gs_name, pattern = "^GOMF"))
length(grep(choosen_gs_name, pattern = "^REAC"))
length(grep(choosen_gs_name, pattern = "^KEGG"))
length(grep(choosen_gs_name, pattern = "^BIOC"))
length(grep(choosen_gs_name, pattern = "^PID"))
length(grep(choosen_gs_name, pattern = "^WP"))
```

What are the genes in each gene set ?

```{r gene_sets_list}
gene_sets_list = lapply(choosen_gs_name, FUN = function(my_gs_name) {
  gene_sets %>%
    dplyr::filter(gs_name == my_gs_name) %>%
    dplyr::pull(gene_symbol)
})
names(gene_sets_list) = choosen_gs_name
```

We can make one heatmap for each category. The only thing which change is the "expression matrix". So, we can prepare all the annotation around the heatmap.

```{r heatmap_annotation, class.source = "fold-hide"}
## Cell annotation
cell_annot_1 = "small_clusters"
cell_annot_1_levels = levels(sobj@meta.data[, cell_annot_1])
sobj@meta.data[, cell_annot_1] = factor(sobj@meta.data[, cell_annot_1],
                                        levels = cell_annot_1_levels)

cell_annot_2 = "kmeans_group"
cell_annot_2_levels = c("group 1", "group 2", "group 3")
sobj@meta.data[, cell_annot_2] = factor(sobj@meta.data[, cell_annot_2],
                                        levels = cell_annot_2_levels)

## Colors
list_colors = list()

list_colors[["clusters"]] = aquarius::c30_palette[c(1:length(levels(sobj@meta.data[, cell_annot_1])))]
names(list_colors[["clusters"]]) = levels(sobj@meta.data[, cell_annot_1])

list_colors[["groups"]] = aquarius::c30_palette[c(1:length(levels(sobj@meta.data[, cell_annot_2])))]
names(list_colors[["groups"]]) = levels(sobj@meta.data[, cell_annot_2])

list_colors[["expression"]] = rev(RColorBrewer::brewer.pal(name = "RdBu", n = 9))
list_colors[["pseudotime"]] = circlize::colorRamp2(breaks = seq(min(sobj$pseudotime),
                                                                max(sobj$pseudotime),
                                                                length.out = 9),
                                                   colors = viridis::viridis(9))
list_colors[["sample_type"]] = c(flfl_pNF = "blue", flfl_dyNF = "green", flfl_MPNST = "brown",
                                 flmn_dyNF = "olivedrab1", flmn_MPNST = "sienna")

list_colors[["cell_cycle_phase"]] = aquarius::gg_color_hue(3)
names(list_colors[["cell_cycle_phase"]]) = c("G1", "G2M", "S")
list_colors[["percent_mito"]] = circlize::colorRamp2(breaks = seq(from = min(sobj$percent.mt),
                                                                  to = max(sobj$percent.mt),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
list_colors[["percent_ribo"]] = circlize::colorRamp2(breaks = seq(from = min(sobj$percent.rb),
                                                                  to = max(sobj$percent.rb),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
list_colors[["nFeature_RNA"]] = circlize::colorRamp2(breaks = seq(from = min(sobj$nFeature_RNA),
                                                                  to = max(sobj$nFeature_RNA),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
list_colors[["log_nCount_RNA"]] = circlize::colorRamp2(breaks = seq(from = min(sobj$log_nCount_RNA),
                                                                    to = max(sobj$log_nCount_RNA),
                                                                    length.out = 9),
                                                       colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))

# Cells order
column_order = order(sobj$pseudotime)

# Heatmap
ha_top = HeatmapAnnotation(pseudotime = sobj$pseudotime,
                           groups = sobj@meta.data[, cell_annot_2],
                           clusters = sobj@meta.data[, cell_annot_1],
                           col = list(pseudotime = list_colors[["pseudotime"]],
                                      groups = list_colors[["groups"]],
                                      clusters = list_colors[["clusters"]]))

ha_bottom = HeatmapAnnotation(percent_mito = sobj$percent.mt,
                              percent_ribo = sobj$percent.rb,
                              nFeature_RNA = sobj$nFeature_RNA,
                              log_nCount_RNA = sobj$log_nCount_RNA,
                              cell_cycle_phase = sobj$Seurat.Phase,
                              sample_type = factor(sobj$sample_type,
                                                   levels = names(list_colors[["sample_type"]])),
                              col = list(sample_type = list_colors[["sample_type"]],
                                         percent_mito = list_colors[["percent_mito"]],
                                         percent_ribo = list_colors[["percent_ribo"]],
                                         nFeature_RNA = list_colors[["nFeature_RNA"]],
                                         log_nCount_RNA = list_colors[["log_nCount_RNA"]],
                                         cell_cycle_phase = list_colors[["cell_cycle_phase"]]))

# Function to make the heatmap from the expression matrix
make_ht = function(mat_expr, ...) {
  ht = Heatmap(as.matrix(mat_expr),
               heatmap_legend_param = list(title = "score", at = c(0, 1), 
                                           labels = c("low", "high")),
               col = list_colors[["expression"]],
               top_annotation = ha_top, bottom_annotation = ha_bottom,
               show_column_names = FALSE,
               column_order = column_order,
               column_gap = unit(2, "mm"),
               cluster_rows = FALSE,
               row_title = NULL,
               row_names_gp = grid::gpar(fontsize = 10, fontface = "plain"),
               row_names_max_width = ComplexHeatmap::max_text_width(rownames(mat_expr),
                                                                    gp = gpar(fontsize = 10)),
               use_raster = FALSE,
               show_heatmap_legend = TRUE,
               border = TRUE,
               ...)
  return(ht)
}
```

## Compute scores for each gene sets

We will compute scores for each gene set with `AUCell`. To do so, we compute a gene ranking, for all cells and all genes in the dataset.

```{r aucell_rank, time_it = TRUE}
gene_expression_matrix = sobj@assays[["RNA"]]@data
aucell_rank = AUCell::AUCell_buildRankings(exprMat = gene_expression_matrix,
                                           nCores = n_threads, plotStats = FALSE)
```

Compute scores :

```{r score_gene_sets, time_it = TRUE}
df_scores = AUCell::AUCell_calcAUC(geneSets = gene_sets_list,
                                   rankings = aucell_rank,
                                   nCores = n_threads)
df_scores = AUCell::getAUC(df_scores) # AUCell object to scores matrix (features x cells)
```

There are missing gene sets :

```{r missing_gene_sets}
names(gene_sets_list)[!(names(gene_sets_list) %in% rownames(df_scores))]
```

We remove them :

```{r sub_gene_sets_list}
gene_sets_list = gene_sets_list[rownames(df_scores)]

length(gene_sets_list) == nrow(df_scores)
```



## Identify a gene set order

We identify an order to represent gene sets along pseudotime.

```{r heatmaps_order, time_it = TRUE}
sobj$cell_CB = colnames(sobj)

## Add a pseudotime column to df_scores
df_features = as.data.frame(t(df_scores)) # cells x features
df_features$cell_CB = rownames(df_features)
df_features = dplyr::left_join(x = df_features,
                               y = sobj@meta.data[, c("cell_CB", "pseudotime")],
                               by = "cell_CB")
df_features$cell_CB = NULL

## Identify order
features_order = aquarius::find_features_order(df_features)
```


## Build the matrix

We prepare the matrix to build the heatmap without row neither column ordering. To do so :

* Cells are ordered by pseudotime
* Features are ordered as previously defined
* Feature score are normalized between 0 and 1 for each feature, to have blue-to-red scale for each row (feature) in the heatmap

```{r mat_expression, class.source = "fold-hide", , fig.width = 15, fig.height = 20}
features_order = features_order[features_order %in% rownames(df_scores)]
mat_expression = df_scores[features_order, ]
mat_expression = Matrix::t(mat_expression)
mat_expression = dynutils::scale_quantile(mat_expression) # between 0 and 1
mat_expression = Matrix::t(mat_expression)
dim(mat_expression)
```


## Draw heatmaps

We have a big amount of gene sets, so we split into categories.

### Hallmark

```{r heatmap_hallmark, class.source = "fold-hide", fig.width = 15, fig.height = 12}
features_order_subset = features_order[which(features_order %in% grep(choosen_gs_name,
                                                                      pattern = "^HALLMARK", value = TRUE))]
sub_mat_expression = mat_expression[features_order_subset, ]

rownames(sub_mat_expression) = lapply(features_order_subset,
                                      FUN = substr,
                                      start = nchar("HALLMARK_") + 1,
                                      stop = 50 + nchar("HALLMARK_"))
ht = make_ht(sub_mat_expression)

ComplexHeatmap::draw(ht,
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


### GO : Biological Processes

```{r heatmap_gobp, class.source = "fold-hide", fig.width = 15, fig.height = 100}
features_order_subset = features_order[which(features_order %in% grep(choosen_gs_name,
                                                                      pattern = "^GOBP", value = TRUE))]
sub_mat_expression = mat_expression[features_order_subset, ]
rownames(sub_mat_expression) = lapply(features_order_subset,
                                      FUN = substr,
                                      start = nchar("GOBP_") + 1,
                                      stop = 50 + nchar("GOBP_"))
ht = make_ht(sub_mat_expression)

ComplexHeatmap::draw(ht,
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


### GO : Molecular Functions

```{r heatmap_gomf, class.source = "fold-hide", fig.width = 15, fig.height = 20}
features_order_subset = features_order[which(features_order %in% grep(choosen_gs_name,
                                                                      pattern = "^GOMF", value = TRUE))]
sub_mat_expression = mat_expression[features_order_subset, ]
rownames(sub_mat_expression) = lapply(features_order_subset,
                                      FUN = substr,
                                      start = nchar("GOMF_") + 1,
                                      stop = 50 + nchar("GOMF_"))
ht = make_ht(sub_mat_expression)

ComplexHeatmap::draw(ht,
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


### GO : Cellular Components

```{r heatmap_gocc, class.source = "fold-hide", fig.width = 15, fig.height = 20}
features_order_subset = features_order[which(features_order %in% grep(choosen_gs_name,
                                                                      pattern = "^GOCC", value = TRUE))]
sub_mat_expression = mat_expression[features_order_subset, ]
rownames(sub_mat_expression) = lapply(features_order_subset,
                                      FUN = substr,
                                      start = nchar("GOCC_") + 1,
                                      stop = 50 + nchar("GOCC_"))
ht = make_ht(sub_mat_expression)

ComplexHeatmap::draw(ht,
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


### KEGG Pathways

```{r heatmap_kegg, class.source = "fold-hide", fig.width = 15, fig.height = 10}
features_order_subset = features_order[which(features_order %in% grep(choosen_gs_name,
                                                                      pattern = "^KEGG", value = TRUE))]
sub_mat_expression = mat_expression[features_order_subset, ]
rownames(sub_mat_expression) = lapply(features_order_subset,
                                      FUN = substr,
                                      start = nchar("KEGG_") + 1,
                                      stop = 50 + nchar("KEGG_"))
ht = make_ht(sub_mat_expression)

ComplexHeatmap::draw(ht,
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```

### BioCarta

```{r heatmap_biocarta, class.source = "fold-hide", fig.width = 15, fig.height = 10}
features_order_subset = features_order[which(features_order %in% grep(choosen_gs_name,
                                                                      pattern = "^BIOCARTA", value = TRUE))]
sub_mat_expression = mat_expression[features_order_subset, ]
rownames(sub_mat_expression) = lapply(features_order_subset,
                                      FUN = substr,
                                      start = nchar("BIOCARTA_") + 1,
                                      stop = 50 + nchar("BIOCARTA_"))
ht = make_ht(sub_mat_expression)

ComplexHeatmap::draw(ht,
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```

### Wiki Pathways

```{r heatmap_wp, class.source = "fold-hide", fig.width = 15, fig.height = 20}
features_order_subset = features_order[which(features_order %in% grep(choosen_gs_name,
                                                                      pattern = "^WP", value = TRUE))]
sub_mat_expression = mat_expression[features_order_subset, ]
rownames(sub_mat_expression) = lapply(features_order_subset,
                                      FUN = substr,
                                      start = nchar("WP_") + 1,
                                      stop = 50 + nchar("WP_"))
ht = make_ht(sub_mat_expression)

ComplexHeatmap::draw(ht,
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


### Reactome

```{r heatmap_reactome, class.source = "fold-hide", fig.width = 15, fig.height = 45}
features_order_subset = features_order[which(features_order %in% grep(choosen_gs_name,
                                                                      pattern = "^REACTOME", value = TRUE))]
sub_mat_expression = mat_expression[features_order_subset, ]
rownames(sub_mat_expression) = lapply(features_order_subset,
                                      FUN = substr,
                                      start = nchar("REACTOME_") + 1,
                                      stop = 50 + nchar("REACTOME_"))
ht = make_ht(sub_mat_expression)

ComplexHeatmap::draw(ht,
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


### Pathways Interaction (PID)

```{r heatmap_pid, class.source = "fold-hide", fig.width = 15, fig.height = 10}
features_order_subset = features_order[which(features_order %in% grep(choosen_gs_name,
                                                                      pattern = "^PID", value = TRUE))]
sub_mat_expression = mat_expression[features_order_subset, ]
rownames(sub_mat_expression) = lapply(features_order_subset,
                                      FUN = substr,
                                      start = nchar("PID_") + 1,
                                      stop = 50 + nchar("PID_"))
ht = make_ht(sub_mat_expression)

ComplexHeatmap::draw(ht,
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```

# Save

We save the dataframe containing score for each gene set :

```{r save_df_scores}
saveRDS(df_scores, file = paste0(out_dir, "/", save_name, "_df_scores.rds"))
```


# R session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```


